### 2018-05-23 abstract.tex
Add all source files to repo. This is the first draft.

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in BLAS library.
  To achieve best performance, in-memory data must be
  prefetched into fast on-chip caches ahead of use.
  Two techniques, software prefetching and data packing,
  are utilized to effectively exploit the capability of on-chip caches.
  These techniques are sufficient for lastly-recent-used (LRU) caches,
  which is often the case in traditional high-performance
  processors used in high-end servers and supercomputers.
  But in recent years the market is meeting a diversity
  in processor design, and some high-performance processors
  are equipped with shared caches with non-LRU replacement policies.
  This rises a challenge to GEMM development in multi-threaded context.
  As several threads try to load data into the shared cache simultaneously,
  inter-thread cache conflicts would increase significantly.
  We present a Shared Cache Partition (SCP) method to
  eliminate inter-thread cache conflicts in GEMM routines,
  by partitioning the shared cache into physically disjoint
  sets and assign different sets to different threads.
  We implement SCP in the OpenBLAS library and
  evaluate the performance on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared (by every 4 cores) psudo-random L2 caches.
  The results show that conflict misses on both L1 and L2
  caches are effectively reduced and GEMM performance get improved
  by $2.75\%$--$6.91\%$.
  %% FIXME BLIS? FT1500A?
\end{abstract}

### 2018-05-23 background.tex
Add all source files to repo. This is the first draft.

\section{Background}\label{sec:background}

\subsection{The Memory Hierachy Review}\label{subsec:hierachy}

Modern architectures take advantage of fast on-chip caches
to overcome the ``memory wall''.
As a compromise between direct mapped cache and full-associative cache,
set-associative cache has been the dominant choice for real-world architectures.
To understand how shared cache can cause inter-thread cache conflicts
and do harm to GEMM performance, we review some details
about the working mechanisim of set-associative caches.

A set-associative cache can be defined by a 4-tuple $(c, l, n, nt)$,
in which $c$ is the cache capacity, $l$ is the cache line size,
$n$ is ways-of-associativity,
and $nt$ means the number of cores sharing the cache.
The whole cache memory is divided into $ns=c/l/n$ sets,
denoted by integral indices $\mathcal{S} = [0, ns)$,
with each set containing $n$ cache lines.
Generally, $ns$ is a power-of-two value.
An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
mapping addresses from the address space $\mathcal{A}$
(either virtual or physical) to cache sets $\mathcal{S}$.
Data at $addr \in \mathcal{A}$ must be fetched
into the cache set $\varphi(addr) \in \mathcal{S}$ if requested.
A classic implementation of the indexing function is shown below,
%% $\varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)$,
with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively.
It is an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$
position in $addr$'s binary form.
\begin{equation*}
  \varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)
  \label{eq:phi}
\end{equation*}

It is easy to prove that any continous memory block no larger than $c/n$
cannot conflict with itself.
We define a quantity, way-capacity, denoted as $wc=c/n$,
to represent this size bound for conflict-free memory blocks.

The main memory, $NL$ levels of caches, and $NT$ processor cores
form a memory hierachy generally organized into
a $NL+2$ level regular tree structure.
Processor cores at level $0$ are leaves,
caches at levels $1$ -- $NL$ are intermedia nodes
and the main memory at level $NL+1$ is the tree root.
Any node in the memory hierachy can be identified by
its layer and its unique index on the layer $(layer, index)$.
The number of nodes on layer $l$ is $NT / nt_l$.

Throughout this paper we will use a 4-core processor with
$NL=2$ levels of caches as the platform for SCP demonstration.
Figure.~\ref{fig:hierachy} shows its memory hierachy and
detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
Each core has 32 128-bit (16B) vector registers,
each capable of store 2 double precision floating-point numbers.
To make a consistency, the processor cores on level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with programmable replacement policy.

\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{figures/cluster-new}
  \caption{Memory hierachy of the 4-core processor
    (nodes labeled with $(layer,index)$)}
  \label{fig:hierachy}
\end{figure}

\begin{table}
  \centering
  \caption{Architectural parameters of the 4-core processor}
  %% (see Section~\ref{subsec:hierachy} for meanings of symbols in the header row)
  \label{tab:cluster}
  \begin{tabular}{lccccccl}
    \toprule
    type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & policy \\
    \midrule
    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & programmable \\
    L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & psudo-random \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Structure of GEMM}\label{subsec:gemm}

GEMM performs a matrix-multiply-accumulate operation $C = \beta C + \alpha A B$,
where $A, B$ and $C$ are matrices of sizes
$M \times K$, $K \times N$ and $M \times N$, respectively,
and $\alpha$, $\beta$ are scalars.
While this operation is algorithmically simple,
so that a 3-deep loop nest suffices to accomplish it,
a high-performance implementation can be quite
sophisticated due to the presence of multi-level memory
hierarchies on modern processors.
Figure.~\ref{fig:gemm} shows the structure of GEMM program from
the OpenBLAS library~\cite{openblas}.
Each loop (in the original 3-deep loop nest) is tiled,
resulting in a total of six loops (referred to as layers 1 -- 6).
%% Loop tiling, together with data packing and prefetching,
%% serves to improve data locality and overlap computation
%% and memory access effectively.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\textwidth]{figures/gemm}
  \caption{Structure of blocked GEMM}
  \label{fig:gemm}
\end{figure*}

In this blocked algorithm, the loops over the $N$, $K$ and $M$
matrix dimensions are tiled by sizes $N_c$, $K_c$ and $M_c$ at 
layers 1, 2 and 3, respectively.
At layers 4 and 5, the $N$ and $M$ dimensions are further tiled by sizes
$N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ submatrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 need not to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices on each level fit into a certain level in the memory hierachy.
Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
the constraints on these tiling factors.
$es$ means the size of matrix element, e.g. 8B for double precision floating-point numbers,
and $c_l$ means the capacity of registers or caches on level $l$.
First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
(the psudo level-0 cache).
Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in next iteration
at layer 5 fit into the L1 cache.
Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in next iteration
at layer 4 fit into the L2 cache.
At last, by (\ref{eq:constraints.l3}), $N_c$ is constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
Larger matrices on higher layers 1 -- 3 live in main memory.

\begin{eqnarray}
  es (M_r + N_r + M_r N_r) & \le & c_{0} \label{eq:constraints.reg}\\
  nt_{1} \cdot es (N_r K_c + 2 M_r K_c) & \le & c_{1} \label{eq:constraints.l1}\\
  nt_{2} \cdot es (M_c K_c + 2 N_r K_c) & \le & c_{2} \label{eq:constraints.l2}\\
  es (N_c K_c + nt_{3} \cdot M_c K_c)   & \le & c_{3} \label{eq:constraints.l3}
\end{eqnarray}

\begin{comment}
Table~\ref{tab:factors} shows the procedure choosing various tiling factors.
In each row, one or more tiling factors (column 1) is determined
by a size constraint (column 3) related to a certain level
in the memory hierachy (column 2).
\begin{table}
  \centering
  \caption{Procedure of choosing tiling factors}
  \label{tab:factors}
  \begin{tabular}{cccl}
    \toprule
    factor & memory & constraint \\
    \midrule
    $M_r$, $N_r$ & registers & $M_r + N_r + M_r N_r \le Regs$ \\
    $K_c$        & L1 cache  & $N_r K_c + 2 M_r K_c \le L1$ \\
    $M_c$        & L2 cache  & $M_c K_c + 2 N_r K_c \le L2$ \\
    $N_c$        & L3 cache  & $N_c K_c + M_c K_c \le L3$ \\
    \bottomrule
  \end{tabular}
\end{table}
\end{comment}

Goto~\cite{gotogemm} factors out the innermost three loops at layers 4 -- 6 for
computing $C_2\ = C_2 + \alpha A_2 B_2$ as an architecture-dependent kernel,
known as  GEBP (GEneral multiply of a Block of $A$ and a Panel of $B$).
It is worthy noting that before GEBP is called (at layer 3),
$A_1[ii:ii+M_c-1][:]$ and $B_1$ are packed into $A_2$ and $B_2$
in a special continous layout,
so consecutive memory access is ensured within the GEBP kernel.
GotoBLAS~\cite{gotoblas}, and its successor, OpenBLAS~\cite{openblas},
implement GEMM programs based on this factorization,
with GEBP highly optimized for the target processor (often coded in assembly).

As stated in Section~\ref{sec:intro}, an GEMM implementation
is made up by a kernel routine and an overall strategy.
In Figure.~\ref{fig:gemm}, GEBP surves as the kernel routine,
and the overall strategy is defined by the three outter loops at layers 1 -- 3.
Developer can control how to partition the workload by choosing
tile factors $M_c$, $N_c$ and $K_c$,
and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
Either the N loop at layer 1 or M loop at layer 3, or both of them,
can be parallelized in a multi-threaded context.
Specifically, the strategy in Figure.~\ref{fig:gemm} chooses a NKM loop order and
parallelizes the M loop at layer 3.
Different threads work on different parts of $A_1$
and share the same $B_1$.
As a result, there are multiple $A_2$ instances, one per thread,
and a single $B_2$ instances, shared by all threads.
To understand how the loop on layer 3 is parallelized on the 4-core processor, 
Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
While all color shaded submatrices are accessed by $T_1$,
only z-curve masked submatrices (a subset of color shaded ones)
are packed by the thread.
That is, each thread packs its own $A_2$ instance and 
$B_2$ is packed by all thread collabratively.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

\begin{figure}[t]
  \centering
  \includegraphics[width=.4\textwidth]{figures/workload}
  \caption{GEMM parallelization with 4 threads and data accessing patterns of thread $T_1$}
  \label{fig:workload}
\end{figure}

\subsection{Cache Conflicts in GEMM}\label{subsec:cache-conflicts}

For set-associative caches, cache misses can be categorized
into capacity misses and conflict misses.
While constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
serve to avoid capacity misses,
conflict misses can not be eliminated completely.
%% We use the term ``cache conflict'' to represent conflict miss events.

Recall Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar that it is expected to
reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
In general, even with carefully selected tile factors,
matrix data can still be evicted out for a few reasons.
For instance, data structures other than the packed matrices,
or program code (in case of a unified cache) may be fetched into a
cache set the matrix data lives in.
We call this kind of cache conflicts intra-thread cache conflicts
because it occurs within the execution of a single thread.
To deal with intra-thread conflicts,
the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to L1 cache
in every iteration at layer 6 before it actually reads them in case of
the data being evicted out accidentally.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefething stratecy
can remedy the penalty caused by intra-thread conflicts.

In contrary to intra-thread cache conflicts,
there exists another kind of cache conflicts,
inter-thread cache conflicts, on architectures with shared caches.
Suppose two working threads, $T_0$ and $T_1$, are running independently
on the 4-core processor in Figure.~\ref{fig:hierachy}
to accomplish a GEMM operation.
At time $t_s$, $T_0$ prefetches some data at address $addr_0$
to cache set $\varphi_{L2}(addr_0)$ of the shared L2 cache.
The data will be actually read at time $t_e$.
Then at some time $t_m$ ($t_s < t_m < t_e$),
$T_1$ prefetches data at addr $addr_1$ to L2 cache.
If $\varphi_L(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,
data at $addr_1$ and $addr_0$ will be fetched to the same cache set.
As the shared L2 cache has a non-LRU policy,
data at $addr_0$ may suffer an eviction before it is actually read,
i.e. a conflict miss occurs.

In GEMM, the cache is utilized aggressively
that almost all cache space is expected to be occupied by packed matrices.
Prefetch instructions from different threads are executed repeatedly,
interleaved in an unpredictable manner.
So inter-thread cache conflicts described above
can occur frequently on non-LRU shared caches,
thus making the technique dealing with intra-thread cache conflicts helpless.
The situation goes worse with more threads sharing the same cache.

### 2018-05-23 evaluation.tex
Add all source files to repo. This is the first draft.

\section{Performance Evaluation}\label{sec:evaluation}

We implement SCP in the OpenBLAS~\cite{openblas} library version 0.3.0-dev,
and evaluate it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of the cluster is almost the same with
the contrived 4-core processor in Figure.~\ref{fig:hierachy}
except that all 16 clusters connect to the main memory
with their L2 caches coherented by hardare.
See Table.~\ref{tab:cluster} for architectural parameters of the cores and caches.
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers (FIXME citation),
relys on the DGEMM variant.

All evaluations run with the same tiling factors
$M_r$, $N_r$, $K_c$, $M_c$ and $N_c$.
The tiling factors are computed as in Section~\ref{subsec:example},
i.e. $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
The parallelism is controlled by two parameters,
the number of active clusters $NC$,
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \cdot NT_C$.
For the Phytium 2000+ processor, the valid ranges for $NC$ and $NT_C$
are $NC \in [1, 16]$ and $NT_C \in [1, 4]$, respectively.
Generally, DGEMM performance are presented in $flops$.
%% floating-point operations per second ($flops$).
As $NT$ varies in our evaluation,
we use another metric derived from $flops$,
the average thread efficiency,
to present performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \cdot \widehat{flops})$
in which $\widehat{flops}$ stands for the theoretical performance peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

We first show how cache sharing can harm GEMM performance
in Section~\ref{subsec:drawback}.
Then Section~\ref{subsec:benefit} demonstrates how SCP can remedy
the penalty caused by cache sharing.
Quantitative analysis of the cache miss rates
is presented in Section~\ref{subsec:analysis}.
Finally, Section~\ref{subsec:privb} discusses
the privatization of shared matrix $B_2$.
%% Three versions of DGEMM are evaluated:
%% (1) Base, the baseline version,
%% (2) SCP,  the SCP version, 
%% and (3) SCP-P, the SCP version plus privatization of shared matrix $B_2$.

\subsection{Penalty of Cache Sharing}\label{subsec:drawback}

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:drawback-4}
    \includegraphics[width=.31\textwidth]{figures/sec1-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:drawback-8}
    \includegraphics[width=.31\textwidth]{figures/sec1-8}
  }
  \\
  \subfigure[$NT=16$]{
    \label{fig:drawback-16}
    \includegraphics[width=.31\textwidth]{figures/sec1-16}
  }
  \subfigure[$NT=32,64$]{
    \label{fig:drawback-32}
    \includegraphics[width=.31\textwidth]{figures/sec1-32}
  }
  \caption{Average thread efficiency with cache sharing}
  \label{fig:drawback}
\end{figure}

As the L2 cache is shared by cores inside the cluster,
the impact of cache sharing can be demonstrated
by comparing DGEMM performance with the same $NT$
but different $NT_C$s.
Figure.~\ref{fig:drawback} shows the evaluation results.
Each curve represents a parallelism configuration.
Different subfigures are configured with different $NT$s,
and different curves in the same subfigure are
configured with different $NT_C$s.
As there is only one configuration for $NT=64$ ($NC=16$, $NT_C=4$)
and no comparation is available,
results of $NT=64$ is combined with $NT=32$ in Figure.~\ref{fig:drawback-32}.

\begin{table}
  \centering
  \caption{Average thread efficiency with cache sharing}
  \label{tab:drawback}
  \begin{tabular}{cccccc}
    \toprule
     & $NT\!\!=\!\!4$ & $NT\!\!=\!\!8$ & $NT\!\!=\!\!16$ & $NT\!\!=\!\!32$ & $NT\!\!=\!\!64$ \\
    \midrule
    $NT_C\!\!=\!\!1$ & 91.46 & 91.30 & 90.30 & -     & - \\   
    $NT_C\!\!=\!\!2$ & 90.12 & 90.25 & 89.49 & 88.26 & - \\
    $NT_C\!\!=\!\!4$ & 84.05 & 80.61 & 79.74 & 79.81 & 71.71 \\
    \bottomrule
  \end{tabular}
\end{table}

From Figure.~\ref{fig:drawback} we can see that
with all $NT$s, $NT_C=1$ achieves the best performance,
and $NT_C=2$ follows, and $NT_C=4$ comes at the lowest efficiency.
Table.~\ref{tab:drawback} summarizes the average value
of $E_{avg}$ over matrix sizes.
The table can be viewed in two directions.
Vertically, results in the same column shows how $E_{avg}$ varies with $NT_C$.
With all $NT$s, $NT_C=1$ outperforms $NT_C=4$ by a large margin.
The gap is $7.4\%$ when $NT=4$ and roughly $10\%$ when $NT>4$.
Horizontally, results in the same row reflects the scalability of
DGEMM with certain $NT_C$s.
With $NT_C=1$, the performance scales linearly with the number of threads.
With $NT_C=2$, $E_{avg}$ suffers a $1.86\%$ lose when $NT$ grows from 4 to 32.
The situation is much worse for $NT_C=4$, $E_{avg}$ drops by $14.19\%$ from
$NT=4$ to $NT=64$!

Evaluation results above show that cache sharing has a considerable
impact on DGEMM performance.

\subsection{Effectiveness of SCP}\label{subsec:benefit}

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:benefit-4}
    \includegraphics[width=.31\textwidth]{figures/sec2-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:benefit-8}
    \includegraphics[width=.31\textwidth]{figures/sec2-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:benefit-16}
    \includegraphics[width=.31\textwidth]{figures/sec2-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:benefit-32}
    \includegraphics[width=.31\textwidth]{figures/sec2-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:benefit-64}
    \includegraphics[width=.31\textwidth]{figures/sec2-64}
  }
  \caption{Average thread efficiency of SCP}
  \label{fig:benefit}
\end{figure}

We evaluate SCP using the same method as in Section~\ref{subsec:drawback}.
SCP is not evaluated with $NT_C=1$ because there is no need
to partition the cache if $NT_C=1$.
Figure.~\ref{fig:benefit} shows the evaluation results.
For the purpose of comparation, results without SCP
(denoted as Base in the figure) are also presented,

\begin{table}
  \centering
  \caption{Average thread efficiency win of SCP}
  \label{tab:win}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{cccccc}
    \toprule
     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
    \midrule
    $NT_C=2$ & 1.77 & 1.69 & 1.82 & 1.78 & - \\
    $NT_C=4$ & 2.75 & 4.62 & 4.41 & 3.11 & 6.91 \\
    \bottomrule
  \end{tabular}
\end{table}

%% \begin{table}
%%   \centering
%%   \caption{Average thread efficiency ($\%$) of SCP}
%%   \label{tab:benefit}
%%   \setlength{\tabcolsep}{3.5pt}
%%   \begin{tabular}{cccccc}
%%     \toprule
%%      & $NT\!\!=\!\!4$ & $NT\!\!=\!\!8$ & $NT\!\!=\!\!16$ & $NT\!\!=\!\!32$ & $NT\!\!=\!\!64$ \\
%%     \midrule
%%     $NT_C\!\!=\!\!2$ & 91.89/1.77 & 91.94/1.69 & 91.31/1.82 & 90.05/1.78 & - \\
%%     $NT_C\!\!=\!\!4$ & 86.80/2.75 & 85.22/4.62 & 84.15/4.41 & 82.92/3.11 & 78.62/6.91 \\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

Under all parallelism configurations, 
SCP performs better than the Base
consistently over all matrix sizes.
Table.~\ref{tab:win} summarizes the performance win of SCP over Base.
%% Table.~\ref{tab:benefit} summarizes the average value
%% of $E_{avg}$ in SCP evaluation.
%% There are two values in each cell, separated by a slash.
%% $E_{avg}$ stands on the left,
%% and performance win over conventional packing
%% in the corresponding parallelism configuration
%% is shown on the right for convenience.
With $NT_C=2$, SCP performs slightly better than Base
by $1.7\%$--$1.8\%$.
The performance gap becomes larger with $NT_C=4$.
ranging from $2.75\%$ to $6.91\%$.
The largest performance gain, $6.91\%$, which is a considerable improvement,
is observed under the maximum parallelism $NT=64$.
%% Looking at Table.~\ref{tab:benefit} vertically,
%% we can see that SCP still cannot fully eliminate the impact
%% of cache sharing.
%% But it do relieve the situation to a certain degree.

\subsection{Cache Miss Rate Analysis}\label{subsec:analysis}

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:papi-4}
    \includegraphics[width=.31\textwidth]{figures/papi-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:papi-8}
    \includegraphics[width=.31\textwidth]{figures/papi-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:papi-16}
    \includegraphics[width=.31\textwidth]{figures/papi-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:papi-32}
    \includegraphics[width=.31\textwidth]{figures/papi-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:papi-64}
    \includegraphics[width=.31\textwidth]{figures/papi-64}
  }
  \caption{Cache miss rates with $NT_C=4$}
  \label{fig:papi}
\end{figure}

To understand why SCP outperforms Base,
and whether SCP actually eliminate inter-thread conflicts,
we use the PAPI~\cite{papi} profiling tool to
analyze the cache behavior of the GEMM program.
Four cache related hardware events are counted,
which are listed and briefly described in Table.~\ref{tab:events}.
With these hardware events,
miss rates of L1 and L2 caches can be calculated with
$\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
The hardware events are mesured in a per-thread manner
and all presented results are the average over threads.

\begin{table}
  \centering
  \caption{Hardware events in GEMM}
  \label{tab:events}
  \begin{tabular}{lll}
    \toprule
    Name & Event & Description \\
    \midrule
    L1M & PAPI\_L1\_DCM & L1 data cache misses \\
    L1H & PAPI\_L1\_DCH & L1 data cache hits \\
    L2M & PAPI\_L2\_DCM & L2 data cache misses \\
    L2H & PAPI\_L2\_DCH & L2 data cache hits \\
    \bottomrule
  \end{tabular}
\end{table}

Figure.~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.
For the sake of space, we only present results with $NT_C=4$
as SCP achieves the largest improvement with full cache sharing.
Table.~\ref{tab:papi} summarizes the average value of cache miss rates
over matrix sizes.
Under all parallelism configurations,
both L1 and L2 cache miss rates are reduced with SCP applied.
The reduced amount is roughly $13\%$ for L1 and nearly $10\%$ for L2.

For the Phytium 2000+ processor whose L1 caches are private,
only L2 caches are specially handled by SCP.
From the results we see that miss rates on both L1 and L2 caches
are reduced effectively by SCP.
While the decrease of L2 miss rate is expected,
why L1 miss rate is also affected?
The reason is that the L2 cache is inclusive.
Conflict misses on the L2 cache not only evict data out from L2,
but also invalidate the evicted data's copy in all upper level L1 caches,
thus leading to more L1 cache misses.
In general, conflict misses on an inclusive cache can affect
performance of all upper level caches.

Results in this section prove that SCP can effectively
eliminate conflict misses on shared caches.
And the benefit propagates to upper level caches, too.

\begin{table}
  \centering
  \caption{Reduction of cache miss rates with $NT_C=4$}
  \label{tab:papi}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{lccccc}
    \toprule
     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
    \midrule
    L1     & 13.87 & 13.75 & 13.54 & 13.84 & 14.21 \\
    L2     & 10.00 & 10.00 & 8.27 & 8.56 & 7.62 \\
    \bottomrule
  \end{tabular}
\end{table}
%% \begin{table}
%%   \centering
%%   \caption{Cache miss rates ($\%$) with $NT_C=4$}
%%   \label{tab:papi}
%%   \setlength{\tabcolsep}{3.5pt}
%%   \begin{tabular}{lccccc}
%%     \toprule
%%      & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
%%     \midrule
%%     L1     & 7.81 & 7.83 & 7.87 & 7.88 & 7.88 \\
%%     L1 SCP & 6.73 & 6.75 & 6.81 & 6.81 & 6.81 \\
%%     L2     & 8.78 & 8.31 & 7.13 & 7.16 & 7.10 \\
%%     L2 SCP & 7.90 & 7.49 & 6.54 & 6.55 & 6.56 \\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

\subsection{Privatization of Shared Matrix}\label{subsec:privb}
In previous evaluations, SCP only set-partitions the shared cache
for $A_2$ matrices which are private to threads.
The shared $B_2$ matrix is still packed
in conventional way-partition style.
As mentioned in Section~\ref{subsec:example},
set-partition can also be applied to $B_2$ after $B_2$ is privatized.
In this section, we will discuss and evaluate this privatization alternative.

\begin{figure}
  %% FIXME draw this figure
  \centering
  \subfigure[Conventional]{
    \label{fig:packb.conventional}
    \includegraphics[width=.50\textwidth]{figures/strategy-conventional}
  }
  \subfigure[Full privatization]{
    \label{fig:packb.full}
    \includegraphics[width=.50\textwidth]{figures/strategy-full}
  }
  \subfigure[Partial privatization]{
    \label{fig:packb.partial}
    \includegraphics[width=.50\textwidth]{figures/strategy-partial}
  }
  \caption{Packing workload ($B_2$) of thread $T_1$ with $NT_C=4$ and $NC=4$}
  %% \caption{Packing strategies of shared matrix $B_2$ with $NT_C=4$ and $NC=4$}
  \label{fig:packb}
\end{figure}

Privatization inccurs more packing overhead because
redundant packing of $B_2$ is enforced.
Figure.~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelism configuration $NT_C=4$ and $NC=4$.
The whole packing workload of $B_2$ is divided into 16 tasks
and the packing workload of thread $T_1$ is masked by z-curves. 
%% and the task assignment is annoted in Figure.~\ref{fig:packb}.
In conventional packing, each thread takes a single task.
In full privatization, all tasks are replicated to all threads.
As a result, the packing overhead grows in proportion to $NT$
in full privatization strategy, which is unacceptable
if GEMM is highly parallelized, e.g. $NT=64$.
Partial privatization in Figure.~\ref{fig:packb.partial}
offers an alternative to full privatization.
In partial privatization, privatization only occurs inside
a cluster so that tasks belong to one cluster need not to be
replicated to other clusters.
Consequently, extra packing overhead of the partial privatization
strategy is bounded by $NT_C$.
For instance, on Phytium 2000+, overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in conventional packing.

Figure.~\ref{fig:privb} shows the evaluation results of the partial
privatization strategy with $NT_C=4$ and $NC=4$.
Results under other parallelism configurations are similar.
For the purpose of comparation,
SCP with and without privatization are both presented
(denoted as SCP-P and SCP in Figure.~\ref{fig:privb}, respectively).
Figure.~\ref{fig:privb.papi} shows the cache miss rates.
Privatization achieves slightly lower miss rates on both L1 and L2 caches.
Figure.~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower cache miss rates,
privatization suffers a performance lose compared to SCP.
Why performance drops while cache miss rates get reduced?
The reason lies in the extra packing overhead introduced by privatization.
Figure.~\ref{fig:privb.breakdown} shows the occupancy ratio
of overall executing time of various overheads.
Overheads in GEMM includes
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While SCP-P involves roughly the same overheads in
packing $A_2$ and synchronization as SCP,
it spends much more time in packing $B_2$.
Table.~\ref{tab:breakdown} lists the breakdown of
executing time. We can see that overhead of packing $B_2$
is multiplied by a factor of $NT_C=4$,
and the total overhead increases from $2.35\%$ to $4.72\%$.

\begin{figure}
  \centering
  \subfigure[Cache miss rates]{
    \label{fig:privb.papi}
    \includegraphics[width=.31\textwidth]{figures/privb-papi}
  }
  \subfigure[Average thread efficiency]{
    \label{fig:privb.ate}
    \includegraphics[width=.31\textwidth]{figures/privb-ate}
  }
  \subfigure[Executing time breakdown]{
    \label{fig:privb.breakdown}
    \includegraphics[width=.65\textwidth]{figures/privb-breakdown}
  }
  \caption{Performance of partial privatization with $NT_C=4$ and $NC=4$}
  \label{fig:privb}
\end{figure}

\begin{table}
  \centering
  \caption{Executing time breakdown with $NT_C=4$ and $NC=4$}
  \label{tab:breakdown}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{lcccc}
    \toprule
     & Sync & PackA & PackB & GEBP\\
    \midrule
    SCP   & 1.01 & 0.53 & 0.80 & 97.65 \\
    SCP-P & 1.00 & 0.52 & 3.20 & 95.28\\
    \bottomrule
  \end{tabular}
\end{table}

Results in this section demonstrats that
privatization of the shared matrix $B_2$ can further
eliminate inter-thread cache conflicts.
But the improvement in cache performance is very small because
$B_2$ only occupies a small portion of the shared L2 cache,
and this marginal improvement is not sufficient to
balance the extra packing overhead introduced
by privatization.

### 2018-05-23 intro.tex
Add all source files to repo. This is the first draft.

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS) (FIXME) defines a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into 3 levels,
level-1 for vector-vector operations,
level-2 for matrix-vetor operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
highly optimized for their processors,
such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
The HPC community have also contributed several high-quality
BLAS implementations, e.g. ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis}\cite{blisport}.

Among three levels, level-3 provides the most opportunities
for optimization because it has the highest computational complexity ($O(n^3)$).
Among all level-3 operations, GEneral Matrix Multiply (GEMM) is
of the most interest because all other level-3 operations can be defined
in terms of GEMM and some level-1 and level-2 operations.
Much effort have been spent on optimizing GEMM for different architectures,
both in general algorithm and in architecture specific ways.
(FIXME citations)

GEMM performs a matrix-multiply-accumulate operation.
Generally, an optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish the computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling the loops,
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In multi-threaded context, different tasks (loop tiles) are assigned to
different threads to parallelize the whole operation.
The main target of GEMM optimization is maximizing the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
On one hand, optimization of the serial kernel routine concentrates on
instruction scheduling to improve instruction throughput.
On the other hand, the overall strategy is responsible for optimizing
memory accesses to protect the kernel from data hunger.
In this paper, we focus on optimizing the memory access of GEMM.

Currently, two techniques, software prefetching and data packing,
are widely adopted in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache ahead of use, so memory accesses will not
cause long latencies and floating-point instructions can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrix blocks the kernel operates on will be packed into continous memory buffer
whose footprint is within the cache capacity before the kernel routine is called.
Because GEMM is a computational intensive operation whose
arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
the overhead caused by data packing is negligible when
the input matrices are large~\cite{gotogemm}.

The software prefetching and data packing approach meets a challenge
when working on architectures with non-LRU shared caches.
In traditional high-performance processors, e.g. Intel Xeon series,
both L1 and L2 caches are private to processor cores
and LRU replacement policy is adopted.
Without considering the last level cache (LLC) if presented,
threads running on different cores prefetch their data into
different caches, and private data of different threads
cannot cause conflict cache misses.
This is not the case if caches are shared by several processor cores.
There are chances that a cache line prefetched by one thread
get evicted out before its lifetime ends because another thread prefetches
another cache line into the same cache set.
A simple solution would be shrinking the size of packed matrix
so data used by several threads can simultaneously live in the shared cache.
Due to the nature of set-associative caches,
this solution cannot completely eliminate inter-thread cache conflicts,
especially if the shared cache has a non-LRU replacement policy.
Evaluation results show that inter-thread cache conflicts on non-LRU shared caches
can heavily harm GEMM performance,
even with packed matrices shrinked to fit the cache size.

Following the trend in recent decades, modern architecture design
are introducing more and more cores on a single processor chip.
To reduce cost of on-chip cache memories and coherence networks,
shared cache may become common in future many-core processors.
And non-LRU replacement policies, for example, psudo-random,
may also be used to further reduce the cache design complexity.
As far as we know, some high-performance processors based on ARM technology,
have already adopted this design. (FIXME list the processors)
As a consequence, a better solution to the inter-thread cache conflicts problem
on non-LRU shared caches is meaningful.

In this paper, we present a Shared Cache Partition (SCP) method,
to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
The key idea is partition the share cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanisim of set-associative caches.

The main contributions of this paper are:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with non-LRU shared caches.
\item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture,
  which has 64 cores, private LRU L1 caches and psudo-random shared L2 caches.
  Evaluation results show that GEMM performance get improved consistently
  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$..
\end{itemize}

To the best of our knowledge, SCP is the first proposal targeted at
solving the inter-thread cache conflicts problem on
architectures with non-LRU share caches.

The rest of the paper is organized as follows.
Section~\ref{sec:background} gives some background on GEMM and
the cache conflicts problem in GEMM problem.
Section~\ref{sec:scp} illustrates the SCP method.
Section~\ref{sec:evaluation} presents evaluation results.
Section~\ref{sec:related} reviews the related work.
Finally, Section~\ref{sec:conclusion} concludes.

### 2018-05-23 related.tex
Add all source files to repo. This is the first draft.

\section{Releated Work}\label{sec:related}
The idea that all level-3 BLAS operations can be built upon
a high-performance GEMM implemention was first proposed
in~\cite{gemmbased1,gemmbased2}.
The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
are instantiated based on this insight.
So optimizing GEMM has always been the central task in developing
dense linear algebra software.
The GEMM optimization has two aspects.
One is developing fast computation kernels and
the other is choosing a proper overall strategy
to optimize memory access.

%% Broadly speaking, there are two aspects of this task,
%% developing fast computation kernels and optimizing memory access.
%% The former corresponds the to the kernel component and
%% the latter is the target of overall strategies.

There are several approaches for obtaining optimized GEMM kernels,
%% (1) assembly programming, (2) auto-tuning, and (3) directive-based programming,
yielding different tradeoffs between performance and portability.
In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
the kernel component of GEMM are written by domain experts in assembly.
In the case of auto-tuning,
ATLAS~\cite{atlas} adopts the auto-tuning method to
automatically generate kernels with different parameters
in C and find the best-performing one by running them
on the actual computing system.
POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
directive-based programming approache.
Given a GEMM kernel in C, POET inserts annotations into
the C code to direct source-to-source compiler transformations.
AUGEM uses a template-based method to match
predefined patterns in the C code and transforms the matched
C code sequence into an optimized assembly code sequence.
\textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
abstractions available in the LLVM compiler infrastructure
and proposes a series of domain-specific yet architecture-independent optimizations
to obtain high-performance kernels in a portable way.

The overall strategy of GEMM mainly concerns about
tiling factors, loop orders and parallelization.
The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
are essential to GEMM performance.
ATLAS~\cite{atlas} relies on auto-tuning to determine optimal
values for these factors.
Analytic techniques~\cite{analytic1,analytic2,blisanalytic} are
can also be used instead of auto-tuning.
These analytic methods generally take into consideration
the cache capacity and associativity
but assume a LRU replacement policy.
As far as we know, this paper is the first to
address the problem of cache sharing with non-LRU replacement policies,
and SCP serves as the first proposal to handle this problem.
\cite{gotogemm} gives a detailed discussion of different looping orders.
GEMM are usually parallelized only on layer 3 (the M loop),
but all loops along the M and N dimensions are potentially parallelizable.
BLIS\cite{blispar} allows the users to specify a sophisticated configuration
that more than one loops of layer 1, 3, 4 and 5
can be simultaneously parallelized.
This nested parallelization can better fit complex architecture features
like multi-sockets and hyperthreading.

### 2018-05-23 scp.tex
Add all source files to repo. This is the first draft.

\section{SCP: The Shared Cache Partition Method}\label{sec:scp}

This section demonstrates the SCP method to handle inter-thread cache conflicts.
The basic idea is that inter-thread cache conflicts can be eliminated
if packed matrices used by different threads are fetched to different cache sets.
In Section~\ref{subsec:example}, we demonstrate SCP by an example.
Then formal description is given in Section~\ref{subsec:formal}.

\subsection{Example}\label{subsec:example}

Suppose we are implementing a double precision GEMM (DGEMM) on
the 4-core processor, i.e. $es=8B$.
We now determine the tiling factors used in DGEMM following the procedure described
by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
e.g. $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
It takes careful consideration to determine which one is the best,
and a tuning process may be needed.
After tuning we got the best values $M_r = 4$ and $N_r = 8$.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures, as well as the program code,
it is reasonable to shrink $M_c$ to a slightly smaller value,
e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
Similar to $M_r$ and $N_r$, a tunning process may be needed to
find the best value for $M_c$.
Here we choose $M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) need not to be
concerned and $N_c$ can be given a large value.
Here we choose $N_c = 1024NT = 4096$.
With tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} list the sizes of packed matrices.

\begin{table}
  \centering
  \caption{Sizes of packed matrices in DGEMM ($es = 8B$)}
  \label{tab:msizes}
  \begin{tabular}{cl|cl}
    \toprule
    matrix & size & matrix & size \\
    \midrule
    $A_2$, $A_3$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
    $A_4$ & $es M_r K_c = 8KB$   & $B_3$, $B_4$ & $es N_r K_c = 16KB$ \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure}
  \centering
  \subfigure[Way-partition]{
    \label{fig:partition.conventional}
    \includegraphics[width=.4\textwidth]{figures/wpart}
  }
  \subfigure[Set-partition]{
    \label{fig:partition.segmented}
    \includegraphics[width=.4\textwidth]{figures/spart}
  }
  \caption{Partition styles of the shared L2 cache (n=16, ns=2048)}
  \label{fig:partition}
\end{figure}

Each $A_2$ is packed into a single continous buffer.
Figure.~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
We see that every cache set contains data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure.~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets,
and inter-thread cache conflicts caused by $A_2$ are eliminated completely!

Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
essentially represent two different partition style of the shared cache,
refered as way-partition and set-partition, respectively.
To guarantee the data layout in set-partition,
$A_2$ cannot be stored in a single continous buffer any more.
Instead, $A_2$ is distributed to 12 continous memory segments, each of size $32KB$.
Distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While the set-partition of shared caches avoids inter-thread cache conflicts,
it introduces some complexity to GEMM implementation
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continous ones.

How about $B_2$, the other packed matrix use in GEBP?
Unlike $A_2$, which is thread private data,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collabratively by all threads.
We have two choices, privatize $B_2$ and apply the set-partition method,
or fall back to conventional way-partition for $B_2$.
$B_2$ can be made thread private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but risks inter-thread conflicts caused by the shared $B_2$ matrix.
The two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Formal Description}\label{subsec:formal}
Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierachy,
The SCP method systematically determines the
the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
We assume that the share matrix $B_2$ is packed in conventional way-partition style.
If desired, the $B_2$ can be privatized and handled in the same way as $A_2$.

Before diving into details, we make a few assumptions,
which are standard practice rarely violated,
on the architectures that SCP can deal with:
\begin{itemize}
\item The architecture has 2 or 3 levels of cache. %% i.e. $NL=2$ or $NL=3$
\item All caches are inclusive set-associative caches. %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
\item Caches on the same level are homogeneous. %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
\item A cache's way-capacity is a multiple of the sum of way-capacity of its children,
i.e. $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
\end{itemize}

The memory layout of a packed matrix is described by a memory descriptor,
which is essentially a subspace of the whole address space $\mathcal{A}$.
Each packed matrix $M$ used by thread $t$,
is associated with a memory descriptor $\mathcal{D}_t^M$
specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
A segment is represented by its start and end addresses $S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partition, the memory descriptor contains only one segment.
With set-partition used in previous section,
the memory descriptor contains a sequence of equally strided segments.

%% FIXME use l as the level iterator?
\begin{algorithm}
  \caption{SCP phase 1: allocate memory buffers}
  \label{alg:scp.phase1}
  \begin{algorithmic}[1]
    \REQUIRE $L_l = (c_l,l_l,n_l,nt_l)$ for $l \in [0, NL]$, $NT$,
    $M_r$, $N_r$, $K_c$, $M_c$, $N_c$
    \ENSURE $\mathcal{D}_p$ and $\mathcal{D}_s$
    \STATE $size_p \gets es M_c K_c$ \label{line:size.p}
    \STATE $size_s \gets es N_c K_c / NT$ \label{line:size.s}
    \STATE $align \gets l_1$ \label{line:align.init}
    \FOR {$l=1$ to $NL$} \label{line:align.for}
    \IF {$nt_l / nt_{l-1} > 1$ \AND $L_l$ is non-LRU} \label{line:align.type}
    \STATE $align \gets lcm(align, wc_l / nt_l)$ \label{line:align.update}
    \ENDIF
    \ENDFOR \label{line:align.endfor}
    \STATE $size_p \gets \lceil size_p / align \rceil \cdot align$ \label{line:align}
    %% \STATE $size \gets NT \cdot size_p + size_s$
    \STATE $addr_p \gets allocate(\mathcal{A}, size_p \cdot NT)$ \label{line:alloc.begin}
    \STATE $addr_s \gets allocate(\mathcal{A}, size_s \cdot NT)$ \label{line:alloc.end}
    \STATE $\mathcal{D}_p \gets \lbrace [addr_p, addr_p + size_p \cdot NT) \rbrace$ \label{line:d.begin}
    \STATE $\mathcal{D}_s \gets \lbrace [addr_s, addr_s + size_s \cdot NT) \rbrace$ \label{line:d.end}
  \end{algorithmic}
\end{algorithm}

SCP runs in three phases.
The first phase allocates memory buffers for all $A_2$ and $B_2$
instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
for storing private and shared matrices respectively.
The second phase computes memory descriptors for $A_2$ instances
by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
$\mathcal{D}_t^{A_2}$ , $t \in [0, NT)$.
Buffers of different threads are disjoint,
i.e. $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
The third phase computes memory descriptors for $B_2$.
Because is $B_2$ is shared among all threads,
$\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
but the part (size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
The working flow of these three phases are listed
in Algorithm~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.

In Algorithm~\ref{alg:scp.phase1},
the per-thread size for memory buffer $\mathcal{D}_p$ and
$\mathcal{D}_s$ are first computed
(line~\ref{line:size.p}--\ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
and $size_p$ is enlarged to a multile of $align$ (line~\ref{line:align}).
$align$ is initialized with the L1 cache line size $l_1$.
Then on each level $l$ (Line~\ref{line:align.for})
with shared non-LRU caches (Line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight in the $wc_l/nt_l$ value is that
each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
in the set-partition of shared caches.
At last, memory are allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin}--\ref{line:d.end}).

\begin{algorithm}
  %% a trick to temporally define customized command
  \renewcommand{\algorithmicprint}{\textbf{call}}
  \renewcommand{\algorithmicwhile}{\textbf{procedure}}
  \renewcommand{\algorithmicendwhile}{\textbf{end procedure}}
  \caption{SCP phase 2: compute descriptors for $A_2$}
  \label{alg:scp.phase2}
  \begin{algorithmic}[1]
    \REQUIRE $L_l = (c_l,l_l,n_l,nt_l)$ for $l \in [0, NL]$, $\mathcal{D}_p$, $NT$
    \ENSURE $\mathcal{D}_t^{A_2}$ for $t \in [0, NT)$
    %% \ENSURE $\mathcal{D}_l^i$ for $l \in [0, NL]$ and $i \in [0, NT/nt_l]$
    \STATE $\mathcal{D}_{NL+1}^0 \gets \mathcal{D}_p$ \label{line:memory.d}
    \STATE $nt_{NL+1} \gets NT$ \label{line:memory.nt}
    \PRINT $subspace(NL+1, 0, \mathcal{D}_{NL+1}^0)$ \label{line:subspace.root}
    \STATE                      % for a blank line
    \WHILE {$subspace\ (l, idx, \mathcal{D}_l^{idx})\ $} \label{line:subspace.begin}
    \IF {$l = 0$} \label{line:shortcut.begin}
    \STATE $\mathcal{D}_{idx}^{A_2} \gets \mathcal{D}_l^{idx}$
    \RETURN
    \ENDIF \label{line:shortcut.end}
    \STATE $nchildren \gets nt_l / nt_{l-1}$ \label{line:nchildren}
    \FOR {$i = 0$ to $nchildren-1$} \label{line:for.begin}
    \STATE $cidx \gets idx \cdot nchildren + i$ \label{line:cidx}
    \IF {$l = NL+1$ \OR $L_l$ is LRU} \label{line:if}
    \STATE $lb \gets min(\mathcal{D}_l^{idx})$
    \STATE $ub \gets max(\mathcal{D}_l^{idx})+1$
    \STATE $len \gets (ub - lb) / nchildren$
    \STATE $\mathcal{D}_{temp} \gets [lb + i \cdot len, lb + (i+1) \cdot len)$
    \ELSE \label{line:else}
    \STATE $len \gets ns_{l} / nchildren$
    \STATE $\mathcal{S}_i \gets [i \cdot len, (i+1) \cdot len)$
    \STATE $\mathcal{D}_{temp} \gets \varphi_l^{-1}(\mathcal{S}_i)$
    \ENDIF \label{line:endif}
    \STATE $\mathcal{D}_{l-1}^{cidx} \gets \mathcal{D}_{l}^{idx} \bigcap \mathcal{D}_{temp}$
    \label{line:childspace}
    \PRINT $subspace(l-1, cidx, \mathcal{D}_{l-1}^{cidx})$ \label{line:recursive}
    \ENDFOR \label{line:for.end}
    \ENDWHILE \label{line:subspace.end}
  \end{algorithmic}
\end{algorithm}

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
The main part is a recursive procedure $subspace$
(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierachy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
$subspace$ has three input parameters, $l$ and $idx$ to
identify the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e. a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
If this is not the case, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(line~\ref{line:for.begin}--\ref{line:for.end}).
For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branches (line~\ref{line:if}--\ref{line:endif})
show two distinct ways to compute $\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partition style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partition on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierachy in a depth-first-search order.

The computation of $\mathcal{D}_t^{B_2}$ is quite simple.
Algorithm~\ref{alg:scp.phase3} divides $\mathcal{D}_s$
into $NT$ partitions, one for each thread.
First, the range of $\mathcal{D}_s$ is obtained
(line~\ref{line:lb}--\ref{line:ub}) and
the capacity of $\mathcal{D}_s$ is equally
divided among $NT$ threads (line~\ref{line:len}).
Then $\mathcal{D}_s$ is partitioned in a way-partition style
and each thread $t$ get buffer space for its part in $B_2$
(line~\ref{line:thread.for}--\ref{line:thread.forend}).

\begin{algorithm}
  \caption{SCP phase 3: compute descriptors for $B_2$}
  \label{alg:scp.phase3}
  \begin{algorithmic}[1]
    \REQUIRE $L_l = (c_l,l_l,n_l,nt_l)$ for $l \in [0, NL]$, $\mathcal{D}_s$
    \ENSURE $\mathcal{D}_t^{B_2}$ for $t \in [0, NT)$
    \STATE $lb \gets min(\mathcal{D}_s)$ \label{line:lb}
    \STATE $ub \gets max(\mathcal{D}_s)+1$ \label{line:ub}
    \STATE $len \gets (ub - lb) / nchildren$ \label{line:len}
    \FOR {$t=0$ to $NT-1$} \label{line:thread.for}
    \STATE $\mathcal{D}_t^{B_2} \gets \lbrace [lb + t \cdot len, lb + (t+1) \cdot len) \rbrace$
    \ENDFOR \label{line:thread.forend}
  \end{algorithmic}
\end{algorithm}

### 2018-05-23 figures/gemm.py

@@ -1,0 +1,593 @@
+#!/usr/bin/python3
+
+import sys
+import numpy as np
+import matplotlib as mpl
+import matplotlib.patches as mp
+import matplotlib.transforms as mtr
+import matplotlib.text as mtext
+import matplotlib.lines as mline
+import matplotlib.pyplot as plt
+
+mpl.rcParams['pdf.fonttype'] = 42
+mpl.rcParams['ps.fonttype'] = 42
+
+def grid(row, col, width=1, height=1, width_ratios=None, height_ratios=None, 
+         left=0, right=0, top=0, bottom=0):
+  hratios = None
+  wratios = None
+  if not height_ratios:
+    height_ratios = np.ones(row)
+    hratios = np.arange(0, row+1)
+  else:
+    hratios = np.empty(row+1)
+    hratios[1:] = np.cumsum(height_ratios)
+  if not width_ratios:
+    width_ratios = np.ones(col)
+    wratios = np.arange(0, col+1)
+  else:
+    wratios = np.empty(col+1)
+    wratios[1:] = np.cumsum(width_ratios)
+  runit = (height-top-bottom)/hratios[-1]
+  cunit = (width-left-right)/wratios[-1]
+  hratios = hratios[:-1]
+  wratios = wratios[:-1]
+  rbases = runit * hratios
+  cbases = cunit * wratios
+  xys = np.array([(left+c, height-top-r) for r in rbases for c in cbases])
+  xys = xys.reshape((row, col, 2))
+  whs = np.array([(w*cunit, -h*runit) for h in height_ratios for w in width_ratios])
+  whs = whs.reshape((row, col, 2))
+  bbs = np.empty((row, col, 4))
+  bbs[:,:,:2] = xys
+  bbs[:,:,2:] = whs
+  return bbs
+
+# def offsettrans(x=0, y=0):
+#   trans = mtr.Affine2D.from_values(1,0,0,1,x,y)
+
+def box(bounds):
+  return mtr.Bbox.from_bounds(*bounds)
+def canonicalbox(box):
+  return mtr.Bbox.from_bounds(box.xmin, box.ymin, abs(box.width), abs(box.height))
+def padbox(box, ratio=-.1):
[…]

### 2018-05-23 figures/ipy.py

@@ -1,0 +1,146 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import os.path as path
+from timing import loadb, statistics
+
+ntcs = [4, 2, 1]
+ncs = [16, 8, 4, 2, 1]
+mems = ['0', '0-1', '0-3', '0-7']
+lb = 3072
+ub = 9216
+x = np.arange(lb, ub+1, 128)
+xindices = (x >> 7) - 1
+# xticks = x[::2]
+# xticklabels = np.array([str(t>>10)+'K' for t in xticks])
+
+def summarize(prog):
+  shape = (len(ntcs), len(ncs), len(mems), 9216/128)
+  results = np.zeros(shape)
+
+  for i0, ntc in enumerate(ntcs):
+    for i1, nc in enumerate(ncs):
+      nt = ntc * nc
+      if nt < 4:
+        continue
+      lb = 3072
+      ub = nt * 1024
+      if nt <= 32 and ub > 6144:
+        ub = 6144
+      if ub > 9216:
+        ub = 9216
+      for i2, mem in enumerate(mems):
+        step = 128
+        size = lb
+        while size <= ub:
+          if size >= 6144:
+            step = 512
+          else:
+            step = 128
+          filename = prog + '_' + str(ntc) + '_' + str(nc) + '_' + mem + '_' + str(size) + '.tout'
+          if not path.exists(filename):
+            print('MISSING', filename)
+            continue
+          data = loadb(filename)
+          gflops = statistics(data, peak=8.8, verbose=False)
+          print(filename, gflops)
+          results[i0, i1, i2, (size>>7)-1] = gflops
+          size += step
+  np.save(prog + '.npy', results)
+  return results
+
+def select(data, ntc, nc, mem):
+  assert ntc * nc >= 4
+
+  i0 = ntcs.index(ntc)
+  i1 = ncs.index(nc)
+  i2 = mems.index(mem)
+  y = data[i0, i1, i2, xindices]
+  indices = y != 0.0
+  return x[indices], y[indices]
+
+def plot(con, scp):
+  nts = { x * y for x in ntcs for y in ncs if x * y >= 4 }
+  nts = list(nts)
+  nts.sort()
+  figprops = dict(figsize=(10*len(nts), 10*len(mems)))
+  fig, axs = plt.subplots(len(mems), len(nts), **figprops)
+  
[…]

### 2018-05-23 figures/papi-trick.py

@@ -1,0 +1,250 @@
+#! python3
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from math import ceil
+import os.path as path
+
+def parse(filename):
+  results = None
+  event_names = []
+  with open(filename, 'rb') as f:
+    num_events, = np.fromfile(f, 'int32', 1, '')
+    for i in range(num_events):
+      length, = np.fromfile(f, 'int32', 1, '')
+      name = np.fromfile(f, 'byte', length, '')
+      event_names.append(name.tostring())
+    num_threads, = np.fromfile(f, 'int32', 1, '')
+    results = np.zeros((num_threads, num_events), 'int64')
+    for i in range(num_threads):
+      results[i] = np.fromfile(f, 'int64', num_events, '')
+  return event_names, results
+
+def summarize(prog):
+  ntcs = [4, 2, 1]
+  ncs = [4, 2, 1]
+  lb = 3072
+  ub = 6144
+  step = 128
+  results = None
+  for i0, ntc in enumerate(ntcs):
+    for i1, nc in enumerate(ncs):
+      nt = ntc * nc
+      if nt < 4:
+        continue
+      size = lb
+      while size <= min(ub, nt * 1024):
+        filename = prog + '_' + str(ntc) + '_' + str(nc) + '_' + str(size) + '.pout'
+        if not path.exists(filename):
+          print('MISSING', filename)
+          size += step
+          continue
+        event_names, data = parse(filename)
+        means = data.mean(axis=0)
+        if id(results) == id(None):
+          shape = (len(ntcs), len(ncs), ub//step, len(event_names))
+          results = np.zeros(shape)
+        results[i0, i1, size//step-1] = means
+        size += step
+  np.save(prog + '-papi.npy', results)
+
+def select(data, ntc, nc, event=None, sizes=None):
+  events = ['L1M',
+            'L1H',
+            'L2M',
+            'L2H',
[…]

### 2018-05-23 figures/papi.py

@@ -1,0 +1,179 @@
+#! python3
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from math import ceil
+import os.path as path
+
+def parse(filename):
+  results = None
+  event_names = []
+  with open(filename, 'rb') as f:
+    num_events, = np.fromfile(f, 'int32', 1, '')
+    for i in range(num_events):
+      length, = np.fromfile(f, 'int32', 1, '')
+      name = np.fromfile(f, 'byte', length, '')
+      event_names.append(name.tostring())
+    num_threads, = np.fromfile(f, 'int32', 1, '')
+    results = np.zeros((num_threads, num_events), 'int64')
+    for i in range(num_threads):
+      results[i] = np.fromfile(f, 'int64', num_events, '')
+  return event_names, results
+
+def summarize(prog):
+  ntcs = [4, 2, 1]
+  ncs = [4, 2, 1]
+  lb = 3072
+  ub = 6144
+  step = 128
+  results = None
+  for i0, ntc in enumerate(ntcs):
+    for i1, nc in enumerate(ncs):
+      nt = ntc * nc
+      if nt < 4:
+        continue
+      size = lb
+      while size <= min(ub, nt * 1024):
+        filename = prog + '_' + str(ntc) + '_' + str(nc) + '_' + str(size) + '.pout'
+        if not path.exists(filename):
+          print('MISSING', filename)
+          size += step
+          continue
+        event_names, data = parse(filename)
+        means = data.mean(axis=0)
+        if id(results) == id(None):
+          shape = (len(ntcs), len(ncs), ub//step, len(event_names))
+          results = np.zeros(shape)
+        results[i0, i1, size//step-1] = means
+        size += step
+  np.save(prog + '-papi.npy', results)
+
+def select(data, ntc, nc, event=None, sizes=None):
+  events = ['L1M',
+            'L1H',
+            'L2M',
+            'L2H',
+            'LIS']
+  ntcs = [4, 2, 1]
+  ncs = [4, 2, 1]
+  lb = 3072
[…]

### 2018-05-23 figures/partitions.py

@@ -1,0 +1,93 @@
+#! python3
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from math import ceil
+import os.path as path
+
+ntc=4
+nways=16
+wc=128
+seg=wc/ntc
+
+height=0.8
+
+colors = ['lightskyblue', 'tan', 'lightgrey', 'pink', 'white']
+hatches = ['/', '..', '\\', 'x', None]
+
+usehatch = True
+
+def wpart():
+  fig, ax = plt.subplots()
+  wpt = 3
+  for t in range(ntc):
+    ways = np.arange(wpt*t, wpt*(t+1))
+    if usehatch:
+      ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
+              color='None', hatch=hatches[t], label='$T'+str(t)+'$')
+    else:
+      ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=0,
+              color=colors[t], label='$T'+str(t)+'$')
+  t = ntc
+  ways = np.arange(wpt*ntc, nways)
+  if usehatch:
+    ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
+            color='None', hatch=hatches[ntc])
+  else:
+    ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3, color=colors[ntc])
+
+  yticks = np.arange(0,nways)
+  ax.set_yticks(yticks+height/2)
+  ax.set_yticklabels(yticks, fontsize='x-large')
+  ax.set_ylabel('Cache ways', fontsize='xx-large')
+  ax.set_xlim(0, wc+seg+4)
+  xticks = np.linspace(0, wc, ntc+1).astype(int)
+  ax.set_xticks(xticks+.5)
+  ax.set_xticklabels([x*16 for x in xticks], fontsize='x-large')
+  # ax.set_xticklabels([str(x)+'K' for x in xticks], fontsize='x-large')
+  ax.set_xlabel('Cache sets', fontsize='xx-large')
+  ax.legend(fontsize='x-large', loc='upper right')
+  ax.tick_params(length=0)
+  ax.set_frame_on(False)
+  ax.set_aspect(.8 * wc/nways)
+  fig.savefig('wpart.pdf')
+
+def spart():
+  fig, ax = plt.subplots()
[…]

### 2018-05-23 figures/privb.py

@@ -1,0 +1,241 @@
+#! python3
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from math import ceil
+import os.path as path
+from timing import loadb, statistics, performance
+
+def papi():
+  lb = 3072
+  ub = 6144
+  step = 128
+  sizes = np.arange(lb, ub+1, step)
+  scp = np.load('v3seg-4x4-1.npy')
+  pri = np.load('v3pb-4x4-1.npy')
+
+  l1ms, l1hs, l2ms, l2hs = scp
+  l1mp, l1hp, l2mp, l2hp = pri
+
+  fig, ax = plt.subplots()
+  ys = l1ms/(l1ms+l1hs)*100
+  yp = l1mp/(l1mp+l1hp)*100
+  l1, = ax.plot(sizes, ys, markersize=10, marker='o', label='$L1$ $SCP$')
+  l2, = ax.plot(sizes, yp, markersize=10, marker='x', label='$L1$ $SCP$-$P$')
+  ys = l2ms/(l2ms+l2hs)*100
+  yp = l2mp/(l2mp+l2hp)*100
+  l3, = ax.plot(sizes, ys, markersize=10, marker='s', label='$L2$ $SCP$')
+  l4, = ax.plot(sizes, yp, markersize=10, marker='+', label='$L2$ $SCP$-$P$')
+
+  ax.set_position([0.1, 0.2, 0.8, 0.75])
+  xticks = sizes
+  if np.max(sizes) > 4096:
+    xticks = xticks[::2]
+  ax.set_xticks(xticks)
+  ax.set_xticklabels(xticks, rotation=45, fontsize='x-large')
+  ax.set_xlabel('Matrix Size ($M=N=K$)', fontsize='xx-large')
+  ax.set_ylim(3, 10)
+  yticks=np.arange(3, 11, 1)
+  yticklabels=yticks
+  ax.set_yticks(yticks)
+  ax.set_yticklabels(yticklabels, fontsize='x-large')
+  ax.set_ylabel('Cache Miss Rate ($\%$)', fontsize='xx-large')
+  ax.legend(fontsize='xx-large', loc='lower right', ncol=2)
+  fig.savefig('privb-papi.pdf')
+
+def breakdown():
+  scp = np.load('breakdown-scp.npy')
+  pri = np.load('breakdown-pri.npy')
+  lb = 3072
+  ub = 6144
[…]

### 2018-05-23 figures/sec1.py

@@ -1,0 +1,76 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import os.path as path
+from timing import loadb, statistics
+
+lb = 3072
+ub = 9216
+x = np.arange(lb, ub+1, 128)
+xindices = (x >> 7) - 1
+
+def select(data, ntc, nc, mem):
+  assert ntc * nc >= 4
+  ntcs = [4, 2, 1]
+  ncs = [16, 8, 4, 2, 1]
+  mems = ['0', '0-1', '0-3', '0-7']
+
+  i0 = ntcs.index(ntc)
+  i1 = ncs.index(nc)
+  i2 = mems.index(mem)
+  y = data[i0, i1, i2, xindices]
+  indices = y != 0.0
+  return x[indices], y[indices]
+
+if __name__ == '__main__':
+  con = np.load('v3.npy')
+  ntcs = [4, 2, 1]
+  ncs = [16, 8, 4, 2, 1]
+  mem = '0-7'
+  nts = [4, 8, 16, 32]
+  markers = ['x', 'o', 's', '+']
+  figs = [plt.subplots() for nt in nts]
+  for i0, ntc in enumerate(ntcs):
+    for i1, nc in enumerate(ncs):
+      nt = ntc * nc
+      if nt not in nts:
+        continue
+      idx = nts.index(nt)
+      fig, ax = figs[idx]
+      marker = markers[i0]
+      xs, ys = select(con, ntc, nc, mem)
+      ys *= 100
+      label = '$NT_C='+str(ntc)+'$' + ' ' + '$NC='+str(nc)+'$'
+      l, = ax.plot(xs, ys, marker=marker, label=label, markersize=10)
+  # handle nt=64 specially
+  i0, ntc = 2, 4
+  i1, nc = 0, 16
+  idx = nts.index(32)
+  fig, ax = figs[idx]
+  marker = markers[i0]
+  xs, ys = select(con, ntc, nc, mem)
+  out = (xs % 256 == 0) & (xs <= 6144)
+  xs = xs[out]
+  ys = ys[out]
+  ys *= 100
+  label = '$NT_C='+str(ntc)+'$' + ' ' + '$NC='+str(nc)+'$'
+  l, = ax.plot(xs, ys, marker=marker, label=label, markersize=10)
+  
+  for i, (fig, ax) in enumerate(figs):
+    ax.set_position([0.1, 0.2, 0.8, 0.75])
+    left, right = ax.get_xlim()
[…]

### 2018-05-23 figures/sec2.py

@@ -1,0 +1,77 @@
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import os.path as path
+from timing import loadb, statistics
+
+lb = 3072
+ub = 9216
+x = np.arange(lb, ub+1, 128)
+xindices = (x >> 7) - 1
+
+def select(data, ntc, nc, mem):
+  assert ntc * nc >= 4
+  ntcs = [4, 2, 1]
+  ncs = [16, 8, 4, 2, 1]
+  mems = ['0', '0-1', '0-3', '0-7']
+
+  i0 = ntcs.index(ntc)
+  i1 = ncs.index(nc)
+  i2 = mems.index(mem)
+  y = data[i0, i1, i2, xindices]
+  indices = y != 0.0
+  return x[indices], y[indices]
+
+if __name__ == '__main__':
+  con = np.load('v3.npy')
+  scp = np.load('v3seg.npy')
+  ntcs = [4, 2]
+  ncs = [16, 8, 4, 2, 1]
+  mems = ['0-7']
+  nts = [4, 8, 16, 32, 64]
+  markers = ['o', 'x', '*', 's']
+  figs = [plt.subplots() for nt in nts]
+  for i0, ntc in enumerate(ntcs):
+    for i1, nc in enumerate(ncs):
+      nt = ntc * nc
+      if nt not in nts:
+        continue
+      idx = nts.index(nt)
+      fig, ax = figs[idx]
+      for mem in mems:
+        xc, yc = select(con, ntc, nc, mem)
+        xs, ys = select(scp, ntc, nc, mem)
+        if nt == 64:
+          out = (xs % 256 == 0) & (xs <= 6144)
+          xc = xc[out]
+          yc = yc[out]
+          xs = xs[out]
+          ys = ys[out]
+        label = '$NT_C='+str(ntc)+'$' + ' ' + '$NC='+str(nc)+'$'
+        yc *= 100
+        ys *= 100
+        if ntc == 2:
+          ax.plot(xc, yc, markersize=10, marker='x', label=label+ ' $Base$')
+          ax.plot(xs, ys, markersize=10, marker='o', label=label+ ' $SCP$')
+        else:
+          ax.plot(xc, yc, markersize=10, marker='+', label=label+ ' $Base$')
+          ax.plot(xs, ys, markersize=10, marker='s', label=label+ ' $SCP$')
+  for i, (fig, ax) in enumerate(figs):
[…]

### 2018-05-23 figures/timing.py

@@ -1,0 +1,193 @@
+#! python3
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from math import ceil
+import os.path as path
+
+def styles(num_threads, colors, markers):
+  clen = len(colors)
+  mlen = len(markers)
+  if (num_threads <= min(clen, mlen)):
+    return list(zip(colors, markers))
+  return [(c, m) for m in markers for c in colors]
+
+class thread_data():
+  def __init__(self):
+    self.kernels = None
+    self.pa = None
+    self.pb = None
+    self.s0 = None
+    self.s1 = None
+    self.s2 = None
+    self.init = None
+    self.performance = None
+    self.owners = None
+
+def loadb(filename):
+# typedef struct {
+#   int32_t kind;
+#   int32_t owner;
+#   double start;
+#   double stop;
+#   double amount;
+# } time_point_t;
+  dt = { 'names': ['kind', 'owner', 'start', 'stop', 'amounts'],
+         'formats': ['int32', 'int32', 'float64', 'float64', 'float64'] }
+  tds = []
+  with open(filename, 'rb') as f:
+    num_threads = np.fromfile(f, 'int32', 1, '')
+    num_points = np.fromfile(f, 'int32', num_threads, '')
+    for i in range(num_threads):
+      rtd = np.fromfile(f, dt, num_points[i], '')
+      td = thread_data()
+      kindices = rtd['kind'] == 0
+      aindices = rtd['kind'] == 1
+      bindices = rtd['kind'] == 2
+      sindices0 = rtd['kind'] == 3
+      sindices1 = rtd['kind'] == 4
+      sindices2 = rtd['kind'] == 5
+      iindices = rtd['kind'] == 6
+      td.kernels = rtd[kindices]
+      td.pa = rtd[aindices]
+      td.pb = rtd[bindices]
+      td.s0 = rtd[sindices0]
+      td.s1 = rtd[sindices1]
+      td.s2 = rtd[sindices2]
+      td.init = rtd[iindices]
+      tds.append(td)
+  return tds
+
+def statistics(data, peak, verbose=False):
+  num_threads = len(data)
+  jobs = np.zeros(num_threads, int)
+  whole = 0.0
+  time = 0.0
+  for i, td in enumerate(data):
[…]

### 2018-05-23 figures/workload.py

@@ -1,0 +1,216 @@
+#! python3
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from math import ceil
+import os.path as path
+
+def boxlabel(ax, box, label, pos, distance=.2, fontdict=dict(), inside=False, arrows=True):
+  assert pos in ['n','s','e','w']
+  lbcolor = 'black'
+  if 'color' in fontdict:
+    lbcolor = fontdict['color']
+  factor = 1
+  if inside:
+    factor = -1
+  x, y = box.get_xy()
+  w = box.get_width()
+  h = box.get_height()
+  if pos == 'n':
+    tx = x + w/2
+    ty = y + h + h*distance*factor
+    a0x0, a0y0 = tx-distance*w, ty
+    a0x1, a0y1 = x, ty
+    a1x0, a1y0 = tx+distance*w, ty
+    a1x1, a1y1 = x+w, ty
+  elif pos == 's':
+    tx = x + w/2
+    ty = y - h*distance*factor
+    a0x0, a0y0 = tx-distance*w, ty
+    a0x1, a0y1 = x, ty
+    a1x0, a1y0 = tx+distance*w, ty
+    a1x1, a1y1 = x+w, ty
+  elif pos == 'e':
+    ty = y+h/2
+    tx = x + w + w*distance*factor
+    a0x0, a0y0 = tx, ty+distance*h
+    a0x1, a0y1 = tx, y+h
+    a1x0, a1y0 = tx, ty-distance*h
+    a1x1, a1y1 = tx, y
+  else:
+    ty = y+h/2
+    tx = x - w*distance*factor
+    a0x0, a0y0 = tx, ty+distance*h
+    a0x1, a0y1 = tx, y+h
+    a1x0, a1y0 = tx, ty-distance*h
+    a1x1, a1y1 = tx, y
+  t = ax.text(tx, ty, label, fontdict=fontdict)
+  if not arrows:
+    return
+  if pos == 'n':
+    a0x0, a0y0 = tx-distance*w, ty
+    a0x1, a0y1 = x+w/10, ty
+    a1x0, a1y0 = tx+distance*w, ty
+    a1x1, a1y1 = x+w-w/10, ty
+    bar0x = [x, x]
+    bar0y = [ty-h*distance/2, ty+h*distance/2]
+    bar1x = [x+w, x+w]
+    bar1y = bar0y
+  elif pos == 's':
[…]

### 2018-05-25 abstract.tex
Fix syntax errors according to the first revision by Prof. Xue

@@
   GEneral Matrix Multiply (GEMM) is the most fundamental
-  computational kernel routine in BLAS library.
-  To achieve best performance, in-memory data must be
-  prefetched into fast on-chip caches ahead of use.
+  computational kernel routine in the BLAS library.
+  To achieve high performance, in-memory data must be
+  prefetched into fast on-chip caches before they are used.
   Two techniques, software prefetching and data packing,
@@
   These techniques are sufficient for lastly-recent-used (LRU) caches,
-  which is often the case in traditional high-performance
-  processors used in high-end servers and supercomputers.
-  But in recent years the market is meeting a diversity
-  in processor design, and some high-performance processors
-  are equipped with shared caches with non-LRU replacement policies.
-  This rises a challenge to GEMM development in multi-threaded context.
+  which are often used in high-performance
+  processors deployed in high-end servers and supercomputers.
+  But in recent years the market has become increasingly more diverse
+  in processor design, resulting in high-performance processors
+  equipped with shared caches with non-LRU replacement policies.
+  This poses a challenge to GEMM development in a multi-threaded context.
   As several threads try to load data into the shared cache simultaneously,
-  inter-thread cache conflicts would increase significantly.
-  We present a Shared Cache Partition (SCP) method to
+  inter-thread cache conflicts will increase significantly.
+  We present a Shared Cache Partitioning (SCP) method to
   eliminate inter-thread cache conflicts in GEMM routines,
@@
   The results show that conflict misses on both L1 and L2
-  caches are effectively reduced and GEMM performance get improved
-  by $2.75\%$--$6.91\%$.
-  %% FIXME BLIS? FT1500A?
+  caches are effectively reduced, resulting in an improvement
+  of the GEMM performance by $2.75\%$--$6.91\%$.
 \end{abstract}

### 2018-05-25 background.tex
Fix syntax errors according to the first revision by Prof. Xue

@@
 To understand how shared cache can cause inter-thread cache conflicts
-and do harm to GEMM performance, we review some details
+and hurts GEMM performance, we review some details
 about the working mechanisim of set-associative caches.
@@
 with each set containing $n$ cache lines.
-Generally, $ns$ is a power-of-two value.
+Generally, $ns$ is a power-of-two.
 An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
@@
 Each core has 32 128-bit (16B) vector registers,
-each capable of store 2 double precision floating-point numbers.
-To make a consistency, the processor cores on level 0
+each capable of storing 2 double precision floating-point numbers.
+For convenience, the processor cores on level 0
 is represented by its registers, which can be viewed as a special
@@
 In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
-so the multiply-accumulate operation at layer 7 need not to scale $C$ again.
+so the multiply-accumulate operation at layer 7 does not have to scale $C$ again.
 Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
@@
 the constraints on these tiling factors.
-$es$ means the size of matrix element, e.g. 8B for double precision floating-point numbers,
-and $c_l$ means the capacity of registers or caches on level $l$.
+$es$ denotes the size of matrix element, e.g. 8B for double precision floating-point numbers,
+and $c_l$ denotes the capacity of registers or caches on level $l$.
 First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
@@
 Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
-$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in next iteration
+$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
 at layer 5 fit into the L1 cache.
 Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
-$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in next iteration
+$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
 at layer 4 fit into the L2 cache.
@@
 $B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
-Larger matrices on higher layers 1 -- 3 live in main memory.
+Larger matrices on higher layers 1 -- 3 always reside in main memory.
 
@@
 and the overall strategy is defined by the three outter loops at layers 1 -- 3.
-Developer can control how to partition the workload by choosing
+Developers can control how to partition the workload by choosing
 tile factors $M_c$, $N_c$ and $K_c$,
@@
 serve to avoid capacity misses,
-conflict misses can not be eliminated completely.
+conflict misses cannot be eliminated completely.
 %% We use the term ``cache conflict'' to represent conflict miss events.
 
-Recall Figure.~\ref{fig:gemm}, the packed matrix $A_2$
+In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
 will be read $N_c/N_r$ times within a single GEBP execution,
 and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
-The packed matrix $B_4$ is similar that it is expected to
+The packed matrix $B_4$ is similar in that it is expected to
 reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
@@
 As GEBP accesses the matrix data much more frequently than other
-data structures, this repeated prefething stratecy
-can remedy the penalty caused by intra-thread conflicts.
+data structures, this repeated prefething strategy
+can reduce the penalty caused by intra-thread conflicts.
 
@@
 can occur frequently on non-LRU shared caches,
-thus making the technique dealing with intra-thread cache conflicts helpless.
-The situation goes worse with more threads sharing the same cache.
+thus making the technique dealing with intra-thread cache conflicts ineffective.
+The situation worsens with more threads sharing the same cache.

### 2018-05-25 evaluation.tex
Fix syntax errors according to the first revision by Prof. Xue

@@
 the contrived 4-core processor in Figure.~\ref{fig:hierachy}
-except that all 16 clusters connect to the main memory
-with their L2 caches coherented by hardare.
-See Table.~\ref{tab:cluster} for architectural parameters of the cores and caches.
+except that L2 caches of all 16 clusters connect to the main memory
+with hardware coherence.
+Table.~\ref{tab:cluster} lists architectural parameters of the cores and caches.
 We restrict our evaluation to DGEMM,
@@
 TOP500 list of world's most powerful supercomputers (FIXME citation),
-relys on the DGEMM variant.
+relies on the DGEMM variant.
 
@@
 are $NC \in [1, 16]$ and $NT_C \in [1, 4]$, respectively.
-Generally, DGEMM performance are presented in $flops$.
+Generally, DGEMM performance is presented in $flops$.
 %% floating-point operations per second ($flops$).
@@
 
-We first show how cache sharing can harm GEMM performance
+We first show how cache sharing can hurt GEMM performance
 in Section~\ref{subsec:drawback}.
-Then Section~\ref{subsec:benefit} demonstrates how SCP can remedy
+Then Section~\ref{subsec:benefit} demonstrates how SCP can reduce
 the penalty caused by cache sharing.
@@
   \centering
-  \caption{Average thread efficiency win of SCP}
+  \caption{Average thread efficiency improvement of SCP}
   \label{tab:win}
@@
 consistently over all matrix sizes.
-Table.~\ref{tab:win} summarizes the performance win of SCP over Base.
+Table.~\ref{tab:win} summarizes the performance improvement of SCP over Base.
 %% Table.~\ref{tab:benefit} summarizes the average value
@@
 The shared $B_2$ matrix is still packed
-in conventional way-partition style.
+in conventional way-partitioning style.
 As mentioned in Section~\ref{subsec:example},
-set-partition can also be applied to $B_2$ after $B_2$ is privatized.
+set-partitioning can also be applied to $B_2$ after $B_2$ is privatized.
 In this section, we will discuss and evaluate this privatization alternative.

### 2018-05-25 intro.tex
Fix syntax errors according to the first revision by Prof. Xue

@@
 such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
-The HPC community have also contributed several high-quality
+The HPC community has also contributed several high-quality
 BLAS implementations, e.g. ATLAS~\cite{atlas},
@@
 
-Among three levels, level-3 provides the most opportunities
+Among the three levels of BLAS routines, level-3 provides the most opportunities
 for optimization because it has the highest computational complexity ($O(n^3)$).
-Among all level-3 operations, GEneral Matrix Multiply (GEMM) is
+Among all the level-3 operations, GEneral Matrix Multiply (GEMM) is
 of the most interest because all other level-3 operations can be defined
 in terms of GEMM and some level-1 and level-2 operations.
-Much effort have been spent on optimizing GEMM for different architectures,
-both in general algorithm and in architecture specific ways.
+Much effort has been spent on optimizing GEMM for different architectures.
+%% both in general algorithm and in architecture specific ways.
 (FIXME citations)
@@
 in the loop nest's iteration space.
-In multi-threaded context, different tasks (loop tiles) are assigned to
+In a multi-threaded context, different tasks (loop tiles) are assigned to
 different threads to parallelize the whole operation.
@@
 On the other hand, the overall strategy is responsible for optimizing
-memory accesses to protect the kernel from data hunger.
+memory accesses to protect the kernel from being idling waiting for data.
 In this paper, we focus on optimizing the memory access of GEMM.
@@
 Software prefetch instructions are utilized
-to load data into cache ahead of use, so memory accesses will not
+to load data into cache before they are used, so memory accesses will not
 cause long latencies and floating-point instructions can execute at peak throughput.
@@
 
-The software prefetching and data packing approach meets a challenge
+The software prefetching and data packing approach meet a challenge
 when working on architectures with non-LRU shared caches.
@@
 and LRU replacement policy is adopted.
-Without considering the last level cache (LLC) if presented,
+Without considering the last level cache (LLC),
 threads running on different cores prefetch their data into
@@
 A simple solution would be shrinking the size of packed matrix
-so data used by several threads can simultaneously live in the shared cache.
+so data used by several threads can simultaneously reside in the shared cache.
 Due to the nature of set-associative caches,
@@
 can heavily harm GEMM performance,
-even with packed matrices shrinked to fit the cache size.
+even with packed matrices shrinked to fit int the cache.
 
 Following the trend in recent decades, modern architecture design
-are introducing more and more cores on a single processor chip.
+is introducing more and more cores on a single processor chip.
 To reduce cost of on-chip cache memories and coherence networks,
@@
 may also be used to further reduce the cache design complexity.
-As far as we know, some high-performance processors based on ARM technology,
+Indeed, some high-performance processors based on ARM technology,
 have already adopted this design. (FIXME list the processors)
@@
 
-In this paper, we present a Shared Cache Partition (SCP) method,
+In this paper, we present a Shared Cache Partitioning (SCP) method,
 to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
@@
   which has 64 cores, private LRU L1 caches and psudo-random shared L2 caches.
-  Evaluation results show that GEMM performance get improved consistently
+  Evaluation results show that GEMM performance improves consistently
   under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$..

### 2018-05-25 scp.tex
Fix syntax errors according to the first revision by Prof. Xue

@@
 In Section~\ref{subsec:example}, we demonstrate SCP by an example.
-Then formal description is given in Section~\ref{subsec:formal}.
+Then a formal description is given in Section~\ref{subsec:formal}.
 
@@
 and a tuning process may be needed.
-After tuning we got the best values $M_r = 4$ and $N_r = 8$.
+After tuning we obtain the best values $M_r = 4$ and $N_r = 8$.
 By (\ref{eq:constraints.l1}),
@@
 Here we choose $M_c = 192$.
-As there is no L3 cache, (\ref{eq:constraints.l3}) need not to be
-concerned and $N_c$ can be given a large value.
+As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
+and $N_c$ can be given a large value.
 Here we choose $N_c = 1024NT = 4096$.
@@
 Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
-essentially represent two different partition style of the shared cache,
-refered as way-partition and set-partition, respectively.
-To guarantee the data layout in set-partition,
+essentially represent two different partitioning style of the shared cache,
+refered as way-partitioning and set-partitioning, respectively.
+To guarantee the data layout in set-partitioning,
 $A_2$ cannot be stored in a single continous buffer any more.
 Instead, $A_2$ is distributed to 12 continous memory segments, each of size $32KB$.
-Distance from one segment to the next is equal to $wc_2=128KB$.
+The distance from one segment to the next is equal to $wc_2=128KB$.
 %% There is no memory fragment because the segment size is a multiple of
@@
 %% FIXME put this to the end of this section
-While the set-partition of shared caches avoids inter-thread cache conflicts,
+While the set-partitioning of shared caches avoids inter-thread cache conflicts,
 it introduces some complexity to GEMM implementation
@@
 the packing of $B_2$ is done collabratively by all threads.
-We have two choices, privatize $B_2$ and apply the set-partition method,
-or fall back to conventional way-partition for $B_2$.
+We have two choices, privatize $B_2$ and apply the set-partitioning method,
+or fall back to conventional way-partitioning for $B_2$.
 $B_2$ can be made thread private if every thread makes
@@
 the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
-We assume that the share matrix $B_2$ is packed in conventional way-partition style.
-If desired, the $B_2$ can be privatized and handled in the same way as $A_2$.
+We assume that the share matrix $B_2$ is packed
+in conventional way-partitioning style.
+If desired, $B_2$ can be privatized and handled in the same way as $A_2$.
 
@@
 in which $s_i < e_i \in \mathcal{A}$.
-With way-partition, the memory descriptor contains only one segment.
-With set-partition used in previous section,
+With way-partitioning, the memory descriptor contains only one segment.
+With set-partitioning used in previous section,
 the memory descriptor contains a sequence of equally strided segments.
@@
 $\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
-but the part (size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
-The working flow of these three phases are listed
+but the part (of size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
+The working flow of these three phases is listed
 in Algorithm~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.
@@
 $align$ is initialized with the L1 cache line size $l_1$.
-Then on each level $l$ (Line~\ref{line:align.for})
-with shared non-LRU caches (Line~\ref{line:align.type}),
+Then on each level $l$ (line~\ref{line:align.for})
+with shared non-LRU caches (line~\ref{line:align.type}),
 $align$ is updated by computing the least-common-multiple of
@@
 each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
-in the set-partition of shared caches.
-At last, memory are allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
+in the set-partitioning of shared caches.
+At last, memory is allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
 and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
@@
 $\mathcal{D}_l^{idx}$ is partitioned in the conventional
-way-partition style.
+way-partitioning style.
 Otherwise, the else branch (line~\ref{line:else}) performs
-a set-partition on $\mathcal{D}_l^{idx}$.
+a set-partitioning on $\mathcal{D}_l^{idx}$.
 Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
@@
 divided among $NT$ threads (line~\ref{line:len}).
-Then $\mathcal{D}_s$ is partitioned in a way-partition style
+Then $\mathcal{D}_s$ is partitioned in a way-partitioning style
 and each thread $t$ get buffer space for its part in $B_2$

### 2018-05-25 background.tex
Syntax fixes

@@
 set-associative cache has been the dominant choice for real-world architectures.
-To understand how shared cache can cause inter-thread cache conflicts
+To understand how cache sharing can cause inter-thread cache conflicts
 and hurts GEMM performance, we review some details
-about the working mechanisim of set-associative caches.
+about the working mechanisim of set-associative caches and
+the memory hierachy of modern processors.
 
@@
 each capable of storing 2 double precision floating-point numbers.
-For convenience, the processor cores on level 0
+For convenience, a processor core on level 0
 is represented by its registers, which can be viewed as a special
@@
 sophisticated due to the presence of multi-level memory
-hierarchies on modern processors.
+hierarchy on modern processors.
 Figure.~\ref{fig:gemm} shows the structure of GEMM program from
@@
 
-As stated in Section~\ref{sec:intro}, an GEMM implementation
+As stated in Section~\ref{sec:intro}, a GEMM implementation
 is made up by a kernel routine and an overall strategy.
@@
 can be parallelized in a multi-threaded context.
-Specifically, the strategy in Figure.~\ref{fig:gemm} chooses a NKM loop order and
-parallelizes the M loop at layer 3.
+Specifically, the implementation in Figure.~\ref{fig:gemm}
+chooses a NKM loop order and parallelizes the M loop at layer 3.
 Different threads work on different parts of $A_1$
@@
 At time $t_s$, $T_0$ prefetches some data at address $addr_0$
-to cache set $\varphi_{L2}(addr_0)$ of the shared L2 cache.
+to cache set $\varphi_2(addr_0)$ of the shared L2 cache.
 The data will be actually read at time $t_e$.
@@
 $T_1$ prefetches data at addr $addr_1$ to L2 cache.
-If $\varphi_L(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,
+If $\varphi_2(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,
 data at $addr_1$ and $addr_0$ will be fetched to the same cache set.

### 2018-05-25 conclusion.tex
Syntax fixes

\section{Conclusion}\label{sec:conclusion}
In this paper, we present a shared cache partitioning method,
SCP, to eliminate inter-thread conflicts in GEMM programs
on architectures with shared caches.
The basic idea is partition the shared cache into physically
disjoint sets and assign different sets to different threads.
Evaluation results show that SCP can effectively reduce
cache misses on the shared cache (and upper level caches),
resulting in a consideration improvement in GEMM performance
in a multi-threaded context.
In future work, we plan to extend SCP to other dense linear
algebra computing routines.

### 2018-05-25 intro.tex
Syntax fixes

@@
 GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
-and BLIS~\cite{blis}\cite{blisport}.
+and BLIS~\cite{blis,blisport}.
 
@@
 of the most interest because all other level-3 operations can be defined
-in terms of GEMM and some level-1 and level-2 operations.
-Much effort has been spent on optimizing GEMM for different architectures.
+in terms of GEMM and some level-1 and level-2 operations~\cite{gemmbased1}.
+Much effort has been spent on optimizing GEMM for different
+architectures~\cite{Liu2012,Wang2015,Volkov:2008,Cui11,blispar}.
 %% both in general algorithm and in architecture specific ways.
-(FIXME citations)
 
-GEMM performs a matrix-multiply-accumulate operation.
+%% GEMM performs a matrix-multiply-accumulate operation.
 Generally, an optimized GEMM implementation consists of two components,
@@
 On the other hand, the overall strategy is responsible for optimizing
-memory accesses to protect the kernel from being idling waiting for data.
+memory accesses to protect the kernel from long memory latencies.
 In this paper, we focus on optimizing the memory access of GEMM.
@@
 To better exploit the capability of on-chip caches,
-matrix blocks the kernel operates on will be packed into continous memory buffer
+matrix blocks the kernel operates on are packed into continous memory buffer
 whose footprint is within the cache capacity before the kernel routine is called.
@@
 when working on architectures with non-LRU shared caches.
-In traditional high-performance processors, e.g. Intel Xeon series,
+In traditional high-performance processors, e.g. the Intel Xeon series,
 both L1 and L2 caches are private to processor cores
@@
 Evaluation results show that inter-thread cache conflicts on non-LRU shared caches
-can heavily harm GEMM performance,
-even with packed matrices shrinked to fit int the cache.
+can heavily hurt GEMM performance,
+even with packed matrices shrinked to fit into the cache.
 
@@
 \item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
-  an emerging high-performance processor based on ARM's AArch64 architecture,
-  which has 64 cores, private LRU L1 caches and psudo-random shared L2 caches.
+  an emerging high-performance processor based on ARM's AArch64 architecture.
+  Phytium 2000+ has 64 cores, private LRU L1 caches
+  and psudo-random shared L2 caches.
   Evaluation results show that GEMM performance improves consistently
-  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$..
+  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$.
 \end{itemize}

### 2018-05-25 related.tex
Syntax fixes

@@
 C code sequence into an optimized assembly code sequence.
+Both the auto-tuning and directive-based programming approaches
+rely on the compiler to transform kernels in C to machine instructions.
+Kernel performance can be improved with compiler techniques,
+including SIMD vectorization~\cite{Larsen:2000,Zhou:2016,Zhou:2016b,
+  Eichenberger2004,GCCSLP2007},
+polyhedral optimization~\cite{Bondhugula2008A,Kong:2013}, 
+and loop tiling~\cite{Lam1991,Spampinato:2014,Xue00}.
 \textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific

### 2018-05-25 scp.tex
Syntax fixes

@@
-\section{SCP: The Shared Cache Partition Method}\label{sec:scp}
+\section{SCP: The Shared Cache Partitioning Method}\label{sec:scp}
 
@@
 With tiling factors determined, the sizes of packed matrices
-can be calculated. Table.~\ref{tab:msizes} list the sizes of packed matrices.
+can be calculated. Table.~\ref{tab:msizes} list the sizes of
+various packed matrices.
 
@@
     \midrule
-    $A_2$, $A_3$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
-    $A_4$ & $es M_r K_c = 8KB$   & $B_3$, $B_4$ & $es N_r K_c = 16KB$ \\
+    %% $A_2$, $A_3$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
+    %% $A_4$ & $es M_r K_c = 8KB$   & $B_3$, $B_4$ & $es N_r K_c = 16KB$ \\
+    $A_2$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
+    $A_3$ & $es M_c K_c = 384KB$ & $B_3$ & $es N_r K_c = 16KB$ \\
+    $A_4$ & $es M_r K_c = 8KB$   & $B_4$ & $es N_r K_c = 16KB$ \\
     \bottomrule
@@
   \centering
-  \subfigure[Way-partition]{
+  \subfigure[Way-partitioning]{
     \label{fig:partition.conventional}
@@
   }
-  \subfigure[Set-partition]{
+  \subfigure[Set-partitioning]{
     \label{fig:partition.segmented}
@@
 $A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
-We see that every cache set contains data from all 4 $A_2$ instances,
+We see that every cache set $s \in [0,2048)$ contains
+data from all 4 $A_2$ instances,
 potentially leading to inter-thread cache conflicts.
@@
 Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
-essentially represent two different partitioning style of the shared cache,
-refered as way-partitioning and set-partitioning, respectively.
+essentially represent two different partitioning styles of the shared cache,
+refered to as way-partitioning and set-partitioning, respectively.
 To guarantee the data layout in set-partitioning,
@@
 How about $B_2$, the other packed matrix use in GEBP?
-Unlike $A_2$, which is thread private data,
+Unlike $A_2$, which is thread private,
 $B_2$ is shared among all threads and
@@
 the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
+Memory layout of other packed matrices on lower layers are
+determined by $A_2$ and $B_2$.
 We assume that the share matrix $B_2$ is packed
@@
 
-The memory layout of a packed matrix is described by a memory descriptor,
-which is essentially a subspace of the whole address space $\mathcal{A}$.
-Each packed matrix $M$ used by thread $t$,
-is associated with a memory descriptor $\mathcal{D}_t^M$
-specifying the memory space it occupies.
+We use a memory descriptor $\mathcal{D}_t^M$ to specify 
+the memory space occupied by packed matrix $M$ used by thread $t$.
+The memory descriptor is essentially a subspace of
+the whole address space, i.e. $\mathcal{D}_t^M \subset \mathcal{A}$.
+%% The memory layout of a packed matrix is described by a memory descriptor,
+%% which is essentially a subspace of the whole address space $\mathcal{A}$.
+%% Each packed matrix $M$ used by thread $t$,
+%% is associated with a memory descriptor $\mathcal{D}_t^M$
+%% specifying the memory space it occupies.
 Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,

### 2018-05-25 evaluation.tex
Add citation to Top500 website

@@
 Second, the LINPACK benchmark, which is used to build the
-TOP500 list of world's most powerful supercomputers (FIXME citation),
+TOP500 list of world's most powerful supercomputers~\cite{top500},
 relies on the DGEMM variant.

### 2018-05-26 abstract.tex
Fix spelling mistakes using spelling checker

@@
   a 64-core AArch64 processor with private LRU L1 caches
-  and shared (by every 4 cores) psudo-random L2 caches.
+  and shared (by every 4 cores) pseudo-random L2 caches.
   The results show that conflict misses on both L1 and L2

### 2018-05-26 background.tex
Fix spelling mistakes using spelling checker

@@
 
-\subsection{The Memory Hierachy Review}\label{subsec:hierachy}
+\subsection{The Memory Hierarchy Review}\label{subsec:hierachy}
 
@@
 and hurts GEMM performance, we review some details
-about the working mechanisim of set-associative caches and
-the memory hierachy of modern processors.
+about the working mechanism of set-associative caches and
+the memory hierarchy of modern processors.
 
@@
 
-It is easy to prove that any continous memory block no larger than $c/n$
+It is easy to prove that any continuous memory block no larger than $c/n$
 cannot conflict with itself.
@@
 The main memory, $NL$ levels of caches, and $NT$ processor cores
-form a memory hierachy generally organized into
+form a memory hierarchy generally organized into
 a $NL+2$ level regular tree structure.
 Processor cores at level $0$ are leaves,
-caches at levels $1$ -- $NL$ are intermedia nodes
+caches at levels $1$ -- $NL$ are intermediate nodes
 and the main memory at level $NL+1$ is the tree root.
-Any node in the memory hierachy can be identified by
+Any node in the memory hierarchy can be identified by
 its layer and its unique index on the layer $(layer, index)$.
@@
 $NL=2$ levels of caches as the platform for SCP demonstration.
-Figure.~\ref{fig:hierachy} shows its memory hierachy and
+Figure.~\ref{fig:hierachy} shows its memory hierarchy and
 detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
@@
   \includegraphics[width=.45\textwidth]{figures/cluster-new}
-  \caption{Memory hierachy of the 4-core processor
+  \caption{Memory hierarchy of the 4-core processor
     (nodes labeled with $(layer,index)$)}
@@
     L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
-    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & psudo-random \\
+    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & pseudo-random \\
     \bottomrule
@@
 with each iteration performing a rank-1 update on
-the $M_r \times N_r$ submatrix of $C$, as shown at layer 7.
+the $M_r \times N_r$ sub-matrix of $C$, as shown at layer 7.
 In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
@@
 Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
-matrices on each level fit into a certain level in the memory hierachy.
+matrices on each level fit into a certain level in the memory hierarchy.
 Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
@@
 $N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
-(the psudo level-0 cache).
+(the pseudo level-0 cache).
 Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
@@
 by a size constraint (column 3) related to a certain level
-in the memory hierachy (column 2).
+in the memory hierarchy (column 2).
 \begin{table}
@@
 $A_1[ii:ii+M_c-1][:]$ and $B_1$ are packed into $A_2$ and $B_2$
-in a special continous layout,
+in a special continuous layout,
 so consecutive memory access is ensured within the GEBP kernel.
@@
 is made up by a kernel routine and an overall strategy.
-In Figure.~\ref{fig:gemm}, GEBP surves as the kernel routine,
-and the overall strategy is defined by the three outter loops at layers 1 -- 3.
+In Figure.~\ref{fig:gemm}, GEBP serves as the kernel routine,
+and the overall strategy is defined by the three outer loops at layers 1 -- 3.
 Developers can control how to partition the workload by choosing
@@
 Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
-While all color shaded submatrices are accessed by $T_1$,
-only z-curve masked submatrices (a subset of color shaded ones)
+While all color shaded sub-matrices are accessed by $T_1$,
+only z-curve masked sub-matrices (a subset of color shaded ones)
 are packed by the thread.
 That is, each thread packs its own $A_2$ instance and 
-$B_2$ is packed by all thread collabratively.
+$B_2$ is packed by all thread collaboratively.
 %% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
@@
 As GEBP accesses the matrix data much more frequently than other
-data structures, this repeated prefething strategy
+data structures, this repeated prefetching strategy
 can reduce the penalty caused by intra-thread conflicts.
@@
 Then at some time $t_m$ ($t_s < t_m < t_e$),
-$T_1$ prefetches data at addr $addr_1$ to L2 cache.
+$T_1$ prefetches data at address $addr_1$ to L2 cache.
 If $\varphi_2(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,

### 2018-05-26 evaluation.tex
Fix spelling mistakes using spelling checker

@@
 Each curve represents a parallelism configuration.
-Different subfigures are configured with different $NT$s,
-and different curves in the same subfigure are
+Different sub-figures are configured with different $NT$s,
+and different curves in the same sub-figure are
 configured with different $NT_C$s.
 As there is only one configuration for $NT=64$ ($NC=16$, $NT_C=4$)
-and no comparation is available,
+and no comparison is available,
 results of $NT=64$ is combined with $NT=32$ in Figure.~\ref{fig:drawback-32}.
@@
 Figure.~\ref{fig:benefit} shows the evaluation results.
-For the purpose of comparation, results without SCP
+For the purpose of comparison, results without SCP
 (denoted as Base in the figure) are also presented,
@@
 $\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
-The hardware events are mesured in a per-thread manner
+The hardware events are measured in a per-thread manner
 and all presented results are the average over threads.
@@
 
-Privatization inccurs more packing overhead because
+Privatization introduces more packing overhead because
 redundant packing of $B_2$ is enforced.
@@
 Results under other parallelism configurations are similar.
-For the purpose of comparation,
+For the purpose of comparison,
 SCP with and without privatization are both presented
@@
 
-Results in this section demonstrats that
+Results in this section demonstrates that
 privatization of the shared matrix $B_2$ can further

### 2018-05-26 intro.tex
Fix spelling mistakes using spelling checker

@@
 level-1 for vector-vector operations,
-level-2 for matrix-vetor operations,
+level-2 for matrix-vector operations,
 and level-3 for matrix-matrix operations.
@@
 On the other hand, the overall strategy is responsible for optimizing
-memory accesses to protect the kernel from long memory latencies.
+memory accesses to protect the kernel from long memory latency.
 In this paper, we focus on optimizing the memory access of GEMM.
@@
 To better exploit the capability of on-chip caches,
-matrix blocks the kernel operates on are packed into continous memory buffer
+matrix blocks the kernel operates on are packed into continuous memory buffer
 whose footprint is within the cache capacity before the kernel routine is called.
@@
 shared cache may become common in future many-core processors.
-And non-LRU replacement policies, for example, psudo-random,
+And non-LRU replacement policies, for example, pseudo-random,
 may also be used to further reduce the cache design complexity.
@@
 and assign different sets to different threads. This can be achieved by
-exploiting the memory mapping mechanisim of set-associative caches.
+exploiting the memory mapping mechanism of set-associative caches.
 
@@
   Phytium 2000+ has 64 cores, private LRU L1 caches
-  and psudo-random shared L2 caches.
+  and pseudo-random shared L2 caches.
   Evaluation results show that GEMM performance improves consistently

### 2018-05-26 related.tex
Fix spelling mistakes using spelling checker

@@
-\section{Releated Work}\label{sec:related}
+\section{Related Work}\label{sec:related}
 The idea that all level-3 BLAS operations can be built upon
-a high-performance GEMM implemention was first proposed
+a high-performance GEMM implementation was first proposed
 in~\cite{gemmbased1,gemmbased2}.
@@
 POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
-directive-based programming approache.
+directive-based programming approach.
 Given a GEMM kernel in C, POET inserts annotations into
@@
 This nested parallelization can better fit complex architecture features
-like multi-sockets and hyperthreading.
+like multi-sockets and hyper-threading.

### 2018-05-26 scp.tex
Fix spelling mistakes using spelling checker

@@
 e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
-Similar to $M_r$ and $N_r$, a tunning process may be needed to
+Similar to $M_r$ and $N_r$, a tuning process may be needed to
 find the best value for $M_c$.
@@
 
-Each $A_2$ is packed into a single continous buffer.
+Each $A_2$ is packed into a single continuous buffer.
 Figure.~\ref{fig:partition.conventional} shows how the
@@
 essentially represent two different partitioning styles of the shared cache,
-refered to as way-partitioning and set-partitioning, respectively.
+referred to as way-partitioning and set-partitioning, respectively.
 To guarantee the data layout in set-partitioning,
-$A_2$ cannot be stored in a single continous buffer any more.
-Instead, $A_2$ is distributed to 12 continous memory segments, each of size $32KB$.
+$A_2$ cannot be stored in a single continuous buffer any more.
+Instead, $A_2$ is distributed to 12 continuous memory segments, each of size $32KB$.
 The distance from one segment to the next is equal to $wc_2=128KB$.
@@
 because the GEBP kernel and the matrix packing routines
-now must work with segmented memory buffers instead of continous ones.
+now must work with segmented memory buffers instead of continuous ones.
 
@@
 $B_2$ is shared among all threads and
-the packing of $B_2$ is done collabratively by all threads.
+the packing of $B_2$ is done collaboratively by all threads.
 We have two choices, privatize $B_2$ and apply the set-partitioning method,
@@
 Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
-and architectural details of the memory hierachy,
+and architectural details of the memory hierarchy,
 The SCP method systematically determines the
@@
 Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
-and $size_p$ is enlarged to a multile of $align$ (line~\ref{line:align}).
+and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
 $align$ is initialized with the L1 cache line size $l_1$.
@@
 (line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
-The $subspace$ procedure traverses the memory hierachy
+The $subspace$ procedure traverses the memory hierarchy
 to compute for each node, including the processor cores and main memory,
@@
 Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
-and traverses the memory hierachy in a depth-first-search order.
+and traverses the memory hierarchy in a depth-first-search order.
 

### 2018-05-26 background.tex
Fix e.g. to e.g., and i.e. to i.e.,

@@
 the constraints on these tiling factors.
-$es$ denotes the size of matrix element, e.g. 8B for double precision floating-point numbers,
+$es$ denotes the size of matrix element, e.g., 8B for double precision floating-point numbers,
 and $c_l$ denotes the capacity of registers or caches on level $l$.
@@
 data at $addr_0$ may suffer an eviction before it is actually read,
-i.e. a conflict miss occurs.
+i.e., a conflict miss occurs.
 

### 2018-05-26 evaluation.tex
Fix e.g. to e.g., and i.e. to i.e.,

@@
 The tiling factors are computed as in Section~\ref{subsec:example},
-i.e. $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
+i.e., $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
 The parallelism is controlled by two parameters,
@@
 in full privatization strategy, which is unacceptable
-if GEMM is highly parallelized, e.g. $NT=64$.
+if GEMM is highly parallelized, e.g., $NT=64$.
 Partial privatization in Figure.~\ref{fig:packb.partial}

### 2018-05-26 intro.tex
Fix e.g. to e.g., and i.e. to i.e.,

@@
 The HPC community has also contributed several high-quality
-BLAS implementations, e.g. ATLAS~\cite{atlas},
+BLAS implementations, e.g., ATLAS~\cite{atlas},
 GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
@@
 when working on architectures with non-LRU shared caches.
-In traditional high-performance processors, e.g. the Intel Xeon series,
+In traditional high-performance processors, e.g., the Intel Xeon series,
 both L1 and L2 caches are private to processor cores

### 2018-05-26 scp.tex
Fix e.g. to e.g., and i.e. to i.e.,

@@
 Suppose we are implementing a double precision GEMM (DGEMM) on
-the 4-core processor, i.e. $es=8B$.
+the 4-core processor, i.e., $es=8B$.
 We now determine the tiling factors used in DGEMM following the procedure described
@@
 Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
-e.g. $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
+e.g., $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
 It takes careful consideration to determine which one is the best,
@@
 it is reasonable to shrink $M_c$ to a slightly smaller value,
-e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
+e.g., $192$, $208$ and $224$ are all proper candidates for $M_c$.
 Similar to $M_r$ and $N_r$, a tuning process may be needed to
@@
 \begin{itemize}
-\item The architecture has 2 or 3 levels of cache. %% i.e. $NL=2$ or $NL=3$
-\item All caches are inclusive set-associative caches. %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
-\item Caches on the same level are homogeneous. %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
+\item The architecture has 2 or 3 levels of cache. %% i.e., $NL=2$ or $NL=3$
+\item All caches are inclusive set-associative caches. %% i.e., $n_l \ge 2$ for $l \in [1, NL]$
+\item Caches on the same level are homogeneous. %% i.e., $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
 \item A cache's way-capacity is a multiple of the sum of way-capacity of its children,
-i.e. $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
+i.e., $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
 \end{itemize}
@@
 The memory descriptor is essentially a subspace of
-the whole address space, i.e. $\mathcal{D}_t^M \subset \mathcal{A}$.
+the whole address space, i.e., $\mathcal{D}_t^M \subset \mathcal{A}$.
 %% The memory layout of a packed matrix is described by a memory descriptor,
@@
 Buffers of different threads are disjoint,
-i.e. $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
+i.e., $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
 The third phase computes memory descriptors for $B_2$.
@@
 among all children of node $(l, idx)$.
-If $subspace$ encounters a level 0 node, i.e. a processor core,
+If $subspace$ encounters a level 0 node, i.e., a processor core,
 then $\mathcal{D}_l^{idx}$ is exactly the thread-private

### 2018-05-26 abstract.tex
Revision from Prof. Xue

--- 2018-05-25 abstract.tex
@@
   Two techniques, software prefetching and data packing,
-  are utilized to effectively exploit the capability of on-chip caches.
-  These techniques are sufficient for lastly-recent-used (LRU) caches,
-  which are often used in high-performance
-  processors deployed in high-end servers and supercomputers.
-  But in recent years the market has become increasingly more diverse
-  in processor design, resulting in high-performance processors
+  have been used to effectively exploit the capability of on-chip lastly-recent-used (LRU) caches,
+  which are popular in traditional high-performance
+  processors used in high-end servers and supercomputers.
+  However, the market has recently witnessed a
+  new diversity in processor design, resulting in high-performance processors
   equipped with shared caches with non-LRU replacement policies.
-  This poses a challenge to GEMM development in a multi-threaded context.
-  As several threads try to load data into the shared cache simultaneously,
+  This poses a challenge to the development of
+  high-performance GEMM in a multi-threaded context.
+  As several threads try to load data into a shared cache simultaneously,
   inter-thread cache conflicts will increase significantly.
   We present a Shared Cache Partitioning (SCP) method to
-  eliminate inter-thread cache conflicts in GEMM routines,
-  by partitioning the shared cache into physically disjoint
-  sets and assign different sets to different threads.
-  We implement SCP in the OpenBLAS library and
-  evaluate the performance on Phytium 2000+,
+  eliminate inter-thread cache conflicts in the
+  GEMM routines,
+  by partitioning a shared cache into physically disjoint
+  sets and assigning different sets to different threads.
+  We have implemented SCP in the OpenBLAS library and
+  evaluated it on Phytium 2000+,
   a 64-core AArch64 processor with private LRU L1 caches
-  and shared (by every 4 cores) psudo-random L2 caches.
-  The results show that conflict misses on both L1 and L2
+  and shared pseudo-random L2 caches (per every four cores).
+  Our experimental results show that conflict misses in both L1 and L2
   caches are effectively reduced, resulting in an improvement

### 2018-05-26 background.tex
Revision from Prof. Xue

--- 2018-05-26 background.tex
@@
 
-\subsection{The Memory Hierarchy Review}\label{subsec:hierachy}
+\subsection{The Memory Hierarchy Review}\label{subsec:hierarchy}
 
@@
 to overcome the ``memory wall''.
-As a compromise between direct mapped cache and full-associative cache,
-set-associative cache has been the dominant choice for real-world architectures.
-To understand how cache sharing can cause inter-thread cache conflicts
+As a compromise between direct mapped caches
+and full-associative caches,
+set-associative caches have been the dominant choice for real-world architectures.
+To understand how cache sharing causes
+inter-thread cache conflicts
 and hurts GEMM performance, we review some details
@@
 $n$ is ways-of-associativity,
-and $nt$ means the number of cores sharing the cache.
+and $nt$ indicates the number of cores sharing the cache.
 The whole cache memory is divided into $ns=c/l/n$ sets,
-denoted by integral indices $\mathcal{S} = [0, ns)$,
+identified by indices in $\mathcal{S} = [0, ns)$,
 with each set containing $n$ cache lines.
@@
 mapping addresses from the address space $\mathcal{A}$
-(either virtual or physical) to cache sets $\mathcal{S}$.
-Data at $addr \in \mathcal{A}$ must be fetched
+(either virtual or physical) to the cache sets in
+$\mathcal{S}$.
+The data at $addr \in \mathcal{A}$ must be fetched
 into the cache set $\varphi(addr) \in \mathcal{S}$ if requested.
@@
 %% $\varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)$,
-with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively.
-It is an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$
-position in $addr$'s binary form.
+with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively:
 \begin{equation*}
@@
 \end{equation*}
+It is an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$
+position in $addr$'s binary form.
 
-It is easy to prove that any continuous memory block no larger than $c/n$
+It is easy to prove that any continuous memory block
+that is no larger than $c/n$
 cannot conflict with itself.
@@
 The main memory, $NL$ levels of caches, and $NT$ processor cores
-form a memory hierarchy generally organized into
-a $NL+2$ level regular tree structure.
-Processor cores at level $0$ are leaves,
-caches at levels $1$ -- $NL$ are intermediate nodes
+form a memory hierarchy are generally organized in terms
+of an $NL+2$ level regular tree structure.
+The processor cores at level $0$ are leaves,
+the caches at levels $1$ -- $NL$ are intermediate nodes,
 and the main memory at level $NL+1$ is the tree root.
 Any node in the memory hierarchy can be identified by
-its layer and its unique index on the layer $(layer, index)$.
-The number of nodes on layer $l$ is $NT / nt_l$.
+its layer and its unique index at the layer $(layer, index)$.
+The number of nodes at layer $l$ is $NT / nt_l$.
 
-Throughout this paper we will use a 4-core processor with
-$NL=2$ levels of caches as the platform for SCP demonstration.
-Figure.~\ref{fig:hierachy} shows its memory hierarchy and
-detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
+Throughout this paper, our example platform will be
+a 4-core processor with
+$NL=2$ levels of caches.
+Figure~\ref{fig:hierarchy} shows its memory hierarchy 
+and Table~\ref{tab:cluster} lists its
+architectural parameters. 
 Each core has 32 128-bit (16B) vector registers,
 each capable of storing 2 double precision floating-point numbers.
-For convenience, a processor core on level 0
+For convenience, a processor core at level 0
 is represented by its registers, which can be viewed as a special
@@
     (nodes labeled with $(layer,index)$)}
-  \label{fig:hierachy}
+  \label{fig:hierarchy}
 \end{figure}
@@
   \caption{Architectural parameters of the 4-core processor}
-  %% (see Section~\ref{subsec:hierachy} for meanings of symbols in the header row)
+  %% (see Section~\ref{subsec:hierarchy} for meanings of symbols in the header row)
   \label{tab:cluster}
@@
     \toprule
-    type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & policy \\
+    Type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & Policy \\
     \midrule
-    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & programmable \\
+    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & Programmable \\
     L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
-    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & pseudo-random \\
+    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & Pseudo-Random \\
     \bottomrule
@@
 hierarchy on modern processors.
-Figure.~\ref{fig:gemm} shows the structure of GEMM program from
+Figure~\ref{fig:gemm} shows the structure of GEMM program from
 the OpenBLAS library~\cite{openblas}.
@@
 with each iteration performing a rank-1 update on
-the $M_r \times N_r$ sub-matrix of $C$, as shown at layer 7.
+the $M_r \times N_r$ submatrix of $C$, as shown at layer 7.
 In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
@@
 matrices on each level fit into a certain level in the memory hierarchy.
-Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
-the constraints on these tiling factors.
-$es$ denotes the size of matrix element, e.g. 8B for double precision floating-point numbers,
-and $c_l$ denotes the capacity of registers or caches on level $l$.
-First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
-$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
-(the pseudo level-0 cache).
-Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
-$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
-at layer 5 fit into the L1 cache.
-Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
-$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
-at layer 4 fit into the L2 cache.
-At last, by (\ref{eq:constraints.l3}), $N_c$ is constrained that 
-$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
-Larger matrices on higher layers 1 -- 3 always reside in main memory.
-
+Equations~(\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3}) express
+the constraints on these tiling factors:
 \begin{eqnarray}
@@
 \end{eqnarray}
+Here,
+$es$ denotes the size of a matrix element, e.g.,
+8B for a double precision floating-point number,
+and $c_l$ denotes the capacity of a register or cache 
+at level $l$.
+By (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are so
+constrained that $M_r$ elements from $A$,
+$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into the registers
+(the pseudo level-0 cache).
+By (\ref{eq:constraints.l1}), $K_c$ is so
+constrained that $B_4$ ($N_r \times K_c$),
+$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
+at layer 5 fit into the L1 cache.
+By (\ref{eq:constraints.l2}), $M_c$ is so
+constrained that $A_3$ ($M_c \times K_c$),
+$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
+at layer 4 fit into the L2 cache.
+Finally, by (\ref{eq:constraints.l3}), $N_c$ is so
+constrained that 
+$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache.
+Larger matrices at higher layers 1 -- 3 always reside in main memory.
 
@@
 As stated in Section~\ref{sec:intro}, a GEMM implementation
-is made up by a kernel routine and an overall strategy.
-In Figure.~\ref{fig:gemm}, GEBP serves as the kernel routine,
-and the overall strategy is defined by the three outer loops at layers 1 -- 3.
-Developers can control how to partition the workload by choosing
+is made up by a kernel routine and an overall workload-partitioning strategy.
+In Figure~\ref{fig:gemm}, GEBP serves as the kernel routine,
+and the overall workload-partitioning strategy is defined by the three outer loops at layers 1 -- 3.
+Developers can decide how to partition the workload by choosing
 tile factors $M_c$, $N_c$ and $K_c$,
 and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
-Either the N loop at layer 1 or M loop at layer 3, or both of them,
+Either the N loop at layer 1 or the M loop at layer 3, or 
+both,
 can be parallelized in a multi-threaded context.
-Specifically, the implementation in Figure.~\ref{fig:gemm}
-chooses a NKM loop order and parallelizes the M loop at layer 3.
+Specifically, the implementation in Figure~\ref{fig:gemm}
+chooses an NKM loop order and parallelizes the M loop at layer 3.
 Different threads work on different parts of $A_1$
 and share the same $B_1$.
-As a result, there are multiple $A_2$ instances, one per thread,
-and a single $B_2$ instances, shared by all threads.
-To understand how the loop on layer 3 is parallelized on the 4-core processor, 
-Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
-While all color shaded sub-matrices are accessed by $T_1$,
-only z-curve masked sub-matrices (a subset of color shaded ones)
+As a result, there are multiple $A_2$ instances
+(one per thread)
+and a single $B_2$ instance (shared by all threads).
+To understand how the loop at layer 3 is parallelized on the 4-core processor, 
+Figure~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
+While all color shaded submatrices are accessed by $T_1$,
+only the z-curve masked submatrices (a subset of color shaded ones)
 are packed by the thread.
-That is, each thread packs its own $A_2$ instance and 
-$B_2$ is packed by all thread collaboratively.
+In other words, each thread packs its own $A_2$ instance and  all the threads collaboratively pack the 
+unique
+$B_2$ instance.
 %% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
@@
 For set-associative caches, cache misses can be categorized
-into capacity misses and conflict misses.
-While constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
-serve to avoid capacity misses,
+into capacity and conflict misses.
+While the constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
+aim to avoid capacity misses,
 conflict misses cannot be eliminated completely.
@@
 will be read $N_c/N_r$ times within a single GEBP execution,
-and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
+and is expected to reside in the 
+L2 cache during all $N_c/N_r$ iterations at layer 4.
 The packed matrix $B_4$ is similar in that it is expected to
-reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
+reside in the L1 cache during all $M_c/M_r$ iterations at layer 5.
 In general, even with carefully selected tile factors,
-matrix data can still be evicted out for a few reasons.
-For instance, data structures other than the packed matrices,
-or program code (in case of a unified cache) may be fetched into a
-cache set the matrix data lives in.
-We call this kind of cache conflicts intra-thread cache conflicts
+matrix data can still be evicted.
+For example, some data structures other than the packed matrices
+or program code (in case of a unified cache) may be fetched into the
+cache set where the matrix data reside in.
+Such cache conflicts are referred to as
+intra-thread cache conflicts
 because it occurs within the execution of a single thread.
 To deal with intra-thread conflicts,
-the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to L1 cache
+the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to the L1 cache
 in every iteration at layer 6 before it actually reads them in case of
@@
 In contrary to intra-thread cache conflicts,
-there exists another kind of cache conflicts,
+there exists another kind of cache conflicts, called
 inter-thread cache conflicts, on architectures with shared caches.
 Suppose two working threads, $T_0$ and $T_1$, are running independently
-on the 4-core processor in Figure.~\ref{fig:hierachy}
+on the 4-core processor in Figure.~\ref{fig:hierarchy}
 to accomplish a GEMM operation.
@@
 Then at some time $t_m$ ($t_s < t_m < t_e$),
-$T_1$ prefetches data at address $addr_1$ to L2 cache.
-If $\varphi_2(addr_1)$ happens to be equal to $\varphi_2(addr_0)$,
-data at $addr_1$ and $addr_0$ will be fetched to the same cache set.
-As the shared L2 cache has a non-LRU policy,
+$T_1$ prefetches data at addr $addr_1$ to the L2 cache.
+If $\varphi_2(addr_1)=\varphi_2(addr_0)$, then the
+data at $addr_1$ and $addr_0$ will be fetched into the same cache set.
+As the shared L2 cache has a non-LRU policy, the
 data at $addr_0$ may suffer an eviction before it is actually read,
@@
 
-In GEMM, the cache is utilized aggressively
-that almost all cache space is expected to be occupied by packed matrices.
+In GEMM, every cache is utilized aggressively
+so that almost all its cache space is expected to be occupied by packed matrices.
 Prefetch instructions from different threads are executed repeatedly,

### 2018-05-26 conclusion.tex
Revision from Prof. Xue

@@
 In this paper, we present a shared cache partitioning method,
-SCP, to eliminate inter-thread conflicts in GEMM programs
-on architectures with shared caches.
-The basic idea is partition the shared cache into physically
+SCP, to reduce inter-thread conflicts in GEMM programs
+on architectures with shared non-LRU caches.
+The basic idea is to partition a shared cache into physically
 disjoint sets and assign different sets to different threads.
 Evaluation results show that SCP can effectively reduce
-cache misses on the shared cache (and upper level caches),
-resulting in a consideration improvement in GEMM performance
+cache misses for a shared cache (and its
+upper level caches),
+resulting in a considerable improvement in GEMM performance
 in a multi-threaded context.
-In future work, we plan to extend SCP to other dense linear
-algebra computing routines.
+%In future work, we plan to extend SCP to other dense linear
+%algebra computing routines.

### 2018-05-26 evaluation.tex
Revision from Prof. Xue

--- 2018-05-25 evaluation.tex
@@
 
-We implement SCP in the OpenBLAS~\cite{openblas} library version 0.3.0-dev,
-and evaluate it on a Phytium 2000+ processor.
+We have implemented SCP in the OpenBLAS~\cite{openblas} library (version 0.3.0-dev) and evaluated
+it on a Phytium 2000+ processor.
 The Phytium 2000+ processor is an emerging high-performance
@@
 cluster containing 4 cores.
-The structure of the cluster is almost the same with
-the contrived 4-core processor in Figure.~\ref{fig:hierachy}
-except that L2 caches of all 16 clusters connect to the main memory
+The structure of a cluster is almost the same as
+the 4-core processor in Figure~\ref{fig:hierarchy}
+except that the L2 caches of all 16 clusters connect to the main memory
 with hardware coherence.
-Table.~\ref{tab:cluster} lists architectural parameters of the cores and caches.
+Table~\ref{tab:cluster} lists the
+architectural parameters for the cores and caches
+in this processor.
 We restrict our evaluation to DGEMM,
@@
 
-All evaluations run with the same tiling factors
-$M_r$, $N_r$, $K_c$, $M_c$ and $N_c$.
-The tiling factors are computed as in Section~\ref{subsec:example},
-i.e. $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
+In our evaluation, we use the tiling factors
+computed as in Section~\ref{subsec:example}:
+$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
 The parallelism is controlled by two parameters,
-the number of active clusters $NC$,
+the number of active clusters $NC$
 and the number of active threads per cluster $NT_C$.
 The total number of threads is $NT = NC \cdot NT_C$.
-For the Phytium 2000+ processor, the valid ranges for $NC$ and $NT_C$
-are $NC \in [1, 16]$ and $NT_C \in [1, 4]$, respectively.
+For the Phytium 2000+ processor, we have
+$NC \in [1, 16]$ and $NT_C \in [1, 4]$. 
 Generally, DGEMM performance is presented in $flops$.
@@
 the average thread efficiency,
-to present performance results consistently.
+to describe our performance results consistently.
 The average thread efficiency is a normalized value computed as
-$E_{avg} = flops / (NT \cdot \widehat{flops})$
-in which $\widehat{flops}$ stands for the theoretical performance peak of a single core.
+$E_{avg} = flops / (NT \cdot \widehat{flops})$,
+where
+$\widehat{flops}$ stands for the theoretical performance peak of a single core.
 %% \begin{equation*}
@@
 
-We first show how cache sharing can hurt GEMM performance
-in Section~\ref{subsec:drawback}.
-Then Section~\ref{subsec:benefit} demonstrates how SCP can reduce
-the penalty caused by cache sharing.
-Quantitative analysis of the cache miss rates
-is presented in Section~\ref{subsec:analysis}.
-Finally, Section~\ref{subsec:privb} discusses
-the privatization of shared matrix $B_2$.
+We show how cache sharing can hurt GEMM performance
+(Section~\ref{subsec:drawback}),
+demonstrate how SCP can reduce
+the penalty caused by cache sharing 
+(Section~\ref{subsec:benefit}), conduct a 
+quantitative analysis of the cache miss rates
+(Section~\ref{subsec:analysis}), and
+discuss the privatization of shared matrix $B_2$
+(Section~\ref{subsec:privb}).
 %% Three versions of DGEMM are evaluated:
@@
 
-As the L2 cache is shared by cores inside the cluster,
+As the L2 cache is shared by cores inside a cluster,
 the impact of cache sharing can be demonstrated
-by comparing DGEMM performance with the same $NT$
-but different $NT_C$s.
-Figure.~\ref{fig:drawback} shows the evaluation results.
+by comparing the DGEMM performance results achieved
+with the same $NT$
+but different $NT_C$'s.
+Figure~\ref{fig:drawback} shows the results.
 Each curve represents a parallelism configuration.
-Different subfigures are configured with different $NT$s,
+Different subfigures are configured with different
+$NT$'s,
 and different curves in the same subfigure are
-configured with different $NT_C$s.
-As there is only one configuration for $NT=64$ ($NC=16$, $NT_C=4$)
-and no comparation is available,
-results of $NT=64$ is combined with $NT=32$ in Figure.~\ref{fig:drawback-32}.
+configured with different $NT_C$'s.
+As there is only one configuration for $NT=64$ ($NC=16$ and $NT_C=4$)
+and no comparison is available, the
+results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.
 
@@
 
-From Figure.~\ref{fig:drawback} we can see that
-with all $NT$s, $NT_C=1$ achieves the best performance,
-and $NT_C=2$ follows, and $NT_C=4$ comes at the lowest efficiency.
-Table.~\ref{tab:drawback} summarizes the average value
-of $E_{avg}$ over matrix sizes.
-The table can be viewed in two directions.
-Vertically, results in the same column shows how $E_{avg}$ varies with $NT_C$.
-With all $NT$s, $NT_C=1$ outperforms $NT_C=4$ by a large margin.
-The gap is $7.4\%$ when $NT=4$ and roughly $10\%$ when $NT>4$.
-Horizontally, results in the same row reflects the scalability of
-DGEMM with certain $NT_C$s.
+We can see that for all $NT$'s, $NT_C=1$ achieves the 
+best performance, followed by 
+$NT_C=2$, and $NT_C=4$ comes at the lowest efficiency.
+Table~\ref{tab:drawback} summarizes the average value
+of $E_{avg}$ across all the matrix sizes.
+Vertically, the results in the same column show how $E_{avg}$ varies with $NT_C$.
+For all $NT$'s, $NT_C=1$ outperforms $NT_C=4$ by a large margin.
+The performance gap is $7.4\%$ when $NT=4$ and roughly $10\%$ when $NT>4$.
+Horizontally,  the results in the same row reflect the scalability of
+DGEMM with certain $NT_C$'s.
 With $NT_C=1$, the performance scales linearly with the number of threads.
-With $NT_C=2$, $E_{avg}$ suffers a $1.86\%$ lose when $NT$ grows from 4 to 32.
-The situation is much worse for $NT_C=4$, $E_{avg}$ drops by $14.19\%$ from
-$NT=4$ to $NT=64$!
+With $NT_C=2$, $E_{avg}$ suffers a $1.86\%$ loss when $NT$ grows from 4 to 32.
+The situation is much worse for $NT_C=4$, in which
+$E_{avg}$ drops by $14.19\%$ from
+$NT=4$ to $NT=64$.
 
-Evaluation results above show that cache sharing has a considerable
+This shows clearly that cache sharing has a considerable
 impact on DGEMM performance.
@@
   }
-  \caption{Average thread efficiency of SCP}
+  \caption{Average thread efficiency under SCP}
   \label{fig:benefit}
@@
 to partition the cache if $NT_C=1$.
-Figure.~\ref{fig:benefit} shows the evaluation results.
-For the purpose of comparation, results without SCP
-(denoted as Base in the figure) are also presented,
+Figure~\ref{fig:benefit} shows the results.
+For comparison purposes, the results without SCP
+(denoted as Base) are also presented.
 
@@
   \centering
-  \caption{Average thread efficiency improvement of SCP}
+  \caption{Average thread efficiency improvement under SCP}
   \label{tab:win}
@@
 
-Under all parallelism configurations, 
-SCP performs better than the Base
-consistently over all matrix sizes.
-Table.~\ref{tab:win} summarizes the performance improvement of SCP over Base.
-%% Table.~\ref{tab:benefit} summarizes the average value
+Under all the parallelism configurations, 
+SCP performs better than Base
+consistently over all the matrix sizes evaluated.
+Table~\ref{tab:win} summarizes the performance improvement of SCP over Base.
+%% Table~\ref{tab:benefit} summarizes the average value
 %% of $E_{avg}$ in SCP evaluation.
@@
 With $NT_C=2$, SCP performs slightly better than Base
-by $1.7\%$--$1.8\%$.
-The performance gap becomes larger with $NT_C=4$.
+by $1.7\%$ -- $1.8\%$.
+The performance gap becomes larger with $NT_C=4$,
 ranging from $2.75\%$ to $6.91\%$.
@@
 is observed under the maximum parallelism $NT=64$.
-%% Looking at Table.~\ref{tab:benefit} vertically,
+%% Looking at Table~\ref{tab:benefit} vertically,
 %% we can see that SCP still cannot fully eliminate the impact
@@
 
-To understand why SCP outperforms Base,
-and whether SCP actually eliminate inter-thread conflicts,
-we use the PAPI~\cite{papi} profiling tool to
-analyze the cache behavior of the GEMM program.
-Four cache related hardware events are counted,
-which are listed and briefly described in Table.~\ref{tab:events}.
-With these hardware events,
-miss rates of L1 and L2 caches can be calculated with
+To understand why SCP outperforms Base and
+why SCP is effective in reducing
+inter-thread conflicts, 
+we use the PAPI profiling tool~\cite{papi} to
+analyze the cache behavior of GEMM. A total of 
+four cache-related hardware events are counted,
+which are listed and briefly described in Table~\ref{tab:events}.
+Based on these hardware events, the
+miss rates of L1 and L2 caches can be estimated  as 
 $\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
-The hardware events are mesured in a per-thread manner
-and all presented results are the average over threads.
+The hardware events are measured in a per-thread manner
+and all the results are the average measurements
+over all the threads.
 
@@
 
-Figure.~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.
-For the sake of space, we only present results with $NT_C=4$
+Figure~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.  We only give the results with $NT_C=4$
 as SCP achieves the largest improvement with full cache sharing.
-Table.~\ref{tab:papi} summarizes the average value of cache miss rates
-over matrix sizes.
-Under all parallelism configurations,
-both L1 and L2 cache miss rates are reduced with SCP applied.
-The reduced amount is roughly $13\%$ for L1 and nearly $10\%$ for L2.
+Table~\ref{tab:papi} summarizes the average of the
+cache miss rates
+across the matrix sizes used.
+Under all the parallelism configurations,
+both L1 and L2 cache miss rates are reduced under
+SCP, 
+by roughly $13\%$ for L1 and nearly $10\%$ for L2.
 
@@
 only L2 caches are specially handled by SCP.
-From the results we see that miss rates on both L1 and L2 caches
+The miss rates for both L1 and L2 caches
 are reduced effectively by SCP.
-While the decrease of L2 miss rate is expected,
-why L1 miss rate is also affected?
-The reason is that the L2 cache is inclusive.
-Conflict misses on the L2 cache not only evict data out from L2,
+While a reduction on the L2 miss rate is expected,
+why is the L1 miss rate also reduced? In this processor,
+every L2 cache is inclusive.
+The conflict misses for the L2 cache not only cause
+the data to be evicted from L2,
 but also invalidate the evicted data's copy in all upper level L1 caches,
-thus leading to more L1 cache misses.
+resulting in more L1 cache misses.
 In general, conflict misses on an inclusive cache can affect
-performance of all upper level caches.
+the performance of all upper level caches. As a 
+result, improving the L2 miss rate also improves the
+L1 miss rate.
 
-Results in this section prove that SCP can effectively
-eliminate conflict misses on shared caches.
-And the benefit propagates to upper level caches, too.
+Our results show that SCP can effectively
+reduce the conflict misses in shared L2 caches.
+In addition, the performance benefit thus obtained
+also propagates to the upper level caches, too.
 
@@
 \subsection{Privatization of Shared Matrix}\label{subsec:privb}
-In previous evaluations, SCP only set-partitions the shared cache
-for $A_2$ matrices which are private to threads.
+In the previous evaluations, SCP only set-partitions the shared cache
+for $A_2$ matrices, which are private to threads.
 The shared $B_2$ matrix is still packed
-in conventional way-partitioning style.
+in a conventional way-partitioning style.
 As mentioned in Section~\ref{subsec:example},
@@
 
-Privatization inccurs more packing overhead because
+Privatization incurs more packing overhead because
 redundant packing of $B_2$ is enforced.
-Figure.~\ref{fig:packb} shows three packing strategies for $B_2$
-in a parallelism configuration $NT_C=4$ and $NC=4$.
+Figure~\ref{fig:packb} shows three packing strategies for $B_2$
+in a parallelism configuration specified by
+$NT_C=4$ and $NC=4$.
 The whole packing workload of $B_2$ is divided into 16 tasks
 and the packing workload of thread $T_1$ is masked by z-curves. 
-%% and the task assignment is annoted in Figure.~\ref{fig:packb}.
-In conventional packing, each thread takes a single task.
+%% and the task assignment is annoted in Figure~\ref{fig:packb}.
+In the conventional packing approach, each thread takes a single task.
 In full privatization, all tasks are replicated to all threads.
 As a result, the packing overhead grows in proportion to $NT$
-in full privatization strategy, which is unacceptable
-if GEMM is highly parallelized, e.g. $NT=64$.
-Partial privatization in Figure.~\ref{fig:packb.partial}
+in a full privatization strategy, which is unacceptable
+if GEMM is highly parallelized, e.g., when $NT=64$.
+Partial privatization in Figure~\ref{fig:packb.partial}
 offers an alternative to full privatization.
 In partial privatization, privatization only occurs inside
-a cluster so that tasks belong to one cluster need not to be
+a cluster so that the tasks for one cluster need not to be
 replicated to other clusters.
-Consequently, extra packing overhead of the partial privatization
+Consequently,  the extra packing overhead of the partial privatization
 strategy is bounded by $NT_C$.
-For instance, on Phytium 2000+, overhead of packing $B_2$
+For example, on Phytium 2000+, overhead of packing $B_2$
 in partial privatization is limited to 4 times
@@
 
-Figure.~\ref{fig:privb} shows the evaluation results of the partial
+Figure~\ref{fig:privb} shows the results of the partial
 privatization strategy with $NT_C=4$ and $NC=4$.
-Results under other parallelism configurations are similar.
-For the purpose of comparation,
+The results under other parallelism configurations are similar.
+For comparison purposes,
 SCP with and without privatization are both presented
-(denoted as SCP-P and SCP in Figure.~\ref{fig:privb}, respectively).
-Figure.~\ref{fig:privb.papi} shows the cache miss rates.
-Privatization achieves slightly lower miss rates on both L1 and L2 caches.
-Figure.~\ref{fig:privb.ate} shows the average thread efficiency.
+(denoted as SCP-P and SCP in Figure~\ref{fig:privb}, respectively).
+Figure~\ref{fig:privb.papi} shows the cache miss rates.
+Privatization achieves slightly lower miss rates for
+both L1 and L2 caches.
+Figure~\ref{fig:privb.ate} shows the average thread efficiency.
 Despite its lower cache miss rates,
-privatization suffers a performance lose compared to SCP.
-Why performance drops while cache miss rates get reduced?
+privatization suffers a performance loss compared to SCP.
+Why is there a performance drop when cache miss rates 
+are reduced?
 The reason lies in the extra packing overhead introduced by privatization.
-Figure.~\ref{fig:privb.breakdown} shows the occupancy ratio
-of overall executing time of various overheads.
-Overheads in GEMM includes
+Figure~\ref{fig:privb.breakdown} shows the occupancy
+ratios of the overall execution times for various overheads.
+Overheads in GEMM include
 (1) packing of $A_2$, (2) packing of $B_2$,
@@
 it spends much more time in packing $B_2$.
-Table.~\ref{tab:breakdown} lists the breakdown of
-executing time. We can see that overhead of packing $B_2$
+Table~\ref{tab:breakdown} gives a breakdown of the
+execution time, indicating that the overhead for packing $B_2$
 is multiplied by a factor of $NT_C=4$,
@@
 
-Results in this section demonstrats that
-privatization of the shared matrix $B_2$ can further
-eliminate inter-thread cache conflicts.
-But the improvement in cache performance is very small because
+Our results demonstrate that
+privatizing the shared matrix $B_2$ can further
+reduce inter-thread cache conflicts.
+However,
+the improvement in cache performance is small. 
+As
 $B_2$ only occupies a small portion of the shared L2 cache,
-and this marginal improvement is not sufficient to
+this marginal improvement is not sufficient to
 balance the extra packing overhead introduced

### 2018-05-26 intro.tex
Revision from Prof. Xue

--- 2018-05-26 intro.tex
@@
 Processor vendors often provide BLAS implementations
-highly optimized for their processors,
+that are highly optimized for their processors,
 such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
@@
 Among the three levels of BLAS routines, level-3 provides the most opportunities
-for optimization because it has the highest computational complexity ($O(n^3)$).
+for optimization because it has the highest computational complexity (of $O(n^3)$).
 Among all the level-3 operations, GEneral Matrix Multiply (GEMM) is
-of the most interest because all other level-3 operations can be defined
+of the most interest because the 
+other level-3 operations can be defined
 in terms of GEMM and some level-1 and level-2 operations~\cite{gemmbased1}.
-Much effort has been spent on optimizing GEMM for different
+In the past, much effort has been spent on optimizing GEMM for different
 architectures~\cite{Liu2012,Wang2015,Volkov:2008,Cui11,blispar}.
@@
 %% GEMM performs a matrix-multiply-accumulate operation.
-Generally, an optimized GEMM implementation consists of two components,
-(1) a highly optimized kernel routine to accomplish the computation, and
+An optimized GEMM implementation consists of two components,
+(1) a highly optimized kernel routine to accomplish its
+computation, and
 (2) an overall strategy to partition the workload into small tasks
@@
 on matrices from a single task.
-The overall strategy partitions the workload by tiling the loops,
+The overall strategy partitions the workload by tiling 
+the underlying loop nest
 and determines a task schedule by choosing a specific traversal order
@@
 different threads to parallelize the whole operation.
-The main target of GEMM optimization is maximizing the floating-point
+The main objective of GEMM optimization is to
+maximize the floating-point
 operation throughput, measured by FLoating-point Operations Per Second ($flops$).
-On one hand, optimization of the serial kernel routine concentrates on
+For the serial kernel routine, the optimization applies
 instruction scheduling to improve instruction throughput.
-On the other hand, the overall strategy is responsible for optimizing
-memory accesses to protect the kernel from long memory latency.
-In this paper, we focus on optimizing the memory access of GEMM.
+For the workload partitioning, the optimization 
+reorganizes memory accesses to reduce long memory latencies for
+the kernel.
+In this paper, we focus on optimizing the memory 
+accesses for GEMM.
 
-Currently, two techniques, software prefetching and data packing,
-are widely adopted in GEMM implementations to speed up memory accesses.
+Two techniques, software prefetching and data packing,
+have been widely used in GEMM implementations to speed up memory accesses.
 Software prefetch instructions are utilized
-to load data into cache before they are used, so memory accesses will not
-cause long latencies and floating-point instructions can execute at peak throughput.
+to load data into cache before they are used, in 
+order to ensure that memory accesses will not incur
+long latencies and floating-point instructions
+can execute at peak throughput.
 To better exploit the capability of on-chip caches,
-matrix blocks the kernel operates on are packed into continuous memory buffer
-whose footprint is within the cache capacity before the kernel routine is called.
+matrices are blocked and then packed into continuous memory buffers
+that can fit into the caches under consideration
+before the kernel routine is called.
 Because GEMM is a computational intensive operation whose
@@
 
-The software prefetching and data packing approach meet a challenge
-when working on architectures with non-LRU shared caches.
-In traditional high-performance processors, e.g. the Intel Xeon series,
+However, software prefetching and data packing cannot be
+effectively
+applied to architectures with non-LRU shared caches.
+In traditional high-performance processors, e.g.,
+the Intel Xeon series,
 both L1 and L2 caches are private to processor cores
@@
 Without considering the last level cache (LLC),
-threads running on different cores prefetch their data into
+the threads running on different cores prefetch their data into
 different caches, and private data of different threads
 cannot cause conflict cache misses.
-This is not the case if caches are shared by several processor cores.
-There are chances that a cache line prefetched by one thread
-get evicted out before its lifetime ends because another thread prefetches
+This is no longer the case if caches are shared by several processor cores.
+In this case, a cache line prefetched by one thread
+may be evicted before its lifetime ends because another thread prefetches
 another cache line into the same cache set.
-A simple solution would be shrinking the size of packed matrix
-so data used by several threads can simultaneously reside in the shared cache.
+A simple-minded solution would be to reduce 
+the size of a packed matrix
+so that the data used by several threads can simultaneously reside in a shared cache.
 Due to the nature of set-associative caches,
@@
 especially if the shared cache has a non-LRU replacement policy.
-Evaluation results show that inter-thread cache conflicts on non-LRU shared caches
+Our evaluation shows that inter-thread cache conflicts
+in non-LRU shared caches
 can heavily hurt GEMM performance,
-even with packed matrices shrinked to fit into the cache.
+even with packed matrices reduced to fit into the cache.
 
@@
 To reduce cost of on-chip cache memories and coherence networks,
-shared cache may become common in future many-core processors.
-And non-LRU replacement policies, for example, pseudo-random,
+shared caches may become common in future many-core processors.
+In addition, non-LRU replacement policies, for example, pseudo-random,
 may also be used to further reduce the cache design complexity.
-Indeed, some high-performance processors based on ARM technology,
+Indeed, some high-performance processors based on the
+ARM technology,
 have already adopted this design. (FIXME list the processors)
-As a consequence, a better solution to the inter-thread cache conflicts problem
-on non-LRU shared caches is meaningful.
+As a result, developing a better solution to reduce
+the inter-thread cache conflicts 
+for non-LRU shared caches is important.
 
-In this paper, we present a Shared Cache Partitioning (SCP) method,
+In this paper, we present a Shared Cache Partitioning (SCP) method
 to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
-The key idea is partition the share cache into physically disjoint sets
+The key idea is to partition a
+share cache into physically disjoint sets
 and assign different sets to different threads. This can be achieved by
@@
 
-The main contributions of this paper are:
+To the best of our knowledge, SCP represents
+the first work addressing
+the inter-thread cache conflicts problem on
+architectures with non-LRU share caches.
+
+The main contributions of this paper are as follows:
 \begin{itemize}
@@
   problem on architectures with non-LRU shared caches.
-\item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
+\item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
   an emerging high-performance processor based on ARM's AArch64 architecture.
@@
   and pseudo-random shared L2 caches.
-  Evaluation results show that GEMM performance improves consistently
-  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$.
+  Our evaluation shows that SCP can improve the
+  GEMM performance consistently
+  under various parallelism configurations, by $2.75\%$--$6.91\%$.
 \end{itemize}
 
-To the best of our knowledge, SCP is the first proposal targeted at
-solving the inter-thread cache conflicts problem on
-architectures with non-LRU share caches.
-
 The rest of the paper is organized as follows.
-Section~\ref{sec:background} gives some background on GEMM and
-the cache conflicts problem in GEMM problem.
-Section~\ref{sec:scp} illustrates the SCP method.
-Section~\ref{sec:evaluation} presents evaluation results.
+Section~\ref{sec:background} gives some background on GEMM and discusses
+the cache conflicts incurred.
+Section~\ref{sec:scp} introduces the SCP method.
+Section~\ref{sec:evaluation} presents and
+analyzes our experimental results.
 Section~\ref{sec:related} reviews the related work.

### 2018-05-26 related.tex
Revision from Prof. Xue

--- 2018-05-25 related.tex
@@
-\section{Releated Work}\label{sec:related}
-The idea that all level-3 BLAS operations can be built upon
-a high-performance GEMM implemention was first proposed
+\section{Related Work}\label{sec:related}
+The idea that all level-3 BLAS operations can be built 
+on top of 
+a high-performance GEMM implementation was first proposed
 in~\cite{gemmbased1,gemmbased2}.
@@
 One is developing fast computation kernels and
-the other is choosing a proper overall strategy
+the other is choosing a proper overall workload-partitioning  strategy
 to optimize memory access.
@@
 In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
-the kernel component of GEMM are written by domain experts in assembly.
+the kernel component of GEMM is 
+written by domain experts in assembly.
 In the case of auto-tuning,
@@
 on the actual computing system.
-POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
-directive-based programming approache.
+POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} use a 
+directive-based programming approach.
 Given a GEMM kernel in C, POET inserts annotations into
@@
 
-The overall strategy of GEMM mainly concerns about
-tiling factors, loop orders and parallelization.
+The overall workload-partitioning strategy used in
+GEMM is mainly concerned with choosing 
+tiling factors, loop orders and parallelization
+techniques.
 The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
@@
 values for these factors.
-Analytic techniques~\cite{analytic1,analytic2,blisanalytic} are
+Analytic techniques~\cite{analytic1,analytic2,blisanalytic} 
 can also be used instead of auto-tuning.
@@
 address the problem of cache sharing with non-LRU replacement policies,
-and SCP serves as the first proposal to handle this problem.
-\cite{gotogemm} gives a detailed discussion of different looping orders.
-GEMM are usually parallelized only on layer 3 (the M loop),
+and SCP serves as the first work to address this problem.
+In \cite{gotogemm}, a detailed discussion on
+choosing
+different loop orders is given.
+GEMM are usually parallelized only at
+layer 3 (the M loop),
 but all loops along the M and N dimensions are potentially parallelizable.
-BLIS\cite{blispar} allows the users to specify a sophisticated configuration
-that more than one loops of layer 1, 3, 4 and 5
+BLIS\cite{blispar} allows developers to specify a sophisticated configuration
+so 
+that more than one loops at layers 1, 3, 4 and 5
 can be simultaneously parallelized.
-This nested parallelization can better fit complex architecture features
+This nested parallelization is well suited to complex architecture features
 like multi-sockets and hyperthreading.

### 2018-05-26 scp.tex
fix the symbol of empty set
Revision from Prof. Xue

@@
 
-This section demonstrates the SCP method to handle inter-thread cache conflicts.
-The basic idea is that inter-thread cache conflicts can be eliminated
-if packed matrices used by different threads are fetched to different cache sets.
-In Section~\ref{subsec:example}, we demonstrate SCP by an example.
-Then a formal description is given in Section~\ref{subsec:formal}.
+This section introduces our SCP method for reducing
+inter-thread cache conflicts.
+The key insight is that inter-thread cache conflicts can be eliminated
+if packed matrices used by different threads are fetched to different cache sets in a shared L2 cache.
+In Section~\ref{subsec:example}, we illustrate 
+SCP by an example.  In Section~\ref{subsec:formal},
+we give the algorithms in our SCP method.
 
-\subsection{Example}\label{subsec:example}
+\subsection{An Example}\label{subsec:example}
 
-Suppose we are implementing a double precision GEMM (DGEMM) on
-the 4-core processor, i.e., $es=8B$.
-We now determine the tiling factors used in DGEMM following the procedure described
-by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
-Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
-e.g., $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
-It takes careful consideration to determine which one is the best,
-and a tuning process may be needed.
-After tuning we obtain the best values $M_r = 4$ and $N_r = 8$.
+Suppose we want to implement a double precision GEMM (DGEMM) on
+the 4-core processor, where $es=8B$.
+We now determine the tiling factors used based on
+the constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3}).
+Many solutions for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}), including
+$(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. Auto
+tuning returns
+$M_r = 4$ and $N_r = 8$ as the best.
 By (\ref{eq:constraints.l1}),
 $K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
-Here we choose $K_c=256$.
+Here, we choose $K_c=256$.
 By (\ref{eq:constraints.l2}),
@@
 \lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
-To leave space for other data structures, as well as the program code,
-it is reasonable to shrink $M_c$ to a slightly smaller value,
-e.g., $192$, $208$ and $224$ are all proper candidates for $M_c$.
-Similar to $M_r$ and $N_r$, a tuning process may be needed to
-find the best value for $M_c$.
-Here we choose $M_c = 192$.
+To leave space for other data structures and the program code,
+it is reasonable to shrink $M_c$ slightly, with,
+for example, $192$, $208$ and $224$ as good
+candidates. Again, a turning process yields
+$M_c = 192$.
 As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
 and $N_c$ can be given a large value.
-Here we choose $N_c = 1024NT = 4096$.
-With tiling factors determined, the sizes of packed matrices
-can be calculated. Table.~\ref{tab:msizes} list the sizes of
-various packed matrices.
+Here, we choose $N_c = 1024NT = 4096$.
+With these tiling factors determined, the sizes of packed matrices
+can be calculated. Table.~\ref{tab:msizes} lists
+the sizes of various packed matrices used.
 
@@
     \toprule
-    matrix & size & matrix & size \\
+    Matrix & Size & Matrix & Size \\
     \midrule
@@
 Each $A_2$ is packed into a single continuous buffer.
-Figure.~\ref{fig:partition.conventional} shows how the
+Figure~\ref{fig:partition.conventional} shows how the
 $A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
-We see that every cache set $s \in [0,2048)$ contains
+Every cache set $s \in [0,2048)$ contains
 data from all 4 $A_2$ instances,
@@
 What if the $A_2$ matrices are distributed in the way shown in
-Figure.~\ref{fig:partition.segmented}?
-$A_2$ of different threads live in strictly disjoint cache sets,
-and inter-thread cache conflicts caused by $A_2$ are eliminated completely!
+Figure~\ref{fig:partition.segmented}?
+$A_2$ of different threads live in strictly disjoint cache sets, enabling
+the inter-thread cache conflicts caused by $A_2$ to be
+eliminated completely!
 
-Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
-essentially represent two different partitioning styles of the shared cache,
-referred to as way-partitioning and set-partitioning, respectively.
-To guarantee the data layout in set-partitioning,
-$A_2$ cannot be stored in a single continuous buffer any more.
-Instead, $A_2$ is distributed to 12 continuous memory segments, each of size $32KB$.
+Figures~\ref{fig:partition.conventional} and~\ref{fig:partition.segmented}
+essentially represent two different partitioning styles for a shared cache,
+referred to as \emph{way-partitioning} and \emph{set-partitioning}, respectively.
+To obtain the data layout in set-partitioning,
+$A_2$ can no longer be stored in a single continuous buffer.
+Instead, $A_2$ will be distributed to 12 continuous memory segments, each of size $32KB$.
 The distance from one segment to the next is equal to $wc_2=128KB$.
@@
 %% FIXME put this to the end of this section
-While the set-partitioning of shared caches avoids inter-thread cache conflicts,
-it introduces some complexity to GEMM implementation
+While avoiding inter-thread cache conflicts,
+set-partitioning
+introduces some complexity to GEMM implementations
 because the GEBP kernel and the matrix packing routines
@@
 
-How about $B_2$, the other packed matrix use in GEBP?
-Unlike $A_2$, which is thread private,
+What about $B_2$, the other packed matrix used in GEBP?
+Unlike $A_2$, which is thread-private,
 $B_2$ is shared among all threads and
 the packing of $B_2$ is done collaboratively by all threads.
-We have two choices, privatize $B_2$ and apply the set-partitioning method,
-or fall back to conventional way-partitioning for $B_2$.
-$B_2$ can be made thread private if every thread makes
+There are two choices (with different tradeoffs), (1) privatize $B_2$ and apply 
+set-partitioning and (2) 
+fall back to the conventional way-partitioning 
+for $B_2$.
+$B_2$ can be made thread-private if every thread makes
 its own pack of the whole $B_1$ matrix.
@@
 The second choice avoids the extra overhead in both time and space
-but risks inter-thread conflicts caused by the shared $B_2$ matrix.
-The two choices are evaluated and compared in Section~\ref{sec:evaluation}.
+but potentially increases the inter-thread conflicts caused by the shared $B_2$ matrix.
+In this work,
+these two choices are evaluated and compared in Section~\ref{sec:evaluation}.
 
-\subsection{Formal Description}\label{subsec:formal}
-Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
+\subsection{Algorithms}\label{subsec:formal}
+
+Given the tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
 and architectural details of the memory hierarchy,
-The SCP method systematically determines the
-the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
-Memory layout of other packed matrices on lower layers are
-determined by $A_2$ and $B_2$.
+SCP will systematically determine the
+the memory layout for the
+packed matrices $A_2$ and $B_2$ used in GEMM.
+The memory layout of the other packed matrices at
+lower layers are determined by the memory layout of
+$A_2$ and $B_2$.
 We assume that the share matrix $B_2$ is packed
-in conventional way-partitioning style.
+in the conventional way-partitioning style.
 If desired, $B_2$ can be privatized and handled in the same way as $A_2$.
 
-Before diving into details, we make a few assumptions,
-which are standard practice rarely violated,
-on the architectures that SCP can deal with:
+we make a few standard assumptions, which usually
+hold for the architecture considered by SCP:
 \begin{itemize}
-\item The architecture has 2 or 3 levels of cache. %% i.e., $NL=2$ or $NL=3$
-\item All caches are inclusive set-associative caches. %% i.e., $n_l \ge 2$ for $l \in [1, NL]$
-\item Caches on the same level are homogeneous. %% i.e., $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
-\item A cache's way-capacity is a multiple of the sum of way-capacity of its children,
+\item The architecture has 2 or 3 levels of cache; %% i.e. $NL=2$ or $NL=3$
+\item All caches are inclusive set-associative caches; %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
+\item Caches at the same level are homogeneous; and %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
+\item A cache's way-capacity is a multiple of the sum of way-capacities of its children,
 i.e., $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
@@
 We use a memory descriptor $\mathcal{D}_t^M$ to specify 
-the memory space occupied by packed matrix $M$ used by thread $t$.
+the memory space occupied by a 
+packed matrix $M$ used by thread $t$.
 The memory descriptor is essentially a subspace of
@@
 %% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
-A segment is represented by its start and end addresses $S_i = [s_i, e_i)$,
+A segment is represented by its start and end addresses,
+$S_i = [s_i, e_i)$,
 in which $s_i < e_i \in \mathcal{A}$.
 With way-partitioning, the memory descriptor contains only one segment.
-With set-partitioning used in previous section,
+With set-partitioning,
 the memory descriptor contains a sequence of equally strided segments.
@@
 instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
-for storing private and shared matrices respectively.
+for storing private and shared matrices, respectively.
 The second phase computes memory descriptors for $A_2$ instances
 by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
-$\mathcal{D}_t^{A_2}$ , $t \in [0, NT)$.
+$\mathcal{D}_t^{A_2}$ , where $t \in [0, NT)$.
 Buffers of different threads are disjoint,
-i.e., $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
+i.e., $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \emptyset$ if $i \ne j$.
 The third phase computes memory descriptors for $B_2$.
@@
 but the part (of size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
-The working flow of these three phases is listed
-in Algorithm~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.
+The work flow of these three phases is listed
+in Algorithms~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.
 
 In Algorithm~\ref{alg:scp.phase1},
-the per-thread size for memory buffer $\mathcal{D}_p$ and
-$\mathcal{D}_s$ are first computed
-(line~\ref{line:size.p}--\ref{line:size.s}).
+the per-thread sizes for memory buffers $\mathcal{D}_p$ and
+$\mathcal{D}_s$ are computed
+(lines~\ref{line:size.p} -- \ref{line:size.s}).
 Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
 extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
-Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
+Then $align$ is found
+(lines~\ref{line:align.init} -- \ref{line:align.endfor}) 
 and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
+Specifically,
 $align$ is initialized with the L1 cache line size $l_1$.
-Then on each level $l$ (line~\ref{line:align.for})
+Then for each level $l$ (line~\ref{line:align.for})
 with shared non-LRU caches (line~\ref{line:align.type}),
 $align$ is updated by computing the least-common-multiple of
-original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
-The insight in the $wc_l/nt_l$ value is that
-each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
-in the set-partitioning of shared caches.
-At last, memory is allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
+the earlier $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
+The insight on using $wc_l/nt_l$ is that
+each way of the level-$l$ cache should be
+equally partitioned among the $nt_l$ threads
+in set-partitioning.
+Finally, memory is allocated (lines~\ref{line:alloc.begin} -- \ref{line:alloc.end})
 and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
-are created (line~\ref{line:d.begin}--\ref{line:d.end}).
+are created (line~\ref{line:d.begin} -- \ref{line:d.end}).
 
@@
 %% FIXME memory buffer? memory space? subspace?
-Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
-The main part is a recursive procedure $subspace$
-(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
+Algorithm~\ref{alg:scp.phase2} describes
+the second phase of SCP.
+The main component is a recursive procedure $subspace$
+(lines~\ref{line:subspace.begin} -- \ref{line:subspace.end}).
 The $subspace$ procedure traverses the memory hierarchy
@@
 a subspace of $\mathcal{D}_p$.
-$subspace$ has three input parameters, $l$ and $idx$ to
-identify the node on which it is running,
+This procedure
+has three input parameters, $l$ and $idx$ for
+identifying the node on which it is running,
 and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
 The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
-among all children of node $(l, idx)$.
-If $subspace$ encounters a level 0 node, i.e., a processor core,
+among all the children of node $(l, idx)$.
+If $subspace$ encounters a level 0 node, i.e.,
+a processor core,
 then $\mathcal{D}_l^{idx}$ is exactly the thread-private
-memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
-(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
-If this is not the case, the number of children is computed (line~\ref{line:nchildren})
+memory buffer $\mathcal{D}_{idx}$ for thread $idx$,
+causing $subspace$ to return early
+(lines~\ref{line:shortcut.begin} -- \ref{line:shortcut.end}).
+Otherwise, the number of children is computed (line~\ref{line:nchildren})
 and $subspace$ iterates over all the child nodes
-(line~\ref{line:for.begin}--\ref{line:for.end}).
-For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
+(lines~\ref{line:for.begin} -- \ref{line:for.end}).
+For each child node $cidx$ at layer $l-1$ (line~\ref{line:cidx}),
 its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
@@
 a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
-The if-else branches (line~\ref{line:if}--\ref{line:endif})
-show two distinct ways to compute $\mathcal{D}_{temp}$.
+The if-else branch (lines~\ref{line:if} -- \ref{line:endif})
+represents two distinct ways for computing
+$\mathcal{D}_{temp}$.
 If the node is main memory or a LRU cache (line~\ref{line:if}),
@@
 First, the range of $\mathcal{D}_s$ is obtained
-(line~\ref{line:lb}--\ref{line:ub}) and
+(lines~\ref{line:lb} -- \ref{line:ub}) and
 the capacity of $\mathcal{D}_s$ is equally
-divided among $NT$ threads (line~\ref{line:len}).
+divided among the $NT$ threads (line~\ref{line:len}).
 Then $\mathcal{D}_s$ is partitioned in a way-partitioning style

### 2018-05-26 background.tex

@@
 \end{eqnarray}
-Here,
-$es$ denotes the size of a matrix element, e.g.,
-8B for a double precision floating-point number,
-and $c_l$ denotes the capacity of a register or cache 
-at level $l$.
+Here, $es$ denotes the size of a matrix element,
+e.g., 8B for a double precision floating-point number,
+and $c_l$ denotes the capacity of a register or cache at level $l$.
 By (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are so
 constrained that $M_r$ elements from $A$,
-$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into the registers
-(the pseudo level-0 cache).
-By (\ref{eq:constraints.l1}), $K_c$ is so
-constrained that $B_4$ ($N_r \times K_c$),
+$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can
+fit into the registers (the pseudo level-0 cache).
+By (\ref{eq:constraints.l1}), $K_c$ is so constrained that $B_4$ ($N_r \times K_c$),
 $A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
 at layer 5 fit into the L1 cache.
-By (\ref{eq:constraints.l2}), $M_c$ is so
-constrained that $A_3$ ($M_c \times K_c$),
+By (\ref{eq:constraints.l2}), $M_c$ is so constrained that $A_3$ ($M_c \times K_c$),
 $B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
 at layer 4 fit into the L2 cache.
-Finally, by (\ref{eq:constraints.l3}), $N_c$ is so
-constrained that 
+Finally, by (\ref{eq:constraints.l3}), $N_c$ is so constrained that 
 $B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache.
@@
 and share the same $B_1$.
-As a result, there are multiple $A_2$ instances
-(one per thread)
+
+As a result, there are multiple $A_2$ instances (one per thread)
 and a single $B_2$ instance (shared by all threads).
 To understand how the loop at layer 3 is parallelized on the 4-core processor, 
-Figure~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
+Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
 While all color shaded submatrices are accessed by $T_1$,
@@
 are packed by the thread.
-In other words, each thread packs its own $A_2$ instance and  all the threads collaboratively pack the 
-unique
-$B_2$ instance.
+In other words, each thread packs its own $A_2$ instance and  all the threads collaboratively pack the unique $B_2$ instance.
 %% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
@@
 Then at some time $t_m$ ($t_s < t_m < t_e$),
-$T_1$ prefetches data at addr $addr_1$ to the L2 cache.
+
+$T_1$ prefetches data at address $addr_1$ to the L2 cache.
 If $\varphi_2(addr_1)=\varphi_2(addr_0)$, then the
 data at $addr_1$ and $addr_0$ will be fetched into the same cache set.
-As the shared L2 cache has a non-LRU policy, the
-data at $addr_0$ may suffer an eviction before it is actually read,
-i.e. a conflict miss occurs.
+As the shared L2 cache has a non-LRU policy,
+the data at $addr_0$ may suffer an eviction before it is actually read,
+i.e., a conflict miss occurs.
 

### 2018-05-26 evaluation.tex

@@
 Generally, DGEMM performance is presented in $flops$.
-%% floating-point operations per second ($flops$).
 As $NT$ varies in our evaluation,
@@
 $E_{avg} = flops / (NT \cdot \widehat{flops})$,
-where
-$\widehat{flops}$ stands for the theoretical performance peak of a single core.
+where $\widehat{flops}$ stands for the theoretical performance
+peak of a single core.
 %% \begin{equation*}
@@
 Each curve represents a parallelism configuration.
-Different subfigures are configured with different
-$NT$'s,
+
+Different subfigures are configured with different $NT$'s,
 and different curves in the same subfigure are
@@
 As there is only one configuration for $NT=64$ ($NC=16$ and $NT_C=4$)
-and no comparison is available, the
-results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.
+and no comparison is available,
+the results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.
 
@@
 to partition the cache if $NT_C=1$.
-Figure~\ref{fig:benefit} shows the results.
+
+Figure.~\ref{fig:benefit} shows the results.
 For comparison purposes, the results without SCP
@@
 As a result, the packing overhead grows in proportion to $NT$
+
 in a full privatization strategy, which is unacceptable
 if GEMM is highly parallelized, e.g., when $NT=64$.
-Partial privatization in Figure~\ref{fig:packb.partial}
+Partial privatization in Figure.~\ref{fig:packb.partial}
 offers an alternative to full privatization.
@@
 privatization strategy with $NT_C=4$ and $NC=4$.
+
 The results under other parallelism configurations are similar.
@@
 
-Our results demonstrate that
-privatizing the shared matrix $B_2$ can further
+Our results demonstrate that privatizing the shared matrix $B_2$ can further
 reduce inter-thread cache conflicts.
-However,
-the improvement in cache performance is small. 
-As
-$B_2$ only occupies a small portion of the shared L2 cache,
+However, the improvement in cache performance is small.
+As $B_2$ only occupies a small portion of the shared L2 cache,
 this marginal improvement is not sufficient to

### 2018-05-26 intro.tex

@@
 the kernel.
-In this paper, we focus on optimizing the memory 
-accesses for GEMM.
+In this paper, we focus on optimizing the memory accesses for GEMM.
 
@@
 
-However, software prefetching and data packing cannot be
-effectively
+However, software prefetching and data packing cannot be effectively
 applied to architectures with non-LRU shared caches.
-In traditional high-performance processors, e.g.,
-the Intel Xeon series,
+In traditional high-performance processors, e.g., the Intel Xeon series,
 both L1 and L2 caches are private to processor cores
@@
   and pseudo-random shared L2 caches.
-  Our evaluation shows that SCP can improve the
-  GEMM performance consistently
+  Our evaluation shows that SCP can improve the GEMM performance consistently
   under various parallelism configurations, by $2.75\%$--$6.91\%$.

### 2018-05-26 intro.tex
Fix two "FIXME"s in the draft
- Citation to BLAS
- List of AArch64 processors with shared L2 cache

@@
 scientific and engineering computing domains.
-Basic Linear Algebra Subprograms (BLAS) (FIXME) defines a collection
+Basic Linear Algebra Subprograms (BLAS)~\cite{blas}  defines a collection
 of APIs which act as standard building blocks for dense matrix operations.
@@
 Indeed, some high-performance processors based on the
-ARM technology,
-have already adopted this design. (FIXME list the processors)
+ARM technology, e.g., the Phytium processor series~\cite{phytium},
+%% I'm sure they have shared L2 caches,
+%% but have not found any technical specification about the replacement policy
+%% XGene from AppliedMicro~\cite{xgene},
+%% ThunderX from Cavium~\cite{thunderx}
+have already adopted this design. 
 As a result, developing a better solution to reduce

### 2018-05-26 background.tex
Fix a syntax error

@@
 The main memory, $NL$ levels of caches, and $NT$ processor cores
-form a memory hierarchy are generally organized in terms
+forming a memory hierarchy are generally organized in terms
 of an $NL+2$ level regular tree structure.

### 2018-05-26 background.tex
Change the loop name N, K, M to jj, kk, ii

@@
 and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
-Either the N loop at layer 1 or the M loop at layer 3, or 
-both,
+Either the $jj$ loop at layer 1 or the $ii$ loop at layer 3, or both,
 can be parallelized in a multi-threaded context.
 Specifically, the implementation in Figure~\ref{fig:gemm}
-chooses an NKM loop order and parallelizes the M loop at layer 3.
+chooses an $jj\textrm{-}kk\textrm{-}ii$ loop order
+and parallelizes the $ii$ loop at layer 3.
 Different threads work on different parts of $A_1$
@@
 and a single $B_2$ instance (shared by all threads).
-To understand how the loop at layer 3 is parallelized on the 4-core processor, 
-Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
+To understand how the $ii$ loop at layer 3 is parallelized on the 4-core processor, 
+Figure.~\ref{fig:workload} shows the data accessing patterns
+of a single thread $T_1$.
 While all color shaded submatrices are accessed by $T_1$,

### 2018-05-26 background.tex
Improve syntax and fix formating errors

@@
 e.g., 8B for a double precision floating-point number,
-and $c_l$ denotes the capacity of a register or cache at level $l$.
+and $c_l$ denotes the capacity of the register file or cache at level $l$.
 By (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are so
@@
 in every iteration at layer 6 before it actually reads them in case of
-the data being evicted out accidentally.
+the data being evicted accidentally.
 As GEBP accesses the matrix data much more frequently than other
@@
 Then at some time $t_m$ ($t_s < t_m < t_e$),
-
 $T_1$ prefetches data at address $addr_1$ to the L2 cache.

### 2018-05-26 evaluation.tex
Remove some comments
Improve syntax and fix formating errors

@@
 Each curve represents a parallelism configuration.
-
 Different subfigures are configured with different $NT$'s,
@@
     \toprule
-     & $NT\!\!=\!\!4$ & $NT\!\!=\!\!8$ & $NT\!\!=\!\!16$ & $NT\!\!=\!\!32$ & $NT\!\!=\!\!64$ \\
+     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
     \midrule
-    $NT_C\!\!=\!\!1$ & 91.46 & 91.30 & 90.30 & -     & - \\   
-    $NT_C\!\!=\!\!2$ & 90.12 & 90.25 & 89.49 & 88.26 & - \\
-    $NT_C\!\!=\!\!4$ & 84.05 & 80.61 & 79.74 & 79.81 & 71.71 \\
+    $NT_C=1$ & 91.46 & 91.30 & 90.30 & -     & - \\   
+    $NT_C=2$ & 90.12 & 90.25 & 89.49 & 88.26 & - \\
+    $NT_C=4$ & 84.05 & 80.61 & 79.74 & 79.81 & 71.71 \\
+    %%  & $NT\!\!=\!\!4$ & $NT\!\!=\!\!8$ & $NT\!\!=\!\!16$ & $NT\!\!=\!\!32$ & $NT\!\!=\!\!64$ \\
+    %% \midrule
+    %% $NT_C\!\!=\!\!1$ & 91.46 & 91.30 & 90.30 & -     & - \\   
+    %% $NT_C\!\!=\!\!2$ & 90.12 & 90.25 & 89.49 & 88.26 & - \\
+    %% $NT_C\!\!=\!\!4$ & 84.05 & 80.61 & 79.74 & 79.81 & 71.71 \\
     \bottomrule
@@
 to partition the cache if $NT_C=1$.
-
 Figure.~\ref{fig:benefit} shows the results.
@@
 as SCP achieves the largest improvement with full cache sharing.
-Table~\ref{tab:papi} summarizes the average of the
-cache miss rates
-across the matrix sizes used.
+Table~\ref{tab:papi} summarizes the average reduction of the
+cache miss rates across the matrix sizes used.
 Under all the parallelism configurations,
@@
 The shared $B_2$ matrix is still packed
-in a conventional way-partitioning style.
+in the conventional way-partitioning style.
 As mentioned in Section~\ref{subsec:example},
@@
 \begin{figure}
-  %% FIXME draw this figure
   \centering
@@
 As a result, the packing overhead grows in proportion to $NT$
-
 in a full privatization strategy, which is unacceptable
@@
 privatization strategy with $NT_C=4$ and $NC=4$.
-
 The results under other parallelism configurations are similar.

### 2018-05-26 related.tex
Change the loop name N, K, M to jj, kk, ii
Improve syntax and fix formating errors

@@
 In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
-the kernel component of GEMM is 
-written by domain experts in assembly.
-In the case of auto-tuning,
+the kernel component of GEMM is written by domain experts in assembly.
 ATLAS~\cite{atlas} adopts the auto-tuning method to
@@
 address the problem of cache sharing with non-LRU replacement policies,
-and SCP serves as the first work to address this problem.
+and SCP serves as the first work to solve this problem.
 In \cite{gotogemm}, a detailed discussion on
-choosing
-different loop orders is given.
-GEMM are usually parallelized only at
-layer 3 (the M loop),
+choosing different loop orders is given.
+GEMM are usually parallelized only at layer 3 (the $ii$ loop),
 but all loops along the M and N dimensions are potentially parallelizable.
 BLIS\cite{blispar} allows developers to specify a sophisticated configuration
-so 
-that more than one loops at layers 1, 3, 4 and 5
+so that more than one loops at layers 1, 3, 4 and 5
 can be simultaneously parallelized.

### 2018-05-26 scp.tex
Remove some comments
Improve syntax and fix formating errors

@@
 Many solutions for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}), including
-$(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. Auto
-tuning returns
-$M_r = 4$ and $N_r = 8$ as the best.
+$(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$.
+Auto tuning returns $M_r = 4$ and $N_r = 8$ as the best.
 By (\ref{eq:constraints.l1}),
@@
 Each $A_2$ is packed into a single continuous buffer.
-Figure~\ref{fig:partition.conventional} shows how the
+Figure.~\ref{fig:partition.conventional} shows how the
 $A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
@@
 The distance from one segment to the next is equal to $wc_2=128KB$.
-%% There is no memory fragment because the segment size is a multiple of
-%% size of $A_4$ (8KB).
-%% FIXME put this to the end of this section
 While avoiding inter-thread cache conflicts,
@@
 set-partitioning and (2) 
-fall back to the conventional way-partitioning 
-for $B_2$.
+fall back to the conventional way-partitioning for $B_2$.
 $B_2$ can be made thread-private if every thread makes
@@
 The second choice avoids the extra overhead in both time and space
-but potentially increases the inter-thread conflicts caused by the shared $B_2$ matrix.
-In this work,
-these two choices are evaluated and compared in Section~\ref{sec:evaluation}.
+but potentially increases the inter-thread conflicts caused
+by the shared $B_2$ matrix.
+In this work, these two choices are evaluated
+and compared in Section~\ref{sec:evaluation}.
 
@@
 SCP will systematically determine the
-the memory layout for the
-packed matrices $A_2$ and $B_2$ used in GEMM.
+the memory layout for the packed matrices $A_2$ and $B_2$ used in GEMM.
 The memory layout of the other packed matrices at
-lower layers are determined by the memory layout of
-$A_2$ and $B_2$.
+lower layers are determined once the memory layout of
+$A_2$ and $B_2$ are determined.
 We assume that the share matrix $B_2$ is packed
@@
 $\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
-%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
-A segment is represented by its start and end addresses,
-$S_i = [s_i, e_i)$,
+A segment is represented by its start and end addresses, $S_i = [s_i, e_i)$,
 in which $s_i < e_i \in \mathcal{A}$.
@@
 
-%% FIXME use l as the level iterator?
 \begin{algorithm}
@@
 and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
-Specifically,
-$align$ is initialized with the L1 cache line size $l_1$.
+Specifically, $align$ is initialized with the L1 cache line size $l_1$.
 Then for each level $l$ (line~\ref{line:align.for})
@@
 
-%% FIXME memory buffer? memory space? subspace?
-Algorithm~\ref{alg:scp.phase2} describes
-the second phase of SCP.
+Algorithm~\ref{alg:scp.phase2} describes the second phase of SCP.
 The main component is a recursive procedure $subspace$
@@
 among all the children of node $(l, idx)$.
-If $subspace$ encounters a level 0 node, i.e.,
-a processor core,
+If $subspace$ encounters a level 0 node, i.e., a processor core,
 then $\mathcal{D}_l^{idx}$ is exactly the thread-private

### 2018-05-26 background.tex
fix syntax error

@@
 The main memory, $NL$ levels of caches, and $NT$ processor cores
-forming a memory hierarchy are generally organized in terms
+form a memory hierarchy which are generally organized in terms
 of an $NL+2$ level regular tree structure.

### 2018-05-26 background.tex
make consistent use of "ijk" loop names and "MNK" matrix dimensions

@@
 
-In this blocked algorithm, the loops over the $N$, $K$ and $M$
-matrix dimensions are tiled by sizes $N_c$, $K_c$ and $M_c$ at 
-layers 1, 2 and 3, respectively.
-At layers 4 and 5, the $N$ and $M$ dimensions are further tiled by sizes
-$N_r$ and $M_r$, respectively.
+In this blocked algorithm, the $jj$, $kk$ and $ii$ loops
+along the $N$, $K$ and $M$ matrix dimensions are
+tiled by sizes $N_c$, $K_c$ and $M_c$ at layers 1, 2 and 3, respectively.
+At layers 4 and 5, the $j$ and $i$ loops along the $N$ and $M$ dimensions
+are further tiled by sizes $N_r$ and $M_r$, respectively.
 As a result, the innermost loop at layer 6 goes over the $K$

### 2018-05-27 abstract.tex
Final revision from Prof. Xue

@@
   Two techniques, software prefetching and data packing,
-  have been used to effectively exploit the capability of on-chip lastly-recent-used (LRU) caches,
+  have been used to effectively exploit the capability of
+  on-chip lastly-recent-used (LRU) caches,
   which are popular in traditional high-performance
@@
   We present a Shared Cache Partitioning (SCP) method to
-  eliminate inter-thread cache conflicts in the
-  GEMM routines,
+  eliminate inter-thread cache conflicts in the GEMM routines,
   by partitioning a shared cache into physically disjoint
@@
   a 64-core AArch64 processor with private LRU L1 caches
-  and shared pseudo-random L2 caches (per every four cores).
-  Our experimental results show that conflict misses in both L1 and L2
-  caches are effectively reduced, resulting in an improvement
-  of the GEMM performance by $2.75\%$--$6.91\%$.
+  and shared pseudo-random L2 caches (per four-core cluster).
+  Our evaluation shows that SCP has effectively reduced the
+  conflict misses in both L1 and L2 caches,
+  resulting in an improvement of the GEMM performance by $2.75\%$ -- $6.91\%$.
 \end{abstract}

### 2018-05-27 background.tex
Final revision from Prof. Xue

@@
 A set-associative cache can be defined by a 4-tuple $(c, l, n, nt)$,
-in which $c$ is the cache capacity, $l$ is the cache line size,
-$n$ is ways-of-associativity,
+where $c$ is the cache capacity, $l$ is the cache line size,
+$n$ is the number of ways-of-associativity,
 and $nt$ indicates the number of cores sharing the cache.
@@
 An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
-mapping addresses from the address space $\mathcal{A}$
+mapping the addresses in the address space $\mathcal{A}$
 (either virtual or physical) to the cache sets in
@@
 \end{equation*}
-It is an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$
+which represents an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$-th
 position in $addr$'s binary form.
-
 It is easy to prove that any continuous memory block
@@
 cannot conflict with itself.
-We define a quantity, way-capacity, denoted as $wc=c/n$,
+We therefore define a quantity, \emph{way-capacity}, denoted as $wc=c/n$,
 to represent this size bound for conflict-free memory blocks.
@@
 The main memory, $NL$ levels of caches, and $NT$ processor cores
-form a memory hierarchy which are generally organized in terms
+form a memory hierarchy are generally organized in terms
 of an $NL+2$ level regular tree structure.
@@
 Each core has 32 128-bit (16B) vector registers,
-each capable of storing 2 double precision floating-point numbers.
+each capable of storing two double-precision floating-point numbers.
 For convenience, a processor core at level 0
 is represented by its registers, which can be viewed as a special
-level-0 cache $L0 = (512B, 16B, 32, 1)$ with programmable replacement policy.
+level-0 cache $L0 = (512B, 16B, 32, 1)$ with a programmable replacement policy.
 
@@
 sophisticated due to the presence of multi-level memory
-hierarchy on modern processors.
-Figure~\ref{fig:gemm} shows the structure of GEMM program from
+hierarchies on modern processors.
+Figure~\ref{fig:gemm} shows the structure of GEMM from
 the OpenBLAS library~\cite{openblas}.
@@
 Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
-matrices on each level fit into a certain level in the memory hierarchy.
-Equations~(\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3}) express
-the constraints on these tiling factors:
+matrices at each level fit into a certain level in the memory hierarchy, with the following
+constraints:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Here, $es$ denotes the size of a matrix element,
-e.g., 8B for a double precision floating-point number,
+where $es$ denotes the size of a matrix element,
+e.g., 8B for a double-precision floating-point number
 and $c_l$ denotes the capacity of the register file or cache at level $l$.
@@
 Finally, by (\ref{eq:constraints.l3}), $N_c$ is so constrained that 
-$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache.
+$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache (if it exists).
 Larger matrices at higher layers 1 -- 3 always reside in main memory.
@@
 implement GEMM programs based on this factorization,
-with GEBP highly optimized for the target processor (often coded in assembly).
+with GEBP highly optimized for the target processor.
 
@@
 In Figure~\ref{fig:gemm}, GEBP serves as the kernel routine,
-and the overall workload-partitioning strategy is defined by the three outer loops at layers 1 -- 3.
+and the workload-partitioning strategy is defined by the three outer loops at layers 1 -- 3.
 Developers can decide how to partition the workload by choosing
 tile factors $M_c$, $N_c$ and $K_c$,
-and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
+and how to schedule the GEBP tasks by interchanging loop orders at layers 1 - 3.
 Either the $jj$ loop at layer 1 or the $ii$ loop at layer 3, or both,
@@
 and parallelizes the $ii$ loop at layer 3.
-Different threads work on different parts of $A_1$
-and share the same $B_1$.
+Different threads work on different portions of $A_1$
+but share the same $B_1$.
 
@@
 To understand how the $ii$ loop at layer 3 is parallelized on the 4-core processor, 
-Figure.~\ref{fig:workload} shows the data accessing patterns
-of a single thread $T_1$.
-While all color shaded submatrices are accessed by $T_1$,
-only the z-curve masked submatrices (a subset of color shaded ones)
-are packed by the thread.
-In other words, each thread packs its own $A_2$ instance and  all the threads collaboratively pack the unique $B_2$ instance.
+Figure~\ref{fig:workload} shows the data accessing patterns
+for a single thread $T_1$.
+While all the color shaded submatrices are accessed by $T_1$,
+only the submatrices highlighted by the z-curves
+(a subset of color shaded ones) are packed by $T_1$.
+In other words, each thread packs its own $A_2$ instance but all the threads pack collaboratively the unique $B_2$ instance.
 %% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
@@
 into capacity and conflict misses.
-While the constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
+While (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
 aim to avoid capacity misses,
@@
 and is expected to reside in the 
-L2 cache during all $N_c/N_r$ iterations at layer 4.
+L2 cache during all the $N_c/N_r$ iterations at layer 4.
 The packed matrix $B_4$ is similar in that it is expected to
-reside in the L1 cache during all $M_c/M_r$ iterations at layer 5.
+reside in the L1 cache during all the
+$M_c/M_r$ iterations at layer~5.
 In general, even with carefully selected tile factors,
-matrix data can still be evicted.
+some matrix elements can still be evicted.
 For example, some data structures other than the packed matrices
 or program code (in case of a unified cache) may be fetched into the
-cache set where the matrix data reside in.
-Such cache conflicts are referred to as
-intra-thread cache conflicts
-because it occurs within the execution of a single thread.
-To deal with intra-thread conflicts,
-the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to the L1 cache
-in every iteration at layer 6 before it actually reads them in case of
-the data being evicted accidentally.
+cache set where these matrix elements reside in.
+Such cache conflicts are known as
+\emph{intra-thread cache conflicts}
+because they occur within the execution of a single thread.
+To deal with intra-thread cache conflicts,
+the GEBP kernel repeatedly prefetches data from $A_4$ and $B_4$ to the L1 cache
+during every iteration at layer 6 before it actually reads them just in case that
+the data may have been evicted accidentally from the L1 cache.
 As GEBP accesses the matrix data much more frequently than other
 data structures, this repeated prefetching strategy
-can reduce the penalty caused by intra-thread conflicts.
+can reduce the penalty caused by intra-thread cacheconflicts.
 
@@
 there exists another kind of cache conflicts, called
-inter-thread cache conflicts, on architectures with shared caches.
+\emph{inter-thread cache conflicts}, on architectures with shared caches.
 Suppose two working threads, $T_0$ and $T_1$, are running independently
@@
 At time $t_s$, $T_0$ prefetches some data at address $addr_0$
-to cache set $\varphi_2(addr_0)$ of the shared L2 cache.
+to the cache set $\varphi_2(addr_0)$ in a shared L2 cache.
 The data will be actually read at time $t_e$.
@@
 So inter-thread cache conflicts described above
-can occur frequently on non-LRU shared caches,
+can occur frequently for shared non-LRU caches,
 thus making the technique dealing with intra-thread cache conflicts ineffective.

### 2018-05-27 conclusion.tex
Final revision from Prof. Xue

@@
 In this paper, we present a shared cache partitioning method,
-SCP, to reduce inter-thread conflicts in GEMM programs
+SCP, to reduce inter-thread cache conflicts in GEMM programs
 on architectures with shared non-LRU caches.
@@
 disjoint sets and assign different sets to different threads.
-Evaluation results show that SCP can effectively reduce
-cache misses for a shared cache (and its
-upper level caches),
-resulting in a considerable improvement in GEMM performance
+Our experimental results show that SCP can effectively reduce the inter-thread
+cache misses for a shared cache (and its upper level caches),
+resulting in a considerable improvement in the GEMM performance
 in a multi-threaded context.

### 2018-05-27 evaluation.tex
Final revision from Prof. Xue

@@
 computed as in Section~\ref{subsec:example}:
-$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
-The parallelism is controlled by two parameters,
+$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$,
+as this combination causes GEMM to deliver the best performance 
+when SCP is not used.
+The degree of parallelism is controlled by two parameters,
 the number of active clusters $NC$
 and the number of active threads per cluster $NT_C$.
-The total number of threads is $NT = NC \cdot NT_C$.
+The total number of threads is $NT = NC \times NT_C$.
 For the Phytium 2000+ processor, we have
 $NC \in [1, 16]$ and $NT_C \in [1, 4]$. 
-Generally, DGEMM performance is presented in $flops$.
+Generally, the DGEMM performance is measured in $flops$.
 As $NT$ varies in our evaluation,
-we use another metric derived from $flops$,
+we also  use another metric derived from $flops$,
 the average thread efficiency,
@@
 The average thread efficiency is a normalized value computed as
-$E_{avg} = flops / (NT \cdot \widehat{flops})$,
+$E_{avg} = flops / (NT \times cdot \widehat{flops})$,
 where $\widehat{flops}$ stands for the theoretical performance
@@
 (Section~\ref{subsec:analysis}), and
-discuss the privatization of shared matrix $B_2$
+discuss the privatization of the shared matrix $B_2$
 (Section~\ref{subsec:privb}).
@@
 
-As the L2 cache is shared by cores inside a cluster,
-the impact of cache sharing can be demonstrated
+In a cluster, the L2 cache is shared by its four cores.
+The impact of cache sharing can be demonstrated
 by comparing the DGEMM performance results achieved
@@
 Figure~\ref{fig:drawback} shows the results.
-Each curve represents a parallelism configuration.
+Each curve represents a parallelization configuration.
 Different subfigures are configured with different $NT$'s,
@@
 configured with different $NT_C$'s.
-As there is only one configuration for $NT=64$ ($NC=16$ and $NT_C=4$)
-and no comparison is available,
+As there is only one configuration for $NT=64$,
+($NC=16$ and $NT_C=4$),
+no comparison is available. Thus,
 the results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.
@@
 
-We can see that for all $NT$'s, $NT_C=1$ achieves the 
+For all $NT$'s, $NT_C=1$ achieves the 
 best performance, followed by 
@@
 This shows clearly that cache sharing has a considerable
-impact on DGEMM performance.
+impact on the DGEMM performance.
 
@@
 
-Under all the parallelism configurations, 
+Under all the parallelization configurations, 
 SCP performs better than Base
-consistently over all the matrix sizes evaluated.
+consistently across all the matrix sizes evaluated.
 Table~\ref{tab:win} summarizes the performance improvement of SCP over Base.
@@
 The largest performance gain, $6.91\%$, which is a considerable improvement,
-is observed under the maximum parallelism $NT=64$.
+is observed under the maximum parallelism with $NT=64$.
 %% Looking at Table~\ref{tab:benefit} vertically,
@@
 why SCP is effective in reducing
-inter-thread conflicts, 
+inter-thread cache conflicts, 
 we use the PAPI profiling tool~\cite{papi} to
@@
 $\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
-The hardware events are measured in a per-thread manner
+In this analysis,
+the hardware events are measured in a per-thread manner
 and all the results are the average measurements
@@
 
-Figure~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.  We only give the results with $NT_C=4$
+Figure~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.
+We only give the results with $NT_C=4$
 as SCP achieves the largest improvement with full cache sharing.
-Table~\ref{tab:papi} summarizes the average reduction of the
-cache miss rates across the matrix sizes used.
-Under all the parallelism configurations,
-both L1 and L2 cache miss rates are reduced under
-SCP, 
+Table~\ref{tab:papi} summarizes their
+average reductions across the matrix sizes used.
+Under all the parallelization configurations,
+both L1 and L2 cache miss rates are reduced under SCP, 
 by roughly $13\%$ for L1 and nearly $10\%$ for L2.
@@
 only L2 caches are specially handled by SCP.
-The miss rates for both L1 and L2 caches
+However,
+the miss rates for both L1 and L2 caches
 are reduced effectively by SCP.
@@
 the data to be evicted from L2,
-but also invalidate the evicted data's copy in all upper level L1 caches,
+but also invalidate the evicted data's copy in all its
+upper level L1 caches,
 resulting in more L1 cache misses.
-In general, conflict misses on an inclusive cache can affect
-the performance of all upper level caches. As a 
-result, improving the L2 miss rate also improves the
-L1 miss rate.
+In general, the conflict misses for
+an inclusive cache can affect
+the performance of all its upper level caches.
+As a result, improving the miss rate of a shared L2 cache
+also improves the miss rates of all its upper level L1 caches.
 
@@
 In addition, the performance benefit thus obtained
-also propagates to the upper level caches, too.
+also propagates to their upper level caches, too.
 
@@
 \subsection{Privatization of Shared Matrix}\label{subsec:privb}
-In the previous evaluations, SCP only set-partitions the shared cache
+In the previous evaluations, SCP only set-partitions the shared caches
 for $A_2$ matrices, which are private to threads.
@@
 set-partitioning can also be applied to $B_2$ after $B_2$ is privatized.
-In this section, we will discuss and evaluate this privatization alternative.
+In this section, we discuss and evaluate this privatization alternative.
 
@@
 Privatization incurs more packing overhead because
-redundant packing of $B_2$ is enforced.
+the packing of $B_2$ must be done redundantly.
 Figure~\ref{fig:packb} shows three packing strategies for $B_2$
-in a parallelism configuration specified by
+in a parallelization configuration specified by
 $NT_C=4$ and $NC=4$.
-The whole packing workload of $B_2$ is divided into 16 tasks
-and the packing workload of thread $T_1$ is masked by z-curves. 
+In each case, the entire packing workload of $B_2$ is divided into 16 tasks.
+So only the packing workload in one thread, $T_1$, 
+which is highlighted by the z-curves (as shown), is illustrated.
 %% and the task assignment is annoted in Figure~\ref{fig:packb}.
-In the conventional packing approach, each thread takes a single task.
-In full privatization, all tasks are replicated to all threads.
-As a result, the packing overhead grows in proportion to $NT$
-in a full privatization strategy, which is unacceptable
+In the conventional packing approach
+(Figure~\ref{fig:packb.conventional}),
+each thread takes a single task.
+In full privatization
+(Figure~\ref{fig:packb.full}),
+all tasks are replicated to all
+the threads.
+As a result, the packing overhead grows in proportion to $NT$, which is unacceptable
 if GEMM is highly parallelized, e.g., when $NT=64$.
-Partial privatization in Figure.~\ref{fig:packb.partial}
-offers an alternative to full privatization.
-In partial privatization, privatization only occurs inside
+In partial privatization 
+(Figure~\ref{fig:packb.partial}), which is 
+an alternative to full privatization,
+privatization only occurs inside
 a cluster so that the tasks for one cluster need not to be
-replicated to other clusters.
-Consequently,  the extra packing overhead of the partial privatization
-strategy is bounded by $NT_C$.
-For example, on Phytium 2000+, overhead of packing $B_2$
+replicated in the other clusters.
+In this case, the extra packing overhead
+is bounded by $NT_C$.
+For example, on Phytium 2000+, the
+overhead of packing $B_2$
 in partial privatization is limited to 4 times
-of that in conventional packing.
+of that in the conventional packing approach.
 
-Figure~\ref{fig:privb} shows the results of the partial
-privatization strategy with $NT_C=4$ and $NC=4$.
-The results under other parallelism configurations are similar.
+Figure~\ref{fig:privb} shows the results for partial
+privatization with $NT_C=4$ and $NC=4$.
+The results for the other 
+configurations are similar.
 For comparison purposes,
-SCP with and without privatization are both presented
-(denoted as SCP-P and SCP in Figure~\ref{fig:privb}, respectively).
-Figure~\ref{fig:privb.papi} shows the cache miss rates.
-Privatization achieves slightly lower miss rates for
-both L1 and L2 caches.
-Figure~\ref{fig:privb.ate} shows the average thread efficiency.
-Despite its lower cache miss rates,
-privatization suffers a performance loss compared to SCP.
-Why is there a performance drop when cache miss rates 
-are reduced?
-The reason lies in the extra packing overhead introduced by privatization.
-Figure~\ref{fig:privb.breakdown} shows the occupancy
-ratios of the overall execution times for various overheads.
-Overheads in GEMM include
+SCP with and without privatization,
+denoted as SCP-P and SCP, respectively,
+are given in Figure~\ref{fig:privb}.
+Figure~\ref{fig:privb.papi} shows the L1 and L2
+cache miss rates, which are 
+slightly lower with privatization (SCP-P) than without
+(SCP). Figure~\ref{fig:privb.ate} shows the average thread efficiency.
+Despite its lower L1 and L2 cache miss rates,
+privatization (SCP-P) suffers a performance loss
+(relative to SCP) due to 
+the extra packing overhead introduced by privatization.
+Figure~\ref{fig:privb.breakdown} compares the
+three types of major overheads in percentage terms 
+over the total execution time, including
 (1) packing of $A_2$, (2) packing of $B_2$,
 and (3) synchronization.
-While SCP-P involves roughly the same overheads in
-packing $A_2$ and synchronization as SCP,
-it spends much more time in packing $B_2$.
-Table~\ref{tab:breakdown} gives a breakdown of the
-execution time, indicating that the overhead for packing $B_2$
+While behaving similarly as SCP
+in (1) and (3), SCP-P is worse than SCP in (2).
+To understand this further,
+Table~\ref{tab:breakdown} gives a breakdown of each's
+execution time in terms of (1) -- (3) and the
+computation time of GEBP. As
+the overhead for packing $B_2$
 is multiplied by a factor of $NT_C=4$,
-and the total overhead increases from $2.35\%$ to $4.72\%$.
+the total overhead has thus increased
+from $2.35\%$ (under SCP) to $4.72\%$ (under SCP-P).
 
@@
   }
-  \subfigure[Executing time breakdown]{
+  \subfigure[Execution time breakdown]{
     \label{fig:privb.breakdown}
@@
   \centering
-  \caption{Executing time breakdown with $NT_C=4$ and $NC=4$}
+  \caption{Execution time breakdown with $NT_C=4$ and $NC=4$}
   \label{tab:breakdown}

### 2018-05-27 intro.tex
Final revision from Prof. Xue

@@
 scientific and engineering computing domains.
-Basic Linear Algebra Subprograms (BLAS)~\cite{blas}  defines a collection
+Basic Linear Algebra Subprograms (BLAS)~\cite{blas}  
+define a collection
 of APIs which act as standard building blocks for dense matrix operations.
-BLAS routines are divided into 3 levels,
+BLAS routines are divided into three levels,
 level-1 for vector-vector operations,
@@
 that are highly optimized for their processors,
-such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
+such as Intel MKL, AMD ACML, and NVIDIA cuBLAS.
 The HPC community has also contributed several high-quality
@@
 In a multi-threaded context, different tasks (loop tiles) are assigned to
-different threads to parallelize the whole operation.
-The main objective of GEMM optimization is to
+different threads to harness parallelism.
+The main objective of the GEMM optimization is to
 maximize the floating-point
@@
 before the kernel routine is called.
-Because GEMM is a computational intensive operation whose
-arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
-the overhead caused by data packing is negligible when
-the input matrices are large~\cite{gotogemm}.
+Because GEMM is a computational intensive operation 
+with an 
+arithmetic complexity of $O(n^3)$ and a memory complexity of $O(n^2)$,
+the overhead caused by data packing is negligible for
+large
+input matrices~\cite{gotogemm}.
 
 However, software prefetching and data packing cannot be effectively
-applied to architectures with non-LRU shared caches.
+applied to architectures with shared non-LRU caches.
 In traditional high-performance processors, e.g., the Intel Xeon series,
 both L1 and L2 caches are private to processor cores
-and LRU replacement policy is adopted.
+and an LRU replacement policy is adopted.
 Without considering the last level cache (LLC),
@@
 cannot cause conflict cache misses.
-This is no longer the case if caches are shared by several processor cores.
+This is no longer the case if a cache is
+shared by several processor cores.
 In this case, a cache line prefetched by one thread
@@
 so that the data used by several threads can simultaneously reside in a shared cache.
-Due to the nature of set-associative caches,
-this solution cannot completely eliminate inter-thread cache conflicts,
+Due to set-associativity,
+this solution is usually ineffective in reducing
+inter-thread cache conflicts,
 especially if the shared cache has a non-LRU replacement policy.
-Our evaluation shows that inter-thread cache conflicts
-in non-LRU shared caches
+Indeed,
+our evaluation shows that inter-thread cache conflicts
+in shared non-LRU caches
 can heavily hurt GEMM performance,
-even with packed matrices reduced to fit into the cache.
+even if packed matrices are reduced (in size)
+to fit into the shared cache.
 
@@
 is introducing more and more cores on a single processor chip.
-To reduce cost of on-chip cache memories and coherence networks,
+To reduce the cost of on-chip cache memories and coherence networks,
 shared caches may become common in future many-core processors.
@@
 the inter-thread cache conflicts 
-for non-LRU shared caches is important.
+for shared non-LRU caches is important.
 
 In this paper, we present a Shared Cache Partitioning (SCP) method
-to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
-The key idea is to partition a
-share cache into physically disjoint sets
+to reduce inter-thread cache conflicts on architectures with shared non-LRU caches.
+The key insight is to partition a
+shared cache into physically disjoint sets
 and assign different sets to different threads. This can be achieved by
-exploiting the memory mapping mechanism of set-associative caches.
+exploiting the memory mapping mechanism in a 
+set-associative cache.
 
@@
 the inter-thread cache conflicts problem on
-architectures with non-LRU share caches.
+architectures with shared non-LRU caches.
 
@@
 \item We present a quantitative analysis of the negative effect of inter-thread cache
-  conflicts on GEMM performance.
+  conflicts on the GEMM performance.
 \item We propose SCP, a method for solving the inter-thread cache conflicts
-  problem on architectures with non-LRU shared caches.
+  problem on architectures with shared non-LRU caches.
 \item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
@@
   Phytium 2000+ has 64 cores, private LRU L1 caches
-  and pseudo-random shared L2 caches.
+  and shared L2 caches with a pseudo-random
+  replacement policy.
   Our evaluation shows that SCP can improve the GEMM performance consistently
-  under various parallelism configurations, by $2.75\%$--$6.91\%$.
+  under various parallelization configurations, by $2.75\%$--$6.91\%$.
 \end{itemize}
@@
 the cache conflicts incurred.
-Section~\ref{sec:scp} introduces the SCP method.
+Section~\ref{sec:scp} introduces our SCP method.
 Section~\ref{sec:evaluation} presents and

### 2018-05-27 related.tex
make consistent use of "ijk" loop names and "MNK" matrix dimensions
Final revision from Prof. Xue

@@
 in~\cite{gemmbased1,gemmbased2}.
-The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
+The GotoBLAS library~\cite{gotoblas} and its successor OpenBLAS~\cite{openblas},
 are instantiated based on this insight.
@@
 and loop tiling~\cite{Lam1991,Spampinato:2014,Xue00}.
+Recently,
 \textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
@@
 the cache capacity and associativity
-but assume a LRU replacement policy.
+but assume an LRU replacement policy.
 As far as we know, this paper is the first to
@@
 GEMM are usually parallelized only at layer 3 (the $ii$ loop),
-but all loops along the M and N dimensions are potentially parallelizable.
-BLIS\cite{blispar} allows developers to specify a sophisticated configuration
+but all loops along the $M$ and $N$ matrix dimensions
+are potentially parallelizable.
+BLIS~\cite{blispar} allows developers to specify a sophisticated configuration
 so that more than one loops at layers 1, 3, 4 and 5

### 2018-05-27 scp.tex
Final revision from Prof. Xue

@@
 
-Suppose we want to implement a double precision GEMM (DGEMM) on
-the 4-core processor, where $es=8B$.
+Suppose we want to implement a double-precision GEMM (DGEMM) on
+our 4-core example processor, where $es=8B$.
 We now determine the tiling factors used based on
@@
 $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$.
-Auto tuning returns $M_r = 4$ and $N_r = 8$ as the best.
+A tuning process returns $M_r = 4$ and $N_r = 8$ as the best.
 By (\ref{eq:constraints.l1}),
@@
 To leave space for other data structures and the program code,
-it is reasonable to shrink $M_c$ slightly, with,
+it is reasonable to shrink $M_c$ slightly, leaving,
 for example, $192$, $208$ and $224$ as good
@@
 Here, we choose $N_c = 1024NT = 4096$.
-With these tiling factors determined, the sizes of packed matrices
+Given these tiling factors, the sizes of packed matrices
 can be calculated. Table.~\ref{tab:msizes} lists
@@
   }
-  \caption{Partition styles of the shared L2 cache (n=16, ns=2048)}
+  \caption{Partitioning styles for the shared L2 cache (n=16, ns=2048)}
   \label{fig:partition}
@@
 Figure.~\ref{fig:partition.conventional} shows how the
-$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
+$A_2$ instances of four threads are distributed in the shared 16-way L2 cache.
 Every cache set $s \in [0,2048)$ contains
-data from all 4 $A_2$ instances,
+data from all the four $A_2$ instances,
 potentially leading to inter-thread cache conflicts.
 What if the $A_2$ matrices are distributed in the way shown in
-Figure~\ref{fig:partition.segmented}?
-$A_2$ of different threads live in strictly disjoint cache sets, enabling
+Figure~\ref{fig:partition.segmented}? In this case,
+the different $A_2$ instances from different threads will
+reside in strictly disjoint cache sets, enabling
 the inter-thread cache conflicts caused by $A_2$ to be
@@
 Instead, $A_2$ will be distributed to 12 continuous memory segments, each of size $32KB$.
-The distance from one segment to the next is equal to $wc_2=128KB$.
+The distance from one segment to the next is $wc_2=128KB$.
 While avoiding inter-thread cache conflicts,
@@
 Unlike $A_2$, which is thread-private,
-$B_2$ is shared among all threads and
-the packing of $B_2$ is done collaboratively by all threads.
+$B_2$ is shared among all the threads and
+the packing of $B_2$ is done collaboratively by all
+the threads.
 There are two choices (with different tradeoffs), (1) privatize $B_2$ and apply 
 set-partitioning and (2) 
-fall back to the conventional way-partitioning for $B_2$.
+fall back to way-partitioning for $B_2$ (as done
+traditionally).
 $B_2$ can be made thread-private if every thread makes
@@
 The second choice avoids the extra overhead in both time and space
-but potentially increases the inter-thread conflicts caused
+but potentially increases the inter-thread cache
+conflicts caused
 by the shared $B_2$ matrix.
@@
 and architectural details of the memory hierarchy,
-SCP will systematically determine the
-the memory layout for the packed matrices $A_2$ and $B_2$ used in GEMM.
+SCP will systematically determine 
+the memory layout for the packed matrices $A_2$ and $B_2$ in GEMM.
 The memory layout of the other packed matrices at
 lower layers are determined once the memory layout of
-$A_2$ and $B_2$ are determined.
+$A_2$ and $B_2$ is determined.
 We assume that the share matrix $B_2$ is packed
@@
 
-we make a few standard assumptions, which usually
-hold for the architecture considered by SCP:
+We make a few standard assumptions, which usually
+hold for the architecture considered:
 \begin{itemize}
-\item The architecture has 2 or 3 levels of cache; %% i.e. $NL=2$ or $NL=3$
+\item The architecture has 2 or 3 levels of caches; %% i.e. $NL=2$ or $NL=3$
 \item All caches are inclusive set-associative caches; %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
-\item Caches at the same level are homogeneous; and %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
+\item The caches at the same level are homogeneous; and %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
 \item A cache's way-capacity is a multiple of the sum of way-capacities of its children,
@@
 the memory space occupied by a 
-packed matrix $M$ used by thread $t$.
+packed matrix $M$ used by a thread $t$.
 The memory descriptor is essentially a subspace of
@@
 A segment is represented by its start and end addresses, $S_i = [s_i, e_i)$,
-in which $s_i < e_i \in \mathcal{A}$.
+where $s_i < e_i \in \mathcal{A}$.
 With way-partitioning, the memory descriptor contains only one segment.
@@
 \begin{algorithm}
-  \caption{SCP phase 1: allocate memory buffers}
+  \caption{SCP (Phase 1): allocating memory buffers
+  for $A_2$ and $B_2$}
   \label{alg:scp.phase1}
@@
 
-SCP runs in three phases.
-The first phase allocates memory buffers for all $A_2$ and $B_2$
+SCP runs in three phases, as shown in 
+Algorithms~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3},
+respectively.  The first phase allocates the
+memory buffers for all $A_2$ and $B_2$
 instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
 for storing private and shared matrices, respectively.
-The second phase computes memory descriptors for $A_2$ instances
+The second phase computes the 
+memory descriptors for $A_2$ instances
 by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
 $\mathcal{D}_t^{A_2}$ , where $t \in [0, NT)$.
-Buffers of different threads are disjoint,
+The buffers of different threads are disjoint,
 i.e., $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \emptyset$ if $i \ne j$.
-The third phase computes memory descriptors for $B_2$.
-Because is $B_2$ is shared among all threads,
+The third phase computes the
+memory descriptors for $B_2$.
+Because $B_2$ is shared among all the threads,
 $\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
-but the part (of size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
-The work flow of these three phases is listed
-in Algorithms~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.
+but the part (of size $\frac{es N_c K_c}{NT}$),
+which is packed by thread $t$.
 
-In Algorithm~\ref{alg:scp.phase1},
+In Algorithm~\ref{alg:scp.phase1} (Phase 1),
 the per-thread sizes for memory buffers $\mathcal{D}_p$ and
@@
   \renewcommand{\algorithmicendwhile}{\textbf{end procedure}}
-  \caption{SCP phase 2: compute descriptors for $A_2$}
+  \caption{SCP (Phase 2): computing the memory descriptors for $A_2$}
   \label{alg:scp.phase2}
@@
 
-Algorithm~\ref{alg:scp.phase2} describes the second phase of SCP.
-The main component is a recursive procedure $subspace$
+In Algorithm~\ref{alg:scp.phase2} (Phase 2),
+the main component is a recursive procedure $subspace$
 (lines~\ref{line:subspace.begin} -- \ref{line:subspace.end}).
-The $subspace$ procedure traverses the memory hierarchy
+This procedure traverses the memory hierarchy
 to compute for each node, including the processor cores and main memory,
 a subspace of $\mathcal{D}_p$.
-This procedure
-has three input parameters, $l$ and $idx$ for
+There are three input parameters, $l$ and $idx$ for
 identifying the node on which it is running,
@@
 $\mathcal{D}_{temp}$.
-If the node is main memory or a LRU cache (line~\ref{line:if}),
+If the node is main memory or an LRU cache (line~\ref{line:if}),
 $\mathcal{D}_l^{idx}$ is partitioned in the conventional
@@
 a set-partitioning on $\mathcal{D}_l^{idx}$.
-Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
+Initially, the main memory is the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
 and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
-because the main memory is shared by all threads.
+because the main memory is shared by all the threads.
 Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
@@
 
-The computation of $\mathcal{D}_t^{B_2}$ is quite simple.
-Algorithm~\ref{alg:scp.phase3} divides $\mathcal{D}_s$
+In Algorithm~\ref{alg:scp.phase3} (Phase 3),
+the computation of $\mathcal{D}_t^{B_2}$ is quite 
+simple. This is done by dividing $\mathcal{D}_s$
 into $NT$ partitions, one for each thread.
@@
 Then $\mathcal{D}_s$ is partitioned in a way-partitioning style
-and each thread $t$ get buffer space for its part in $B_2$
-(line~\ref{line:thread.for}--\ref{line:thread.forend}).
+and each thread $t$ obtains the buffer space for its part in $B_2$
+(lines~\ref{line:thread.for} -- \ref{line:thread.forend}).
 
 \begin{algorithm}
-  \caption{SCP phase 3: compute descriptors for $B_2$}
+  \caption{SCP (Phase 3): computing the memory descriptors for $B_2$}
   \label{alg:scp.phase3}

### 2018-05-28 evaluation.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references

@@
 
-We have implemented SCP in the OpenBLAS~\cite{openblas} library (version 0.3.0-dev) and evaluated
-it on a Phytium 2000+ processor.
+We have implemented SCP in the OpenBLAS~\cite{openblas} library
+%% (version 0.3.0-dev)
+and evaluated it on a Phytium 2000+ processor.
 The Phytium 2000+ processor is an emerging high-performance
@@
 with hardware coherence.
+\textcolor{red}{
 Table~\ref{tab:cluster} lists the
 architectural parameters for the cores and caches
-in this processor.
+in this processor
+and Table~\ref{tab:software} shows the software environment used in out evaluation.}
 We restrict our evaluation to DGEMM,
@@
 relies on the DGEMM variant.
+
+\begin{table}
+  \centering
+  \caption{Software environment used in evaluation}
+  \label{tab:software}
+  \begin{tabular}{lll}
+    \toprule
+    Environment & Software & Version \\
+    \midrule
+    Operating System & GNU/Linux & 4.4.0 AArch64 \\
+    Compiler & GNU/GCC & 6.3.0 \\
+    Thread Model & OpenMP & 4.0 \\
+    BLAS & OpenBLAS & 0.3.0-dev \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
 As $NT$ varies in our evaluation,
-we also  use another metric derived from $flops$,
+we use another metric derived from $flops$,
 the average thread efficiency,
@@
 The average thread efficiency is a normalized value computed as
-$E_{avg} = flops / (NT \times cdot \widehat{flops})$,
+$E_{avg} = flops / (NT \times \widehat{flops})$,
 where $\widehat{flops}$ stands for the theoretical performance
@@
 
-\begin{figure}
-  \centering
-  \subfigure[$NT=4$]{
-    \label{fig:drawback-4}
-    \includegraphics[width=.31\textwidth]{figures/sec1-4}
-  }
-  \subfigure[$NT=8$]{
-    \label{fig:drawback-8}
-    \includegraphics[width=.31\textwidth]{figures/sec1-8}
-  }
-  \\
-  \subfigure[$NT=16$]{
-    \label{fig:drawback-16}
-    \includegraphics[width=.31\textwidth]{figures/sec1-16}
-  }
-  \subfigure[$NT=32,64$]{
-    \label{fig:drawback-32}
-    \includegraphics[width=.31\textwidth]{figures/sec1-32}
-  }
-  \caption{Average thread efficiency with cache sharing}
-  \label{fig:drawback}
-\end{figure}
-
 In a cluster, the L2 cache is shared by its four cores.
@@
 the results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.
+
+\begin{figure}
+  \centering
+  \subfigure[$NT=4$]{
+    \label{fig:drawback-4}
+    \includegraphics[width=.45\textwidth]{figures/sec1-4}
+  }
+  \subfigure[$NT=8$]{
+    \label{fig:drawback-8}
+    \includegraphics[width=.45\textwidth]{figures/sec1-8}
+  }
+  \\
+  \subfigure[$NT=16$]{
+    \label{fig:drawback-16}
+    \includegraphics[width=.45\textwidth]{figures/sec1-16}
+  }
+  \subfigure[$NT=32,64$]{
+    \label{fig:drawback-32}
+    \includegraphics[width=.45\textwidth]{figures/sec1-32}
+  }
+  \caption{Average thread efficiency with cache sharing}
+  \label{fig:drawback}
+\end{figure}
 
@@
     \label{fig:benefit-4}
-    \includegraphics[width=.31\textwidth]{figures/sec2-4}
+    \includegraphics[width=.45\textwidth]{figures/sec2-4}
   }
@@
     \label{fig:benefit-8}
-    \includegraphics[width=.31\textwidth]{figures/sec2-8}
+    \includegraphics[width=.45\textwidth]{figures/sec2-8}
   }
@@
     \label{fig:benefit-16}
-    \includegraphics[width=.31\textwidth]{figures/sec2-16}
+    \includegraphics[width=.45\textwidth]{figures/sec2-16}
   }
@@
     \label{fig:benefit-32}
-    \includegraphics[width=.31\textwidth]{figures/sec2-32}
+    \includegraphics[width=.45\textwidth]{figures/sec2-32}
   }
@@
     \label{fig:benefit-64}
-    \includegraphics[width=.31\textwidth]{figures/sec2-64}
+    \includegraphics[width=.45\textwidth]{figures/sec2-64}
   }
@@
     \label{fig:papi-4}
-    \includegraphics[width=.31\textwidth]{figures/papi-4}
+    \includegraphics[width=.45\textwidth]{figures/papi-4}
   }
@@
     \label{fig:papi-8}
-    \includegraphics[width=.31\textwidth]{figures/papi-8}
+    \includegraphics[width=.45\textwidth]{figures/papi-8}
   }
@@
     \label{fig:papi-16}
-    \includegraphics[width=.31\textwidth]{figures/papi-16}
+    \includegraphics[width=.45\textwidth]{figures/papi-16}
   }
@@
     \label{fig:papi-32}
-    \includegraphics[width=.31\textwidth]{figures/papi-32}
+    \includegraphics[width=.45\textwidth]{figures/papi-32}
   }
@@
     \label{fig:papi-64}
-    \includegraphics[width=.31\textwidth]{figures/papi-64}
+    \includegraphics[width=.45\textwidth]{figures/papi-64}
   }
@@
     \label{fig:packb.conventional}
-    \includegraphics[width=.50\textwidth]{figures/strategy-conventional}
+    \includegraphics[width=.65\textwidth]{figures/strategy-conventional}
   }
@@
     \label{fig:packb.full}
-    \includegraphics[width=.50\textwidth]{figures/strategy-full}
+    \includegraphics[width=.65\textwidth]{figures/strategy-full}
   }
@@
     \label{fig:packb.partial}
-    \includegraphics[width=.50\textwidth]{figures/strategy-partial}
+    \includegraphics[width=.65\textwidth]{figures/strategy-partial}
   }
@@
 In each case, the entire packing workload of $B_2$ is divided into 16 tasks.
-So only the packing workload in one thread, $T_1$, 
+So only the packing workload of one thread, $T_1$, 
 which is highlighted by the z-curves (as shown), is illustrated.
-%% and the task assignment is annoted in Figure~\ref{fig:packb}.
+%% and the task assignment is annotated in Figure~\ref{fig:packb}.
 In the conventional packing approach
@@
 denoted as SCP-P and SCP, respectively,
-are given in Figure~\ref{fig:privb}.
+are both given in Figure~\ref{fig:privb}.
 Figure~\ref{fig:privb.papi} shows the L1 and L2
@@
 and (3) synchronization.
+\textcolor{red}{
+The solid and dashed horizontal lines represent
+the average value of total overhead across matrix sizes
+under SCP and SCP-P, respectively.}
 While behaving similarly as SCP
@@
     \label{fig:privb.papi}
-    \includegraphics[width=.31\textwidth]{figures/privb-papi}
+    \includegraphics[width=.45\textwidth]{figures/privb-papi}
   }
@@
     \label{fig:privb.ate}
-    \includegraphics[width=.31\textwidth]{figures/privb-ate}
+    \includegraphics[width=.45\textwidth]{figures/privb-ate}
   }
@@
     \label{fig:privb.breakdown}
-    \includegraphics[width=.65\textwidth]{figures/privb-breakdown}
+    \includegraphics[width=.90\textwidth]{figures/privb-breakdown}
   }
@@
 However, the improvement in cache performance is small.
-As $B_2$ only occupies a small portion of the shared L2 cache,
+As $B_2$ only occupies a small portion of the shared L2 cache during GEMM execution,
 this marginal improvement is not sufficient to

### 2018-05-28 related.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references

@@
 like multi-sockets and hyperthreading.
+
+\textcolor{red}{
+Because GEMM is highly optimized and parallelized,
+other higher level dense linear algebra operations,
+such as Cholesky, LU and QR factorization,
+can just simply call GEMM to benefit from multi-core and many-core processors.
+In other words, this programming pattern in dense linear algebra software
+harness a fork-and-join style parallelism.
+To reduce synchronization overhead,
+The PLASMA project~\cite{plasma2009,plasma2010,plasma2017}
+utilizes a task-driven parallelization paradigm
+in which dense linear algorithms are formulated
+into directed-acyclic-graphs (DAG) of fine-grained tasks.
+The tasks are dynamically scheduled based on data dependencies
+between them (represented by edges in the DAG).
+The principle of SCP also applies to this DAG based parallelization.
+Input matrices of different tasks can be packed into
+different sets of a shared cache and
+tasks can be scheduled based on the cache sets their input matrices reside in,
+i.e., tasks whose input matrices live in the same cache sets
+are scheduled to the same core in priority.
+}

### 2018-05-28 scp.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references

@@
 $A_2$ and $B_2$ is determined.
-We assume that the share matrix $B_2$ is packed
+We assume that the shared matrix $B_2$ is packed
 in the conventional way-partitioning style.
@@
 \begin{itemize}
-\item The architecture has 2 or 3 levels of caches; %% i.e. $NL=2$ or $NL=3$
-\item All caches are inclusive set-associative caches; %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
-\item The caches at the same level are homogeneous; and %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
+\item The architecture has 2 or 3 levels of caches. %% i.e. $NL=2$ or $NL=3$
+\item All caches are inclusive set-associative caches. %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
+\item The caches at the same level are homogeneous. %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
 \item A cache's way-capacity is a multiple of the sum of way-capacities of its children,
@@
 and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
-are created (line~\ref{line:d.begin} -- \ref{line:d.end}).
+are created (lines~\ref{line:d.begin} -- \ref{line:d.end}).
 

### 2018-05-28 figures/privb.py

@@ -73,20 +73,27 @@ def breakdown():
   # print(ypri)
   fig, ax = plt.subplots()
   ax.set_position([0.1, 0.2, 0.8, 0.75])
+  xmin, xmax = lb-step, ub+step
+  scp_mean = 100-np.mean(scp, axis=0)[-1]
+  pri_mean = 100-np.mean(pri, axis=0)[-1]
 
+  l0 = ax.hlines(scp_mean, xmin, xmax, linestyles='solid', linewidth=2,
+                 color='blue', label='$SCP$ $overhead$')
   scp_o = ax.bar(x0, scp[:, 0], width=width, bottom=0,
                  color='yellow', label='$SCP$ $Sync$')
   scp_a = ax.bar(x0, scp[:, 1], width=width, bottom=yscp[:, 0],
                  color='skyblue', label='$SCP$ $PackA$')
   scp_b = ax.bar(x0, scp[:, 2], width=width, bottom=yscp[:, 1],
                  color='tan', label='$SCP$ $PackB$')
+  l1 = ax.hlines(pri_mean, xmin, xmax, linestyles='dashed', linewidth=2,
+                 color='red', label='$SCP$-$P$ $overhead$')
   pri_o = ax.bar(x1, pri[:, 0], width=width, bottom=0,
                  color='yellow', hatch='//', label='$SCP$-$P$ $Sync$')
   pri_a = ax.bar(x1, pri[:, 1], width=width, bottom=ypri[:, 0],
                  color='skyblue', hatch='//', label='$SCP$-$P$ $PackA$')
   pri_b = ax.bar(x1, pri[:, 2], width=width, bottom=ypri[:, 1],
                  color='tan', hatch='//', label='$SCP$-$P$ $PackB$')
-  ax.set_xlim(lb-step, ub+step)
+  ax.set_xlim(xmin, xmax)
   ax.set_xticks(xticks)
   ax.set_xticklabels(xticks, rotation=45, fontsize='x-large')
   ax.set_xlabel('Matrix Size ($M=N=K$)', fontsize='xx-large')
@@ -96,11 +103,17 @@ def breakdown():
   ax.set_yticks(yticks)
   ax.set_yticklabels(yticklabels, fontsize='x-large')
   ax.set_ylabel('Occupancy Ratio ($\%$)', fontsize='xx-large')
[…]

### 2018-05-29 abstract.tex
Mearge revision from Prof. Xue

@@
   Our evaluation shows that SCP has effectively reduced the
-  conflict misses in both L1 and L2 caches,
-  resulting in an improvement of the GEMM performance by $2.75\%$ -- $6.91\%$.
+  conflict misses in both L1 and L2 caches in a highly-optimized GEMM implementation,
+  resulting in an improvement of its performance by $2.75\%$ -- $6.91\%$.
 \end{abstract}

### 2018-05-29 background.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references
Mearge revision from Prof. Xue

@@
 A classic implementation of the indexing function is shown below,
-%% $\varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)$,
+%% $\varphi(addr) = (addr \gg \log_2(l)) ~\&~ (ns-1)$,
 with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively:
 \begin{equation*}
-  \varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)
+  \varphi(addr) = (addr \gg \log_2(l)) ~\&~ (ns-1)
   \label{eq:phi}
 \end{equation*}
-which represents an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$-th
+which represents an extraction of the $\log_2(ns)$ bits starting from the $\log_2(l)$-th
 position in $addr$'s binary form.
@@
 Throughout this paper, our example platform will be
-a 4-core processor with
+a four-core processor with
 $NL=2$ levels of caches.
@@
   \centering
-  \includegraphics[width=.45\textwidth]{figures/cluster-new}
-  \caption{Memory hierarchy of the 4-core processor
+  \includegraphics[width=.65\textwidth]{figures/cluster-new}
+  \caption{Memory hierarchy of the four-core processor
     (nodes labeled with $(layer,index)$)}
@@
   \centering
-  \caption{Architectural parameters of the 4-core processor}
+  \caption{Architectural parameters of the four-core processor}
   %% (see Section~\ref{subsec:hierarchy} for meanings of symbols in the header row)
@@
 e.g., 8B for a double-precision floating-point number
-and $c_l$ denotes the capacity of the register file or cache at level $l$.
+and $c_l$ ($0\leqslant l \leqslant 3$) denotes the capacity of the register file or cache at level $l$.
 By (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are so
@@
 and a single $B_2$ instance (shared by all threads).
-To understand how the $ii$ loop at layer 3 is parallelized on the 4-core processor, 
+To understand how the $ii$ loop at layer 3 is parallelized on the four-core processor, 
 Figure~\ref{fig:workload} shows the data accessing patterns
@@
   \centering
-  \includegraphics[width=.4\textwidth]{figures/workload}
-  \caption{GEMM parallelization with 4 threads and data accessing patterns of thread $T_1$}
+  \includegraphics[width=.5\textwidth]{figures/workload}
+  \caption{GEMM parallelization with four threads and data accessing patterns of thread $T_1$}
   \label{fig:workload}
@@
 data structures, this repeated prefetching strategy
-can reduce the penalty caused by intra-thread cacheconflicts.
+can reduce the penalty caused by intra-thread cache conflicts.
 
@@
 Suppose two working threads, $T_0$ and $T_1$, are running independently
-on the 4-core processor in Figure.~\ref{fig:hierarchy}
+on the four-core processor in Figure.~\ref{fig:hierarchy}
 to accomplish a GEMM operation.
 At time $t_s$, $T_0$ prefetches some data at address $addr_0$
-to the cache set $\varphi_2(addr_0)$ in a shared L2 cache.
+to the cache set $\varphi_2(addr_0)$ in the shared L2 cache.
 The data will be actually read at time $t_e$.

### 2018-05-29 evaluation.tex
Mearge revision from Prof. Xue

@@
 The 64 cores are organized into 16 clusters with each
-cluster containing 4 cores.
+cluster containing four cores.
 The structure of a cluster is almost the same as
-the 4-core processor in Figure~\ref{fig:hierarchy}
+the four-core processor in Figure~\ref{fig:hierarchy}
 except that the L2 caches of all 16 clusters connect to the main memory
 with hardware coherence.
-\textcolor{red}{
 Table~\ref{tab:cluster} lists the
@@
 in this processor
-and Table~\ref{tab:software} shows the software environment used in out evaluation.}
+and Table~\ref{tab:software} shows the software environment used.
 We restrict our evaluation to DGEMM,
@@
 and (3) synchronization.
-\textcolor{red}{
 The solid and dashed horizontal lines represent
-the average value of total overhead across matrix sizes
-under SCP and SCP-P, respectively.}
+the average values of the total overheads
+across the matrix sizes
+under SCP and SCP-P, respectively.
 While behaving similarly as SCP

### 2018-05-29 intro.tex
Mearge revision from Prof. Xue

@@
 \begin{itemize}
+\item We propose SCP, a method for reducing 
+	the inter-thread cache conflicts
+  for GEMM on architectures with shared non-LRU caches.
 \item We present a quantitative analysis of the negative effect of inter-thread cache
-  conflicts on the GEMM performance.
-\item We propose SCP, a method for solving the inter-thread cache conflicts
-  problem on architectures with shared non-LRU caches.
+  conflicts on the GEMM performance
+  on architectures with shared non-LRU caches.
 \item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
@@
   and shared L2 caches with a pseudo-random
-  replacement policy.
-  Our evaluation shows that SCP can improve the GEMM performance consistently
-  under various parallelization configurations, by $2.75\%$--$6.91\%$.
+  replacement policy. By reducing the inter-thread
+  cache conflicts effectively,
+  SCP can improve the 
+  performance of a highly-optimized
+  GEMM implementation  consistently
+  under various parallelization configurations, by 
+  $2.75\%$ -- $6.91\%$.
 \end{itemize}

### 2018-05-29 related.tex
Mearge revision from Prof. Xue

--- 2018-05-27 related.tex
@@
 like multi-sockets and hyperthreading.
+
+As GEMM is highly optimized and parallelized,
+other higher-level dense linear algebra operations,
+such as Cholesky, LU and QR factorization, can
+enjoy its performance benefits by building themselves
+on top of GEMM.
+In other words, this programming pattern in dense linear algebra software
+harness a fork-and-join style parallelism.
+To reduce synchronization overhead,
+the PLASMA project~\cite{plasma2009,plasma2010,plasma2017}
+makes use of a task-driven parallelization paradigm,
+in which dense linear algorithms are formulated in
+terms of 
+directed-acyclic-graphs (DAGs) of fine-grained tasks.
+The tasks are dynamically scheduled based on their
+data dependencies (represented by the graph edges).
+The basic idea behind SCP also applies to such
+DAG-based parallelization.
+Input matrices of different tasks can be packed into
+different sets of a shared cache. As a result, 
+the tasks can be scheduled based on the cache sets 
+where the input matrices reside in, so that
+the tasks whose input matrices reside 
+in the same cache set
+are scheduled to the same core.

### 2018-05-29 background.tex
Make consistent use of Figure and Table instead of Figure. and Table.

@@
 In other words, each thread packs its own $A_2$ instance but all the threads pack collaboratively the unique $B_2$ instance.
-%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
+%% We will use the OpenBLAS GEMM implementation in Figure~\ref{fig:gemm}
 %% in discussion and evaluation throughout the paper.
@@
 
-In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
+In Figure~\ref{fig:gemm}, the packed matrix $A_2$
 will be read $N_c/N_r$ times within a single GEBP execution,
@@
 Suppose two working threads, $T_0$ and $T_1$, are running independently
-on the four-core processor in Figure.~\ref{fig:hierarchy}
+on the four-core processor in Figure~\ref{fig:hierarchy}
 to accomplish a GEMM operation.

### 2018-05-29 scp.tex
Mearge revision from Prof. Xue
Make consistent use of Figure and Table instead of Figure. and Table.

@@
 Given these tiling factors, the sizes of packed matrices
-can be calculated. Table.~\ref{tab:msizes} lists
+can be calculated. Table~\ref{tab:msizes} lists
 the sizes of various packed matrices used.
@@
     \label{fig:partition.conventional}
-    \includegraphics[width=.4\textwidth]{figures/wpart}
+    \includegraphics[width=.48\textwidth]{figures/wpart}
   }
@@
     \label{fig:partition.segmented}
-    \includegraphics[width=.4\textwidth]{figures/spart}
+    \includegraphics[width=.48\textwidth]{figures/spart}
   }
-  \caption{Partitioning styles for the shared L2 cache (n=16, ns=2048)}
+  \caption{Partitioning styles for a shared L2 cache (n=16 and ns=2048)}
   \label{fig:partition}
@@
 Each $A_2$ is packed into a single continuous buffer.
-Figure.~\ref{fig:partition.conventional} shows how the
+Figure~\ref{fig:partition.conventional} shows how the
 $A_2$ instances of four threads are distributed in the shared 16-way L2 cache.
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \caption{SCP (Phase 1): allocating memory buffers

### 2018-05-30 figures/partitions.py
Colorize figure 4

@@ -17,20 +17,28 @@ colors = ['lightskyblue', 'tan', 'lightgrey', 'pink', 'white']
 hatches = ['/', '..', '\\', 'x', None]
 
 usehatch = True
+usecolor = True
 
 def wpart():
   fig, ax = plt.subplots()
   wpt = 3
+
   for t in range(ntc):
     ways = np.arange(wpt*t, wpt*(t+1))
-    if usehatch:
+    if usehatch and usecolor:
+      ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
+              color=colors[t], hatch=hatches[t], label='$T'+str(t)+'$')
+    elif usehatch:
       ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
               color='None', hatch=hatches[t], label='$T'+str(t)+'$')
     else:
-      ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=0,
+      ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
               color=colors[t], label='$T'+str(t)+'$')
   t = ntc
   ways = np.arange(wpt*ntc, nways)
+  if usehatch and usecolor:
+    ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
+            color=colors[ntc], hatch=hatches[ntc])
   if usehatch:
     ax.barh(ways, wc*np.ones(len(ways)), height=height, left=0, linewidth=.3,
             color='None', hatch=hatches[ntc])
@@ -57,7 +65,10 @@ def spart():
   fig, ax = plt.subplots()
   ways = np.arange(0, 12)
   for t in range(ntc):
-    if usehatch:
+    if usehatch and usecolor:
+      ax.barh(ways, seg*np.ones(len(ways)), height=height, left=t*seg, linewidth=.3,
+              color=colors[t], hatch=hatches[t], label='$T'+str(t)+'$')
+    elif usehatch:
       ax.barh(ways, seg*np.ones(len(ways)), height=height, left=t*seg, linewidth=.3,
               hatch=hatches[t], label='$T'+str(t)+'$', color='None')
     else:
@@ -65,6 +76,9 @@ def spart():
               color=colors[t], label='$T'+str(t)+'$')
   t = ntc
   ways = np.arange(12, nways)
+  if usehatch and usecolor:
[…]

### 2018-08-12 evaluation.tex
Make consistent use of Figure and Table instead of Figure. and Table.
Update after minor revision

@@
 and Table~\ref{tab:software} shows the software environment used.
+\textcolor{red}{
+All caches of the Phytium 2000+ processor are physically indexed,
+i.e., the physical address determines where the data should
+be placed in the cache. In our implementation, physically continuous
+memory buffers (allocated with the \texttt{HUGETLBFS} filesystem
+provided by the Linux kernel) are used,
+so in-cache matrix data of different threads
+are guaranteed to be isolated from each other.
+}
 We restrict our evaluation to DGEMM,
@@
 to partition the cache if $NT_C=1$.
-Figure.~\ref{fig:benefit} shows the results.
+Figure~\ref{fig:benefit} shows the results.
 For comparison purposes, the results without SCP

### 2018-08-12 related.tex
Update after minor revision

@@
 are scheduled to the same core.
+
+\textcolor{red}{
+SCP is a cache partitioning technique.
+While SCP is specific to non-LRU caches and the GEMM program,
+there are general cache partitioning techniques \cite{cp0, cp1, cp2}
+aiming at estimating the interference between multiple applications
+sharing the same cache and improving overall system performance.
+In these research, either new hardware circuit is added to the processor
+or sophisticated cache replacement policies are required.
+SCP is a pure software method working on commercial processors.
+}

### 2018-08-13 response/response.tex
Add response to the minor revision

\documentclass[]{article}

\usepackage{booktabs} % For formal tables
\begin{document}

\title{Response}
\date{}
\author{}

\maketitle

%% We thank all the reviewers for their advices.

\section{Response to Reviewer 1}

\begin{verbatim}
Although the explanation of the problem and the solution is very
clear in general, I think that it would be useful to include an
explanation of how the mapping from virtual to physical addresses
could create additional conflicts in the shared cache, and an analysis
of whether this could happen in the target system or not.
\end{verbatim}

CPU caches can be either virtually indexed or physically indexed.
Nowadays, most processors use physically indexed caches,
that is, the physical address determines where the data should
be placed in the cache.
For physically indexed caches, SCP allocates physically continuous
memory buffers (using the \texttt{HUGETLBFS} filesystem provided
by the Linux kernel)
and partitions the buffer among threads sharing the same cache. 
For virtually indexed caches, virtually continuous
memory buffers (allocated with malloc)
are used instead of physically continuous ones.
Under both situations, in-cache matrix data of different threads
are guaranteed to be isolated from each other
and memory mapping from virtual to physical addresses
would not introduce additional conflicts.

Explanation on this problem has been added to the first paragraph
of section 4 (at the end of page 10).

\section{Response to Reviewer 2}

\begin{verbatim}
Why are the selected matrix sizes (3072 to 4096-6144) chosen?
\end{verbatim}

These sizes are large enough to make GEMM run stably at peak performance
given the total number of threads participating in the computation,
that is, 3072--4096 for 4 threads and 4096--6144 for 8, 16, 32 and 64 threads.

\begin{verbatim}
Which is the behavior of the approach for smaller matrices?
\end{verbatim}

The performance of GEMM on small matrices may show sub-optimal
and unstable performance.
It is common practice for research on GEMM optimization to omit small
matrices as GEMM is meant to run on large ones.
There is another variant of BLAS specially designed for
large amounts of small matrices,
Batched-BLAS, which is out of the scope of our paper.

\begin{verbatim}
Why is there a systematic drop for the largest tested matrices on
Figs. 5 and 6?
\end{verbatim}

There are two kinds of performance drop on Fig.~5 and Fig.~6.

(1) Performance drops as matrix size increases with $NT \ge 16$.
There is a synchronization point at the end of each iteration
of the loop at layer 2 (Fig.~2).
The synchronization is an all-to-all style
communication, whose overhead is proportional to $NT^2$.
As matrix size grows, GEMM suffers more synchronization because
the loop at layer 2 runs more iterations.
When $NT$ is 4 or 8, overhead of this synchronization is small,
but it increases quickly as $NT$ increases.

(2) Performance drops at matrix sizes which are multiple of 1024,
e.g., 3072, 4096, 5120, 6144. 
After profiling, we find that at these sizes,
GEMM suffers more memory contention.
As an example, the table below shows the profiling results
under matrix sizes 3968 and 4096
with parallelism configuration $NT_C=4$ and $NC=4$.
$BW_{A}$ and $BW_{B}$ are measured bandwidth of packing A and B, respectively.
$E_{avg}$ is the floating-point efficiency.
From the table we can see that there is a big performance gap
in $BW_{A}$ and $BW_{B}$.
$E_{avg}$ is also a little lower at 4096 than at 3840,
which may also be caused by the memory contention.
In this example, $NT=NT_C \times NC=16$.
When $M=N=K=4096$, each thread packs a $256 \times 256$ submatrix
from $A$ and a $256 \times 256$ submatrix from $B$,
both of size $V=512KB$.
When $M=N=K=3840$, the submatrix to be packed are of size $V=480KB$.
If the 16 threads starts packing roughly at the same time,
they will access memory addresses
$$addr+0V, addr+1V, \cdots, addr+15V$$
Due to the organization of DDR memory hardware,
$V=512K$ is more likely to cause memory contention than $V=480KB$
because the memory banks are accessed in an imbalanced manner.

\begin{table}[h]
  \centering
  \caption{Profiling results with $NT_C=4$ and $NC=4$}
  \vspace{1em}
  \begin{tabular}{cccc}
    \toprule
         & $BW_A$   & $BW_B$   & $E_{avg}$\\
    \midrule
    3840 & 0.81GB/s & 0.44GB/s & 87.2\%   \\
    4096 & 1.44GB/s & 0.89GB/s & 86.1\%   \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{verbatim}
Is there any evidence that this methodology transparently applies to 
other BLAS-3 routines in OpenBLAS (it should be trivial to test) and
for all matrix layouts (transposed/non transposed)?
\end{verbatim}

As other level-3 routines, for instance, TRSM, TRMM, SYMM and HEMM,
call GEMM internally, they should also benefit from our proposal.
So the SCP methodology can applies transparently to other BLAS-3 routines.

Internally, different matrix layouts (row-major/column-major)
are canonicalized to an implementation defined form
by optionally swapping the $A$ and $B$ operands in $C=\alpha A B+\beta C$.
For example, row-major $C=\alpha A B+\beta C$ can be transformed to
column-major $C^T=\alpha B^T A^T + \beta C^T$.
Then by data packing, matrix orientation information (transposed/non-transposed)
is eliminated out and matrix data is packed into the special
in-memory representation used by the kernel routine, as demonstrated in Fig. 2.
SCP works after these transformations and does not rely on
specific matrix layout or orientations.
So the SCP methodology applies transparently to
all matrix layouts and orientations.

\section{Response to Reviewer 3}

\begin{verbatim}
There is a typo on Page 7, line 43, turning -> tuning.
\end{verbatim}

Already fixed in the paper.

\begin{verbatim}
As the authors mentioned the related work on the tuning implementation
of GEMM only, they should also review the work regarding cache 
partitioning because this paper discussion is mainly on cache
partitioning. Some important papers are: ...
\end{verbatim}

The recommended citations are added to the ``related work`` section.

\begin{verbatim}
The proposed method works well, but it is so specific to DGEMM and
Phytium 2000+ processor. To confirm that their approach is much more
widely applicable even in the context of DGEMM, the discussion when
using the different types of processors with a cache hierarchy for
multi-core environments such as Xeon processors and KNL is strongly
required.
\end{verbatim}

While the Xeon and Xeon Phi processors series
have complex memory hierarchies,
all their caches use an LRU replacement policy.
So they are not target processors of our SCP method,
which must have shared, non-LRU caches.
%% Xeon processors have private, LRU L1 and L2 caches,
%% and a shared LRU last-level-cache (L3).
%% Xeon Phi (KNC, KNL) processors have LRU L1 and L2 caches.
As a result, evaluation on Xeon and Xeon Phi processors would not
improve the confirmability.

The SCP method is specific to GEMM, but its benefits can transparently
propagates to other BLAS-3 operations.
SCP is not specific to the Phytium 2000+ processor and
should work on any processors with shared non-LRU caches.
Unfortunately, we have not get another processor with shared
non-LRU caches like Phytium 2000+.

\end{document}

### 2018-08-13 response/response.tex
Response revised by Prof Xue

@@
 \usepackage{booktabs} % For formal tables
+\usepackage{xcolor}
+
 \begin{document}
 
-\title{Response}
+\title{Author Response}
 \date{}
@@
 
-%% We thank all the reviewers for their advices.
+Thanks to all the reviewers for their comments.
 
@@
 
-CPU caches can be either virtually indexed or physically indexed.
-Nowadays, most processors use physically indexed caches,
-that is, the physical address determines where the data should
-be placed in the cache.
-For physically indexed caches, SCP allocates physically continuous
-memory buffers (using the \texttt{HUGETLBFS} filesystem provided
-by the Linux kernel)
-and partitions the buffer among threads sharing the same cache. 
-For virtually indexed caches, virtually continuous
-memory buffers (allocated with malloc)
-are used instead of physically continuous ones.
-Under both situations, in-cache matrix data of different threads
-are guaranteed to be isolated from each other
-and memory mapping from virtual to physical addresses
-would not introduce additional conflicts.
-
-Explanation on this problem has been added to the first paragraph
-of section 4 (at the end of page 10).
+Added. See the first block of sentences in 
+\textcolor{blue}{blue} 
+in Section 4.
 
@@
 
-These sizes are large enough to make GEMM run stably at peak performance
-given the total number of threads participating in the computation,
-that is, 3072--4096 for 4 threads and 4096--6144 for 8, 16, 32 and 64 threads.
+Added. See the second block of sentences in 
+\textcolor{blue}{blue} 
+in Section 4.
 
@@
 
-The performance of GEMM on small matrices may show sub-optimal
-and unstable performance.
-It is common practice for research on GEMM optimization to omit small
+The performance of GEMM on small matrices may be
+sub-optimal
+and unstable.
+It is common to omit small
 matrices as GEMM is meant to run on large ones.
-There is another variant of BLAS specially designed for
+There is a variant of BLAS specially designed for
 large amounts of small matrices,
-Batched-BLAS, which is out of the scope of our paper.
+Batched-BLAS, which is out of the scope of this paper.
 
@@
 
-There are two kinds of performance drop on Fig.~5 and Fig.~6.
+There are two kinds of performance drops here:
 
-(1) Performance drops as matrix size increases with $NT \ge 16$.
+(1) {\bf Performance drops as the matrix size increases 
+under $NT \ge 16$.}
 There is a synchronization point at the end of each iteration
 of the loop at layer 2 (Fig.~2).
-The synchronization is an all-to-all style
+The synchronization is an all-to-all 
 communication, whose overhead is proportional to $NT^2$.
-As matrix size grows, GEMM suffers more synchronization because
-the loop at layer 2 runs more iterations.
-When $NT$ is 4 or 8, overhead of this synchronization is small,
-but it increases quickly as $NT$ increases.
+As the matrix size increases, GEMM suffers
+from more synchronization because
+the loop at layer 2 runs for more iterations.
+When $NT$ is 4 or 8, the overhead is small but
+increases quickly as $NT$ increases.
 
-(2) Performance drops at matrix sizes which are multiple of 1024,
-e.g., 3072, 4096, 5120, 6144. 
-After profiling, we find that at these sizes,
-GEMM suffers more memory contention.
+(2) {\bf Performance drops at matrix sizes that are
+multiples of 1024.}
+After some  profiling, we found that at these sizes,
+GEMM suffers from more memory contention.
 As an example, the table below shows the profiling results
-under matrix sizes 3968 and 4096
-with parallelism configuration $NT_C=4$ and $NC=4$.
-$BW_{A}$ and $BW_{B}$ are measured bandwidth of packing A and B, respectively.
+for matrix sizes 3968 and 4096
+under $NT_C=4$ and $NC=4$.
+$BW_{A}$ ($BW_{B}$) is the measured bandwidth for 
+packing A (B), and
 $E_{avg}$ is the floating-point efficiency.
-From the table we can see that there is a big performance gap
+We can see that there is a big performance gap
 in $BW_{A}$ and $BW_{B}$.
@@
 both of size $V=512KB$.
-When $M=N=K=3840$, the submatrix to be packed are of size $V=480KB$.
-If the 16 threads starts packing roughly at the same time,
-they will access memory addresses
+When $M=N=K=3840$, the submatrix to be packed is 
+of size $V=480KB$.
+If the 16 threads start packing roughly at the same time,
+they will access the following memory addresses:
 $$addr+0V, addr+1V, \cdots, addr+15V$$
-Due to the organization of DDR memory hardware,
+Due to the organization of the DDR memory hardware,
 $V=512K$ is more likely to cause memory contention than $V=480KB$
-because the memory banks are accessed in an imbalanced manner.
+because the memory banks are accessed in a non-balanced manner.
 
@@
 
-As other level-3 routines, for instance, TRSM, TRMM, SYMM and HEMM,
-call GEMM internally, they should also benefit from our proposal.
-So the SCP methodology can applies transparently to other BLAS-3 routines.
+As other level-3 routines, e.g., TRSM, TRMM, SYMM and HEMM,
+call GEMM internally, they should also benefit from our work.
+So the SCP methodology can apply transparently to 
+these other BLAS-3 routines as well.
 
@@
 by optionally swapping the $A$ and $B$ operands in $C=\alpha A B+\beta C$.
-For example, row-major $C=\alpha A B+\beta C$ can be transformed to
-column-major $C^T=\alpha B^T A^T + \beta C^T$.
-Then by data packing, matrix orientation information (transposed/non-transposed)
-is eliminated out and matrix data is packed into the special
+For example, the row-major $C=\alpha A B+\beta C$ can be transformed to
+the column-major $C^T=\alpha B^T A^T + \beta C^T$.
+Then by means of data packing, the
+matrix orientation information (transposed vs. non-transposed)
+is eliminated and the matrix data are
+packed into the special
 in-memory representation used by the kernel routine, as demonstrated in Fig. 2.
-SCP works after these transformations and does not rely on
-specific matrix layout or orientations.
-So the SCP methodology applies transparently to
-all matrix layouts and orientations.
+
+SCP, which works after these transformations, 
+does not rely on
+specific matrix layouts or orientations.
 
@@
 
-Already fixed in the paper.
+Fixed.
 
@@
 
-The recommended citations are added to the ``related work`` section.
+The recommended references are now cited and
+discussed in Section 5 (highlighted
+in \textcolor{blue}{blue}).
 
@@
 
-While the Xeon and Xeon Phi processors series
-have complex memory hierarchies,
-all their caches use an LRU replacement policy.
-So they are not target processors of our SCP method,
-which must have shared, non-LRU caches.
-%% Xeon processors have private, LRU L1 and L2 caches,
-%% and a shared LRU last-level-cache (L3).
-%% Xeon Phi (KNC, KNL) processors have LRU L1 and L2 caches.
-As a result, evaluation on Xeon and Xeon Phi processors would not
-improve the confirmability.
+For Xeon and Xeon Phi processors, the caches with
+the LRU replacement policy are used.
+However,
+SCP is designed for processors with non-LRU caches.
 
-The SCP method is specific to GEMM, but its benefits can transparently
-propagates to other BLAS-3 operations.
+SCP works for all other BLAS-3 routines, e.g., TRSM, TRMM, SYMM and HEMM, that call GEMM internally.
 SCP is not specific to the Phytium 2000+ processor and
-should work on any processors with shared non-LRU caches.
-Unfortunately, we have not get another processor with shared
-non-LRU caches like Phytium 2000+.
+should work for any processors with shared non-LRU caches.
 

### 2018-08-13 evaluation.tex
Response revised by Prof Xue

--- 2018-05-29 evaluation.tex
@@
 and Table~\ref{tab:software} shows the software environment used.
+\textcolor{blue}{
+The L2 caches in the Phytium 2000+ processor are physically indexed.
+In our implementation, physically continuous
+memory buffers (allocated by the \texttt{HUGETLBFS} filesystem
+provided by the Linux kernel) are used for storing
+matrices in packed form. Therefore,
+for a given shared L2 cache,
+the matrix data accessed by different threads 
+are guaranteed to be free of conflict cache misses.
+}
 We restrict our evaluation to DGEMM,
@@
 %% \end{equation*}
+
+\textcolor{blue}{
+The mairix sizes selected range from 3072 to 
+4096 when $NT \in \{4, 8\}$ and to 
+6144 when $NT \in \{16, 32, 64\}$. 
+These sizes are large enough to make GEMM run stably at peak performance
+the given the total number of threads participating in the computation.
+}
 

### 2018-08-13 related.tex
Response revised by Prof Xue

--- 2018-05-29 related.tex
@@
 are scheduled to the same core.
+
+\textcolor{blue}{
+SCP is a cache partitioning technique.
+While SCP is specific to GEMM running on architectures
+with non-LRU caches,
+there are general cache partitioning techniques \cite{cp0, cp1, cp2}
+that focus on estimating the interference across 
+the multiple applications
+sharing the same cache and improving the
+overall system performance.
+In these research efforts, either hardware modifications are required
+or some sophisticated cache replacement policies are 
+needed.
+In contrast,
+SCP is a pure software approach that works for 
+commercial processors.
+}

### 2018-08-13 response/response.tex
Remove colored highlights

@@
 
-Added. See the first block of sentences in 
-\textcolor{blue}{blue} 
-in Section 4.
+Added. See the new sentences added in lines -2 -- -1 in
+page 10 and lines 1 -- 3 in Page 11 (Section 4).
 
@@
 
-Added. See the second block of sentences in 
-\textcolor{blue}{blue} 
-in Section 4.
+Added. See the 2nd last paragraph added just before
+Section 4.1.
 
@@
 The recommended references are now cited and
-discussed in Section 5 (highlighted
-in \textcolor{blue}{blue}).
+discussed in the last paragraph of Section 5.
 

### 2018-08-13 evaluation.tex
Remove colored highlights

@@
 and Table~\ref{tab:software} shows the software environment used.
-\textcolor{blue}{
 The L2 caches in the Phytium 2000+ processor are physically indexed.
@@
 are guaranteed to be free of conflict cache misses.
-}
 We restrict our evaluation to DGEMM,
@@
 
-\textcolor{blue}{
 The mairix sizes selected range from 3072 to 
@@
 the given the total number of threads participating in the computation.
-}
 

### 2018-08-25 evaluation.tex
Add content from response letter to the final version as the editor requests

@@
 
+{\color{red}{
+In Figure~\ref{fig:drawback} and \ref{fig:benefit},
+The performance drops slightly as the matrix size increases 
+under $NT \ge 16$.
+The reason is that there is a synchronization point
+at the end of each iteration
+of the loop at layer 2 (Fig.~2).
+The synchronization is an all-to-all 
+communication, whose overhead is proportional to $NT^2$.
+When $NT$ is 4 or 8, the overhead is small but
+it increases quickly as $NT$ increases.
+Performance also drops at matrix sizes that are
+multiples of 1024 because
+GEMM suffers from more memory contention at these sizes,.
+As an example, we profiled the GEMM execution
+at matrix sizes 3968 and 4096 under $NT_C=4$ and $NC=4$.
+The measured bandwidth for packing A (B) are
+1.44(0.89) GB/s at size 3968 and 0.81(0.44) GB/s at size 4096.
+We can see that there is a big performance gap
+in memory bandwidth.
+Due to the organization of the DDR memory hardware,
+The selected tiling factors $M_c$, $N_c$ and $K_c$
+made GEMM more likely to cause memory contention at sizes
+that are multiples of 1024
+because the memory banks are accessed in a non-balanced manner.
+
+In GEMM implementation, different matrix layouts (row-major/column-major)
+are first canonicalized to an implementation defined form
+by optionally swapping the $A$ and $B$ operands in $C=\alpha A B+\beta C$.
+And matrix orientation information (transposed vs. non-transposed)
+is eliminated by means of data packing.
+SCP, which works after these transformations, 
+does not rely on specific matrix layouts or orientations.
+
+As other level-3 routines, e.g., TRSM, TRMM, SYMM and HEMM,
+call GEMM internally, they should also benefit from SCP,
+i.e., the SCP methodology can apply transparently to 
+other BLAS-3 routines as well.
+}}
+
 \subsection{Cache Miss Rate Analysis}\label{subsec:analysis}

### 2018-08-27 evaluation.tex
Revision from Prof. Xue before final submission

--- 2018-08-13 evaluation.tex
@@
 
+In Figures~\ref{fig:drawback} and \ref{fig:benefit},
+the performance drops slightly as the matrix size increases 
+when $NT \ge 16$. Let us explain
+the reason behind. There is a synchronization point
+at the end of each iteration
+of the loop at layer 2 (Figure~2).
+This synchronization is an all-to-all 
+communication, whose overhead is proportional to $NT^2$.
+This overhead is small when $NT$ is 4 or 8 but
+increases quickly as $NT$ increases. In addition, the
+performance also drops when matrix sizes are 
+multiples of 1024, as
+GEMM suffers from more memory contention at these matrix
+sizes. 
+To understand this better,
+we have profiled the GEMM execution
+for the matrix sizes 3968 and 4096 under $NT_C=4$ and $NC=4$.
+The measured bandwidth for packing A (B) are
+1.44 (0.89) GB/s for the
+matrix size 3968 but drops significantly to 
+0.81 (0.44) GB/s for the matrix size 4096. 
+As the memory banks in the Phytium 2000+ processor 
+are accessed in a non-balanced manner,
+the tiling factors $M_c$, $N_c$ and $K_c$ selected
+make it more likely for GEMM to suffer from  
+memory contention at matrix sizes that are multiples of 1024.
+
+%In GEMM implementation, different matrix layouts (row-major/column-major)
+%are first canonicalized to an implementation defined form
+%by optionally swapping the $A$ and $B$ operands in $C=\alpha A B+\beta C$.
+%And matrix orientation information (transposed vs. non-transposed)
+%is eliminated by means of data packing.
+%SCP, which works after these transformations, 
+%does not rely on specific matrix layouts or orientations.
+
 \subsection{Cache Miss Rate Analysis}\label{subsec:analysis}

### 2018-08-27 intro.tex
Revision from Prof. Xue before final submission

@@
 the first work addressing
-the inter-thread cache conflicts problem on
+the inter-thread cache conflicts problem for GEMM on
 architectures with shared non-LRU caches.
+As other level-3 routines, such as TRSM, TRMM, SYMM and HEMM,
+call GEMM internally, they should also benefit from SCP,
+i.e., the SCP methodology can apply transparently to these
+BLAS-3 routines as well.
 
